### 2014-02-18 paper/paper.tex
will now work to resolve multi-node issue

\documentclass{article}

% The geometry package allows for easy page formatting.
\usepackage{geometry}
\geometry{letterpaper}

% Load up special logo commands.
\usepackage{doc}

% Package for formatting URLs.
\usepackage{url}

% Packages and definitions for graphics files.
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{natbib}
\usepackage{tikz}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

%
% Set the title, author, and date.
%
\title{Panthera: A Study of Caching in Distributed Computing}
\author{Dhaivat Pandya}
\date{}

%
% The document proper.
%
\begin{document}

% Add the title section.
\maketitle

% Add an abstract.
\abstract{
Describe your paper in 100-200 words, give or take.  The command-line \texttt{wc} utility is really useful here!  This particular sample paper is meant to demonstrate a variety of \LaTeX\ directives for producing a well-structured, consistently-formatted scholarly document.  The actual content and outline may vary according to the needs of your specific research topic.
}

\section{Introduction}
The Hadoop distributed system \cite{hadoop} is an open source version of the revolutionary MapReduce system developed at Google. With it, developers can take easily take advantage of large, multi-node clusters to solve computational problems.

Latency in distributed systems can have significant effects, and a reduction of the same comes with tremendous benefits. As the RAMCloud project has outlined \cite{ramcloud}, low latency can greatly extend the applications of distributed computing systems.

In this paper, we discuss \textit{Panthera}, a cache layer for Hadoop.

\section{Constraints}

There are several systems built on Hadoop that are in widespread use \cite{hbase, cloudbatch, pig}. Thus, for \textit{Panthera} to be practical, it must operate independently of the existing Hadoop codebase. Additionally, for non-cache related requests, \textit{Panthera} must add insignificant latency. Finally, data and metadata request latency should be significantly with \textit{Panthera} in comparison to a vanilla Hadoop installation.

\section{Architecture}
\begin{figure}
\begin{center}
	\begin{tikzpicture}
		\begin{tikzpicture}[scale=1.5, transform shape]
		\draw[style=solid] (0,0) circle (1) node {DataNode 1};
		\draw[style=dashed] (0, 0) circle(1.3) node {DataNode 1};
		\draw[style=solid] (3.0, 2.5) node {NameNode}  circle (1);
		\draw[style=solid] (3.0, 0) node{DataNode 2} circle (1);
		\draw[style=solid] (6.0, 0) node{DataNode 3} circle (1);
		\draw[loosely dashed, line width=2pt]  (2.5,0.3)-- (2.5, 2.3);
		\draw[loosely dashed, line width=2pt]  (0, 0.3) -- (2.5, 2.3);
		\draw[loosely dashed, line width=2pt]  (5.0, 0.3) -- (2.5, 2.3);
	\end{tikzpicture}
	\end{tikzpicture}
\end{center}
\end{figure}

\textit{Panthera} runs as a standalone program on every slave node (usually also a DataNode) of a given Hadoop Distributed File System cluster.

\bibliographystyle{plain}

\bibliography{references}

\end{document}

### 2014-02-18 analysis/box-plot.r

@@ -4,10 +4,10 @@ dat <- read.csv("latencies.csv", head = TRUE, sep=",")
 cached_data = unlist(na.omit(dat[1]))
 noncached_data = unlist(na.omit(dat[2]))
 
-df <- data.frame(cached = cached_data, noncached = noncached_data)
-png("boxplot.png")
+df <- data.frame(metacached = cached_data, non_metacached = noncached_data)
 pdf("boxplot.pdf")
-boxplot(df, ylim=c(0, 8000000))
-title("Panthera (cached) vs Status Quo (non-cached)")
+boxplot(df, ylim=c(0, 40000000), col=c('cornflowerblue', 'chocolate1'), xaxt="n")
+title("Metacached (Panthera) vs Non-metacached")
+axis(1, at=1:2, labels=c("metacached (panthera)", "non-metacached"))
 title(ylab = "Access Time (nanoseconds, lower is better)")
 dev.off()

### 2014-02-18 analysis/histogram.r

@@ -4,6 +4,6 @@ cached_latency = unlist(na.omit(dat[1]))
 no_cached_latency = unlist(na.omit(dat[2]))
 
 pdf("histogram.pdf")
-hist(no_cached_latency, breaks=1000, col='black', xlim=c(0, 16000000), ylim=c(0, 600))
-hist(cached_latency, breaks=100, col='green', add=T)
+hist(no_cached_latency, breaks=1000, col='chocolate1', xlim=c(0, 16000000), ylim=c(0, 250))
+hist(cached_latency, breaks=100, col='cornflowerblue', add=T)
 dev.off()

### 2014-02-18 analysis/point-plot.r

@@ -1,0 +1,10 @@
+dat <- read.csv("latencies.csv")
+
+cached_latency = unlist(na.omit(dat[1]))
+no_cached_latency = unlist(na.omit(dat[2]))
+
+pdf("point-plot.pdf")
+plot(cached_latency, col="dodgerblue2", pch=1, cex = 1.5, lwd=2)
+title("Metacached (Panthera) vs Non-metacached")
+points(no_cached_latency, col="chocolate1", pch=1, cex = 1.5, lwd=2)
+dev.off()

### 2014-02-18 src/caches/get_file_info_cache.go
fixed part of the issue

@@ -6,7 +6,9 @@ to call, so the cache layer can respond instead of having to go to the server.
 */
 
 import (
+	//local packages
 	"namenode_rpc"
+	"util"
 )
 
 type GetFileInfoCache struct {
@@ -36,6 +38,7 @@ func (gfi_cache *GetFileInfoCache) Disable() {
 //Query the cache. Returns nil if req is not found in the cache or the Enabled is set to false.
 //gfi_cache.Cache.Query should NOT be called since it suses the wrong kind of equality comparator
 func (gfi_cache *GetFileInfoCache) Query(req namenode_rpc.ReqPacket) namenode_rpc.ResponsePacket {
+	util.DebugLogger.Println("in GetFileInfoCache.Query()")
 	//if the cache is not enabled, we keep returning nil
 	if !gfi_cache.enabled {
 		return nil
@@ -55,11 +58,15 @@ func (gfi_cache *GetFileInfoCache) Query(req namenode_rpc.ReqPacket) namenode_rp
 		return false
 	})
 
+	util.DebugLogger.Println("Defined Equals method.")
+
 	res := gfi_cache.Cache.QueryCustom(req, equals)
+	util.DebugLogger.Println("Done querying cache.")
 
 	if res == nil {
 		return nil
 	}
+
 	return res
 }
 

### 2014-02-18 src/data_requests/data_req_processor.go

@@ -330,11 +330,16 @@ func (p *Processor) RecordNoCacheLatency() {
 	util.NoCacheLatencyLog.Println(duration.Nanoseconds())
 }
 
-func (p *Processor) handleDataResponse(conn net.Conn, dataNode net.Conn) {
+func (p *Processor) handleDataResponse(conn net.Conn, dataNode net.Conn) error {
 	util.DebugLogger.Println("current request: ", p.currentRequest)
 	dataResponse := datanode_rpc.NewDataResponse()
 	err := dataResponse.LiveLoad(dataNode)
 	byt, _ := dataResponse.Bytes()
+	if len(byt) == 0 {
+		util.DebugLogger.Println("Bytes length is 0; returning from HandleDataResponse()")
+		return errors.New("bytes read = 0")
+	}
+
 	util.DebugLogger.Println("Data response bytes len: ", len(byt))
 	util.DebugLogger.Println("dataResponse.DataLength: ", dataResponse.DataLength)
 	util.DebugLogger.Println("dataResponse.data bytes: ", dataResponse.Data)
@@ -345,7 +350,7 @@ func (p *Processor) handleDataResponse(conn net.Conn, dataNode net.Conn) {
 		if !res {
 			util.DataReqLogger.Println("Could not retry successfully. Returning from HandleDataNode")
 			//connections were closed, no point in continuing with the processor
-			return
+			return nil
 		}
 	}
 	
@@ -356,7 +361,7 @@ func (p *Processor) handleDataResponse(conn net.Conn, dataNode net.Conn) {
 		if p.skipResponse {
 			util.DataReqLogger.Println("Response skipped.")
 			p.skipResponse = false
-			return
+			return nil
 		}
 
 		pair := datanode_rpc.NewRequestResponse(p.currentRequest, dataResponse)
@@ -372,6 +377,8 @@ func (p *Processor) handleDataResponse(conn net.Conn, dataNode net.Conn) {
 		}
 		util.DataReqLogger.Println(p.id, "Wrote response to client.")
 	}
+
+	return nil
 }
 
 func (p *Processor) handlePutDataResponse(conn net.Conn, dataNode net.Conn) {
@@ -397,9 +404,9 @@ func (p *Processor) handlePutDataResponse(conn net.Conn, dataNode net.Conn) {
 	}
 }
 
-func (p *Processor) loadResponse(conn net.Conn, dataNode net.Conn) {
[…]

### 2014-02-18 src/hdfs_requests/processor.go

@@ -14,6 +14,8 @@ import (
 	"strings"
 	"strconv"
 	"io"
+	"encoding/hex"
+	"reflect"
 
 	//used for the the DataNodeMap
 	"configuration"
@@ -27,6 +29,8 @@ type PacketNumber uint32
 var TIMECOUNTER time.Time
 
 type Processor struct {
+	currentRequest *namenode_rpc.RequestPacket
+
 	//array of the past requests received by this processor
 	//presumably by the same client (i.e. one client per processor)
 	PastRequests []namenode_rpc.RequestPacket
@@ -407,6 +411,8 @@ func (p *Processor) readRequestPacket(conn net.Conn) (*namenode_rpc.RequestPacke
 	reqPacket := namenode_rpc.NewRequestPacket()
 	reqPacket.Load(finalBuf)
 
+	p.currentRequest = reqPacket
+
 	return reqPacket, nil
 }
 
@@ -440,10 +446,15 @@ func (p *Processor) HandleRequestPacket(conn net.Conn, hdfs net.Conn) error {
 	reqPacket, modified := p.preprocessRequestPacket(reqPacket)
 
 	//check the cache and write the corresponding request
+	util.DebugLogger.Println("Calling process method...")
 	respPacket := p.Process(reqPacket)
+	util.DebugLogger.Println("Done with process method, respPacket.")
+
 	//hit in the cache
 	if respPacket != nil {
-		conn.Write(respPacket.Bytes())
+		util.DebugLogger.Println("Writing resp packet, type: ", reflect.TypeOf(respPacket), " bytes: \n", hex.Dump(respPacket.GetBuf()))
+		conn.Write(respPacket.GetBuf())
+		util.DebugLogger.Println("Done writing resp packet, bytes") 
 	} else {
 		p.CacheRequest(reqPacket)
 	}
@@ -496,6 +507,7 @@ func (p *Processor) HandleConnectionReimp(conn net.Conn, hdfs net.Conn) {
 		//if it is the second packet, it is the authentication packet
 		//after that, process them as request packets
 		default:
+			util.DebugLogger.Println("Starting to handle request packet...")
 			err := p.HandleRequestPacket(conn, hdfs)
 			if err != nil {
 				util.DebugLog("Error handling request packet.")
@@ -677,6 +689,7 @@ func (p *Processor) HandleHDFS(conn net.Conn, hdfs net.Conn) {
 			if readErr == io.EOF {
[…]

### 2014-02-18 src/main/main.go

@@ -112,6 +112,7 @@ func runDataNodeMap(dataNodeMap configuration.DataNodeMap, cache *caches.DataCac
 		listener, err := net.Listen("tcp", ":" + string(port))
 		if err != nil || listener == nil {
 			log.Println("Could not listen on relay port:", port, " because: ", err.Error(), listener, location)
+			util.DebugLogger.Println("Could not listen on relay port: ", port, " because: ", err.Error(), listener, location)
 		}
 
 		fmt.Println("Listener: ", listener)
@@ -143,7 +144,7 @@ func main() {
 	cacheSet.GetListingCache = caches.NewGetListingCache(getListingCacheSize)
 	
 	//disable the metadata cache for now
-	cacheSet.Disable()
+	//cacheSet.Disable()
 
 	server, err := net.Listen("tcp", config.serverHost + ":" + config.serverPort)
 	log.SetOutput(ioutil.Discard)

### 2014-02-18 src/namenode_rpc/namenode_rpc.go

@@ -65,11 +65,31 @@ type ResponsePacket interface {
 
 	//common fields that need to have a Getter
 	GetPacketNumber() uint32
+
+	//return the bytes that were used to load a field
+	GetBuf() []byte
 }
 
 type ResponseObject struct {
 	NameLength uint16
 	ObjectName []byte
+}
+
+//this method figures out, given a request, what kind of new packet to build
+func BuildResponsePacket(buf []byte, packetNumber uint32, currRequest *RequestPacket) ResponsePacket {
+	if currRequest == nil {
+		return NewGenericResponsePacket(buf, packetNumber)
+	}
+
+	switch string(currRequest.MethodName) {
+	case "getFileInfo":
+		gfir := NewGetFileInfoResponse()
+		gfir.Load(buf)
+		return gfir
+	default:
+		return NewGenericResponsePacket(buf, packetNumber)
+	}
+
 }
 
 //this is used when we aren't exactly sure (or don't care)
@@ -90,6 +110,10 @@ type GenericResponsePacket struct {
 
 	ParameterLength uint16
 	ParameterValue []byte
+}
+
+func (grp *GenericResponsePacket) GetBuf() []byte {
+	return grp.Buf
 }
 
 func NewGenericResponsePacket(buf []byte, packNum uint32) *GenericResponsePacket {
@@ -406,6 +430,10 @@ type GetFileInfoResponse struct {
 
 	//the buf passed to the Load() call
 	LoadedBytes []byte
+}
+
+func (gfir *GetFileInfoResponse) GetBuf() []byte {
+	return gfir.LoadedBytes
 }
 
 func NewGetFileInfoResponse() *GetFileInfoResponse {

### 2014-02-18 src/util/logging.go

@@ -4,23 +4,24 @@ import (
 	"fmt"
 	"log"
 	"os"
+	"io/ioutil"
 )
 
 //this logger is handled by the DataReqLogger
 var DataReqLogger *log.Logger
-var DataReqLogFile = "/home/dhaivat/dev/hadoopproxy/logs/data_req_log"
+var DataReqLogFile = "/home/dhaivat/dev/panthera/logs/data_req_log"
 
 var LoggingEnabled bool = true
 var DebugLoggingEnabled bool = true
 
 var NoCacheLatencyLog *log.Logger
-var NoCacheLatencyLogFile = "/home/dhaivat/dev/hadoopproxy/logs/no_cache_latency"
+var NoCacheLatencyLogFile = "/home/dhaivat/dev/panthera/logs/no_cache_latency"
 
 var CachedLatencyLog *log.Logger
-var CachedLatencyLogFile = "/home/dhaivat/dev/hadoopproxy/logs/cached_latency"
+var CachedLatencyLogFile = "/home/dhaivat/dev/panthera/logs/cached_latency"
 
 var DebugLogger *log.Logger
-var DebugLogFile = "/home/dhaivat/dev/hadoopproxy/logs/debug.log"
+var DebugLogFile = "/home/dhaivat/dev/panthera/logs/debug.log"
 
 //NOTE this is a relative path; in deployment, the executable needs to be in the same directory
 //this file otherwise there will be problems in loading the logging configuration (it will
@@ -29,8 +30,12 @@ var LoggingConfFile = "logging.json"
 
 //load, with a configuration file, the places where the logs are 
 //supposed to be going
-func InitLoggingConfiguration() {
-	loggingContents := string(ioutil.ReadFile(LoggingConfFile))
+func InitLoggingConfiguration() error {
+	_, err := ioutil.ReadFile(LoggingConfFile)
+	if err != nil {
+		return err
+	}
+	return nil
 }
 
 func InitDataReqLogger() error {

### 2014-02-18 src/hdfs_requests/processor.go
gfi cache responding correctly

@@ -62,6 +62,9 @@ type Processor struct {
 
 	//we need the datanode map to make replacements in block reports
 	dataNodeMap *configuration.DataNodeMap
+
+	//skip the response if it has already been filled by the cache
+	skipResponse bool
 }
 
 //Object constructor. It needs the event_chan to talk with other processors,
@@ -77,6 +80,7 @@ func NewProcessor(event_chan chan ProcessorEvent, cacheSet *caches.CacheSet,
 	p.cacheSet = cacheSet
 	p.HandledFirstPacket = false
 	p.PacketsProcessed = 0
+	p.skipResponse = false
 
 	cacheLogFile, err := os.OpenFile("cache_times", os.O_RDWR | os.O_APPEND | os.O_CREATE, 0666)
 	if err != nil {
@@ -452,6 +456,7 @@ func (p *Processor) HandleRequestPacket(conn net.Conn, hdfs net.Conn) error {
 
 	//hit in the cache
 	if respPacket != nil {
+		p.skipResponse = true
 		util.DebugLogger.Println("Writing resp packet, type: ", reflect.TypeOf(respPacket), " bytes: \n", hex.Dump(respPacket.GetBuf()))
 		conn.Write(respPacket.GetBuf())
 		util.DebugLogger.Println("Done writing resp packet, bytes") 
@@ -739,7 +744,11 @@ func (p *Processor) HandleHDFS(conn net.Conn, hdfs net.Conn) {
 
 			//BREAAAAAAAAAAAAAAAAAAAAKKKKKINNNNNGGGG CHANGE
 			util.DebugLogger.Println("About to write to connection...")
-			conn.Write(genericResp.GetBuf())
+			if !p.skipResponse {
+				conn.Write(genericResp.GetBuf())
+			} else {
+				p.skipResponse = false
+			}
 			util.DebugLogger.Println("Written to connection.")
 		}
 	}

### 2014-02-18 src/namenode_rpc/namenode_rpc.go

@@ -81,6 +81,8 @@ func BuildResponsePacket(buf []byte, packetNumber uint32, currRequest *RequestPa
 		return NewGenericResponsePacket(buf, packetNumber)
 	}
 
+	return NewGenericResponsePacket(buf, packetNumber)
+	/*
 	switch string(currRequest.MethodName) {
 	case "getFileInfo":
 		gfir := NewGetFileInfoResponse()
@@ -88,7 +90,7 @@ func BuildResponsePacket(buf []byte, packetNumber uint32, currRequest *RequestPa
 		return gfir
 	default:
 		return NewGenericResponsePacket(buf, packetNumber)
-	}
+	} */
 
 }
 

### 2014-02-19 src/util/logging.go
logging meta cache latency times

@@ -9,19 +9,25 @@ import (
 
 //this logger is handled by the DataReqLogger
 var DataReqLogger *log.Logger
-var DataReqLogFile = "/home/dhaivat/dev/panthera/logs/data_req_log"
+var DataReqLogFile = "../../logs/data_req_log"
 
 var LoggingEnabled bool = true
 var DebugLoggingEnabled bool = true
 
 var NoCacheLatencyLog *log.Logger
-var NoCacheLatencyLogFile = "/home/dhaivat/dev/panthera/logs/no_cache_latency"
+var NoCacheLatencyLogFile = "../../logs/no_cache_latency"
 
 var CachedLatencyLog *log.Logger
-var CachedLatencyLogFile = "/home/dhaivat/dev/panthera/logs/cached_latency"
+var CachedLatencyLogFile = "../../logs/cached_latency"
 
 var DebugLogger *log.Logger
-var DebugLogFile = "/home/dhaivat/dev/panthera/logs/debug.log"
+var DebugLogFile = "../../logs/debug.log"
+
+var MetaCachedLatencyLogger *log.Logger
+var MetaCachedLatencyLogFile = "../../logs/meta_cache_latency"
+
+var NonMetaCachedLatencyLogger *log.Logger
+var NonMetaCachedLatencyLogFile = "../../logs/non_meta_cache_latency"
 
 //NOTE this is a relative path; in deployment, the executable needs to be in the same directory
 //this file otherwise there will be problems in loading the logging configuration (it will
@@ -77,8 +83,28 @@ func InitCachedLatencyLogger() error {
 
 	CachedLatencyLog = log.New(cachedLogFile, "", 0)
 	
-	//write the units for the data
-	CachedLatencyLog.Println("(latency times in nanoseconds)")
+	return nil
+}
+
+func InitNonMetaCachedLatencyLogger() error {
+	file, err := os.OpenFile(NonMetaCachedLatencyLogFile, os.O_RDWR | os.O_CREATE | os.O_APPEND, 0666)
+	if err != nil {
+		return err
+	}
+
+	NonMetaCachedLatencyLogger = log.New(file, "", 0)
+	
+	return nil
+}
+
+func InitMetaCachedLatencyLogger() error {
+	metaCacheLogFile, err := os.OpenFile(MetaCachedLatencyLogFile, os.O_RDWR | os.O_CREATE | os.O_APPEND, 0666)
[…]

### 2014-02-19 src/main/main.go
piling commit

@@ -57,7 +57,7 @@ func loop(server net.Listener, caches *caches.CacheSet, dnMap *configuration.Dat
 		util.DebugLog("Connecting to HDFS host: " + string(config.hdfsHostname) + ":" + string(config.hdfsPort))
 		hdfs, hdfs_err := net.Dial("tcp", config.hdfsHostname + ":" + config.hdfsPort)
 		if hdfs_err != nil {
-			util.LogError(err.Error())
+			util.LogError(hdfs_err.Error())
 			continue
 		}
 		util.DebugLog("Dialed HDFS...")
@@ -128,8 +128,8 @@ func main() {
 	util.LoggingEnabled = false
 	util.Log("Starting...")
 
-	config.hdfsHostname = "127.0.0.1"
-	config.hdfsPort = "1102"
+	config.hdfsHostname = "162.243.105.47"
+	config.hdfsPort = "54310"
 
 	config.serverHost = "0.0.0.0"
 	config.serverPort = "1035"

### 2014-02-19 src/util/logging.go

@@ -9,19 +9,19 @@ import (
 
 //this logger is handled by the DataReqLogger
 var DataReqLogger *log.Logger
-var DataReqLogFile = "/home/dhaivat/dev/panthera/logs/data_req_log"
+var DataReqLogFile = "/root/panthera/logs/data_req_log"
 
 var LoggingEnabled bool = true
 var DebugLoggingEnabled bool = true
 
 var NoCacheLatencyLog *log.Logger
-var NoCacheLatencyLogFile = "/home/dhaivat/dev/panthera/logs/no_cache_latency"
+var NoCacheLatencyLogFile = "/root/panthera/logs/no_cache_latency"
 
 var CachedLatencyLog *log.Logger
-var CachedLatencyLogFile = "/home/dhaivat/dev/panthera/logs/cached_latency"
+var CachedLatencyLogFile = "/root/panthera/logs/cached_latency"
 
 var DebugLogger *log.Logger
-var DebugLogFile = "/home/dhaivat/dev/panthera/logs/debug.log"
+var DebugLogFile = "/root/panthera/logs/debug.log"
 
 //NOTE this is a relative path; in deployment, the executable needs to be in the same directory
 //this file otherwise there will be problems in loading the logging configuration (it will

### 2014-02-19 src/hdfs_requests/processor.go
latency calculation

@@ -460,7 +460,7 @@ func (p *Processor) HandleRequestPacket(conn net.Conn, hdfs net.Conn) error {
 
 	//preprocess the request
 	//reqPacket, _ = p.Preprocess(reqPacket)
-
+	p.cacheStartTime = &t
 	reqPacket, modified := p.preprocessRequestPacket(reqPacket)
 
 	//check the cache and write the corresponding request
@@ -533,7 +533,6 @@ func (p *Processor) HandleConnectionReimp(conn net.Conn, hdfs net.Conn) {
 		default:
 			util.DebugLogger.Println("Starting to handle request packet...")
 			t := time.Now()
-			p.cacheStartTime = &t
 			err := p.HandleRequestPacket(conn, hdfs)
 			if err != nil {
 				util.DebugLog("Error handling request packet.")

### 2014-02-19 src/hdfs_requests/processor.go
working on latency issues

@@ -12,7 +12,6 @@ import (
 	"time"
 	"os"
 	"strings"
-	"strconv"
 	"io"
 	"encoding/hex"
 	"reflect"
@@ -118,13 +117,15 @@ func NewProcessor(event_chan chan ProcessorEvent, cacheSet *caches.CacheSet,
 func (p *Processor) CacheRequest(req *namenode_rpc.RequestPacket) {
 	//fmt.Println("Method name from CacheRequest(): ", string(req.MethodName))
 
+	/*
 	if(string(req.MethodName) == "getFileInfo") {
 		//follow through with the GetFileInfoCache
 		fmt.Println("Caching request..., cache size: ", len(p.cacheSet.GfiCache.Cache.RequestResponse))
 		p.cacheSet.GfiCache.Cache.AddRequest(req)
 		fmt.Println("Cached request. Cache size: ", len(p.cacheSet.GfiCache.Cache.RequestResponse))
 		log.Println("Cached GFI Request: ")
-	} else if(string(req.MethodName) == "getListing") {
+	} else */
+	if(string(req.MethodName) == "getListing") {
 		p.cacheSet.GetListingCache.Cache.AddRequest(req)
 	}
 }
@@ -391,10 +392,8 @@ func (p *Processor) readRequestPacketLength(conn net.Conn) (uint32, []byte, erro
 	} */
 	//return packetLength, readErr
 	byteBuffer := bytes.NewBuffer(buf)
-	fmt.Println("read request packet buf: ", buf, byteBuffer.Bytes(), string(byteBuffer.Bytes()))
 	binary.Read(byteBuffer, binary.BigEndian, &packetLength)
 
-	util.DebugLog("Read request packet length: " + strconv.FormatInt(int64(packetLength), 10))
 	return packetLength, buf, readErr	
 }
  
@@ -418,6 +417,7 @@ func (p *Processor) readRequestPacket(conn net.Conn) (*namenode_rpc.RequestPacke
 
 	reqPacket := namenode_rpc.NewRequestPacket()
 	reqPacket.Load(finalBuf)
+	fmt.Println("Read request with method name: ", string(reqPacket.MethodName))
 
 	p.currentRequest = reqPacket
 
@@ -460,8 +460,10 @@ func (p *Processor) HandleRequestPacket(conn net.Conn, hdfs net.Conn) error {
 
 	//preprocess the request
 	//reqPacket, _ = p.Preprocess(reqPacket)
-	p.cacheStartTime = &t
[…]

### 2014-02-24 poster/paper.tex
added poster

 
  \documentclass[final]{beamer} % beamer 3.10: do NOT use option hyperref={pdfpagelabels=false} !

  %\documentclass[final,hyperref={pdfpagelabels=false}]{beamer} % beamer 3.07: get rid of beamer warnings
  \mode<presentation> {  %% check http://www-i6.informatik.rwth-aachen.de/~dreuw/latexbeamerposter.php for examples
    \usetheme{Berlin}    %% you should define your own theme e.g. for big headlines using your own logos 
  }
\setbeamercolor{block body}{bg=white, fg=black}
\setbeamerfont{block title}{size=\Large}
  \usepackage[english]{babel}
  \usepackage[latin1]{inputenc}
  \usepackage{amsmath,amsthm, amssymb, latexsym}
  \usepackage{float}
\usepackage{color}
\usepackage{pgfplots}
\usepackage{anyfontsize}
  %\usepackage{times}\usefonttheme{professionalfonts}  % times is obsolete
  \usefonttheme[onlymath]{serif}
  \boldmath
  \usepackage[orientation=portrait,size=a0,scale=1.2, debug]{beamerposter}                       % e.g. for DIN-A0 poster
  %\usepackage[orientation=portrait,size=a1,scale=1.4,grid,debug]{beamerposter}                  % e.g. for DIN-A1 poster, with optional grid and debug output
  %\usepackage[size=custom,width=200,height=120,scale=2,debug]{beamerposter}                     % e.g. for custom size poster
  %\usepackage[orientation=portrait,size=a0,scale=1.0,printer=rwth-glossy-uv.df]{beamerposter}   % e.g. for DIN-A0 poster with rwth-glossy-uv printer check
  % ...
\definecolor{OliveGreen}{HTML}{eeeeee}

\begin{document}

  \begin{frame}
    \begin{center}

      \textcolor{black}{
      \textbf{\fontsize{110}{40}{\selectfont Panthera: A Study of Caching in 
      Distributed Computing}}}
    \end{center}

    \begin{columns}[t]
     \begin{column}{0.32\textwidth}
      \begin{block}{Introduction}
      Distributed computing has seen tremendous growth in the past few years.
In particular, Hadoop, an open source project that allows developers to
write distributed applications, has found use in academia and industry.
But, Hadoop does not effectively utilize RAM memory and local storage to
better performance. \textbf{This project develops and tests caching mechanisms
to reduce waiting time in Hadoop.}
      \end{block}

    \begin{block}{Problem}
	\begin{itemize}
	 \item Develop a caching layer for Hadoop that can cache both data and metadata.
	 \item The solution must not depend on a modified Hadoop version; it should be a "drop-in" addition.
	 \item The layer should reduce file access latency/waiting time significantly in comparison to a vanilla Hadoop installation.
	\end{itemize}

      \end{block}

      \begin{block}{Hypothesis}
       The cache layer designed (called \textbf{Panthera}) will have lower file
       access latency when compared to a standard Hadoop installation. Additionally, it
       will be possible to design \textbf{Panthera} such that the Hadoop source code
       remains unaffected.
      \end{block}

	  \begin{block}{Latency Types}
		%discuss the types of latency and make the comparison
		%between hard disk latency and regular lateny
		In the Hadoop File System Architecture, there are several types of latency involved:
		\begin{itemize}
		\item \textbf{Network latency: } Information shared between client and server (e.g. between a DataNode and a client) has an attached waiting time. With \textbf{Panthera}, this latency is eliminated since the data is cached on the client.
		\item \textbf{Hard drive latency: } Reading from the hard drive also has significant latency. Hadoop generally reads from the hard drive when accessing files.
		\item \textbf{Memory latency: } Reading from RAM memory has an extremely small
		waiting time associated with it. \textbf{Memory waiting time is less than 1/1000th hard drive waiting time. Panthera converts hard drive latency to memory latency, greatly reducing file access time.}
		\end{itemize}
	  \end{block}

		\begin{block}{Data vs Metadata}
			Data within the Hadoop file system consists of the actual contents of a file.
			Metadata is the information about files and directories, e.g. time a file was created, the list of files in a directory, etc. \textbf{Panthera performs both metadata
			and data caching; both have significantly different technical implications
			and challenges.}
		\end{block}
	  \begin{block}{Panthera Metadata Architecture}
	  \includegraphics[scale=1.2]{assets/panthera_metadata.pdf}
      \end{block}

      \begin{block}{Panthera Data Architecture}
      \includegraphics[scale=1.2]{assets/panthera_architecture.pdf}
      \end{block}

     \end{column}
     %##########################################################################################################

    \begin{column}{0.32\textwidth}

	\begin{block}{Panthera - Implementation}
	\begin{itemize}
		\item \textbf{Panthera is implemented on Google's Go programming language}. The language has excellent concurrency primitives and memory management, which are perfect for
		\textbf{Panthera} which must run reliably with hundreds of clients.
		\item \textbf{Panthera} implements a large portion of the Hadoop File System protocol, enabling it to "speak" with the DataNode and NameNode.
	\end{itemize}
	\end{block}

	\begin{block}{Testing Methodology - Single Node vs Multinode}
	All testing procedures were conducted on both \textbf{single node} and \textbf{multi-node} setups.\textbf{Single node} testing shows Panthera's performance without regard to network latency, whereas \textbf{multi node} testing incorporates network factors.
	\end{block}

	\begin{block}{Testing Methodology - Procedure}
	\begin{itemize}
		\item \textbf{Metadata}: Repeatedly request directory listing for a directory with 100 files in it.
		\item \textbf{Data}: Repeatedly request file contents for a 1MB large file.
	\end{itemize}
	\end{block}

	\begin{block}{Results}
		\includegraphics[scale=1.2]{assets/box-plot-data.pdf}
		\vspace{1em}
		\includegraphics[scale=1.2]{assets/box-plot-metadata.pdf}
		\vspace{1em}
		\includegraphics[scale=1.2]{assets/box-plot-meta-multinode.pdf}
	\end{block}

      \end{column}
      %##########################################################

      \begin{column}{0.32\textwidth}
    \begin{block}{Statistics}
    \begin{itemize}
    		\item \textbf{Data (single node, noncached)}
    		\begin{itemize}
    			\item Mean: 12548998 nanoseconds
    			%\item Median: 1292182 nanoseconds
    		\end{itemize}
    		\item \textbf{Data (single node, cached/Panthera)}
    		\begin{itemize}
    			\item Mean: 511185.3 nanoseconds
    			%\item Median: 149004 nanoseconds
    		\end{itemize}
    		\item \textbf{Noncached/Panthera}: \textbf{95.9\% improvement in latency with Panthera}, waiting time of noncached is 24.5 times higher than that of Panthera.

    		\item \textbf{Metadata (single node, noncached)}
    		\begin{itemize}
    			\item Mean: 8445265 nanoseconds
    			%\item Median: 1563000 nanoseconds
    		\end{itemize}

    		\item \textbf{Metadata (single node, cached)}
    		\begin{itemize}
    			\item Mean: 1055012 nanoseconds
    			%\item Median: 179500 nanoseconds
    		\end{itemize}

    		\item \textbf{Metadata (multinode, noncached)}
		\begin{itemize}
			\item Mean: 1638184 nanoseconds    		
    		\end{itemize}

    		\item \textbf{Metadata (multinode, Panthera cached)}
    		\begin{itemize}
			\item Mean: 176868.7 nanoseconds    		
    		\end{itemize}

    		\item \textbf{Metadata - noncached/Panthera: } \textbf{90.3\% improvement with the Panthera metadata cache}. Panthera (developed during this project) decreases waiting time by a factor of 8.

	\end{itemize}    	    
    \end{block}

    	\begin{block}{Conclusion}
	The results validate the hypothesis presented. Panthera was successfully developed as a "drop-in" solution and it was able to reduce file access latency in comparison to a standard Hadoop installation. In particular, \textbf{the Panthera data cache led to a 95.9\% or 9 fold decrease in waiting time} and the \textbf{the Panthera metadata cache also gave a 90.3\% or 8 fold decrease in latency}. Additionally, the \textbf{the cache developed during this project had a far lower standard deviation for access times and fewer outliers}. For large scale systems such as those in use at major companies and universities, such a drastic decrease in latency can have incredible effects on the efficiency of distributed computing systems.

	\end{block}

	\begin{block}{Applications}
	Hadoop has diverse applications and since Panthera was created as a "drop-in" solution, the \textbf{applications span an extremely wide range of disciplines}.
	\begin{itemize}
		\item Bioinformatics (e.g. CrossBow from Johns Hopkins University)
		\item Numerical/scientific computing 
		\item Image processing
		\item Artificial Intelligence/Machine Learning (e.g. Apache Mahout)
		\item Market Research (e.g. Target)
	\end{itemize}
	\end{block}

	\begin{block}{Future Work}
	\begin{itemize}
		\item \textbf{All code related to this project will be open sourced} to further development in distributed computing.
		\item Currently working on testing a \textbf{bioinformatics toolkit with Panthera}; initial results are very promising and show significant reduction in total running time.
		\item Smart scheduling using Panthera; arranges tasks so that the cache is used most efficiently.

	\end{itemize}
	\end{block}
      \end{column}
    \end{columns}
  \end{frame}

  \end{document}

### 2014-02-24 analysis/box-plot.r

@@ -6,7 +6,7 @@ noncached_data = unlist(na.omit(dat[2]))
 
 df <- data.frame(metacached = cached_data, non_metacached = noncached_data)
 pdf("boxplot.pdf")
-boxplot(df, ylim=c(0, 40000000), col=c('cornflowerblue', 'chocolate1'), xaxt="n")
+boxplot(df, ylim=c(0, 4000000), col=c('cornflowerblue', 'chocolate1'), xaxt="n")
 title("Metacached (Panthera) vs Non-metacached")
 axis(1, at=1:2, labels=c("metacached (panthera)", "non-metacached"))
 title(ylab = "Access Time (nanoseconds, lower is better)")

### 2014-02-24 src/build/numpy/build/src.linux-x86_64-2.7/numpy/__config__.py

@@ -1,0 +1,33 @@
+# This file is generated by /home/dhaivat/dev/panthera/src/build/numpy/-c
+# It contains system_info results at the time of building this package.
+__all__ = ["get_info","show"]
+
+blas_info={'libraries': ['blas'], 'library_dirs': ['/usr/lib'], 'language': 'f77'}
+lapack_info={'libraries': ['lapack'], 'library_dirs': ['/usr/lib'], 'language': 'f77'}
+atlas_threads_info={}
+blas_opt_info={'libraries': ['blas'], 'library_dirs': ['/usr/lib'], 'define_macros': [('NO_ATLAS_INFO', 1)], 'language': 'f77'}
+openblas_info={}
+atlas_blas_threads_info={}
+lapack_opt_info={'libraries': ['lapack', 'blas'], 'library_dirs': ['/usr/lib'], 'define_macros': [('NO_ATLAS_INFO', 1)], 'language': 'f77'}
+atlas_info={}
+lapack_mkl_info={}
+blas_mkl_info={}
+atlas_blas_info={}
+mkl_info={}
+
+def get_info(name):
+    g = globals()
+    return g.get(name, g.get(name + "_info", {}))
+
+def show():
+    for name,info_dict in globals().items():
+        if name[0] == "_" or type(info_dict) is not type({}): continue
+        print(name + ":")
+        if not info_dict:
+            print("  NOT AVAILABLE")
+        for k,v in info_dict.items():
+            v = str(v)
+            if k == "sources" and len(v) > 200:
+                v = v[:60] + " ...\n... " + v[-60:]
+            print("    %s = %s" % (k,v))
+    

### 2014-02-24 src/build/numpy/build/src.linux-x86_64-2.7/numpy/core/include/numpy/__multiarray_api.c
removed builds

@@ -1,307 +1,0 @@
-
-/* These pointers will be stored in the C-object for use in other
-    extension modules
-*/
-
-void *PyArray_API[] = {
-        (void *) PyArray_GetNDArrayCVersion,
-        (void *) &PyBigArray_Type,
-        (void *) &PyArray_Type,
-        (void *) &PyArrayDescr_Type,
-        (void *) &PyArrayFlags_Type,
-        (void *) &PyArrayIter_Type,
-        (void *) &PyArrayMultiIter_Type,
-        (int *) &NPY_NUMUSERTYPES,
-        (void *) &PyBoolArrType_Type,
-        (void *) &_PyArrayScalar_BoolValues,
-        (void *) &PyGenericArrType_Type,
-        (void *) &PyNumberArrType_Type,
-        (void *) &PyIntegerArrType_Type,
-        (void *) &PySignedIntegerArrType_Type,
-        (void *) &PyUnsignedIntegerArrType_Type,
-        (void *) &PyInexactArrType_Type,
-        (void *) &PyFloatingArrType_Type,
-        (void *) &PyComplexFloatingArrType_Type,
-        (void *) &PyFlexibleArrType_Type,
-        (void *) &PyCharacterArrType_Type,
-        (void *) &PyByteArrType_Type,
-        (void *) &PyShortArrType_Type,
-        (void *) &PyIntArrType_Type,
-        (void *) &PyLongArrType_Type,
-        (void *) &PyLongLongArrType_Type,
-        (void *) &PyUByteArrType_Type,
-        (void *) &PyUShortArrType_Type,
-        (void *) &PyUIntArrType_Type,
-        (void *) &PyULongArrType_Type,
-        (void *) &PyULongLongArrType_Type,
-        (void *) &PyFloatArrType_Type,
-        (void *) &PyDoubleArrType_Type,
-        (void *) &PyLongDoubleArrType_Type,
-        (void *) &PyCFloatArrType_Type,
-        (void *) &PyCDoubleArrType_Type,
-        (void *) &PyCLongDoubleArrType_Type,
-        (void *) &PyObjectArrType_Type,
-        (void *) &PyStringArrType_Type,
-        (void *) &PyUnicodeArrType_Type,
-        (void *) &PyVoidArrType_Type,
-        (void *) PyArray_SetNumericOps,
-        (void *) PyArray_GetNumericOps,
-        (void *) PyArray_INCREF,
-        (void *) PyArray_XDECREF,
[…]

### 2014-02-24 src/build/numpy/build/src.linux-x86_64-2.7/numpy/core/include/numpy/__umath_generated.c

@@ -1,1164 +1,0 @@
-
-
-/** Warning this file is autogenerated!!!
-
-    Please make changes to the code generator program (numpy/core/code_generators/generate_umath.py)
-**/
-
-static PyUFuncGenericFunction _arg_functions[] = { CFLOAT__arg, CDOUBLE__arg, CLONGDOUBLE__arg };
-static void * _arg_data[] = { (void *)NULL, (void *)NULL, (void *)NULL };
-static char _arg_signatures[] = { NPY_CFLOAT, NPY_FLOAT, NPY_CDOUBLE, NPY_DOUBLE, NPY_CLONGDOUBLE, NPY_LONGDOUBLE };
-static PyUFuncGenericFunction _ones_like_functions[] = { BOOL__ones_like, BYTE__ones_like, UBYTE__ones_like, SHORT__ones_like, USHORT__ones_like, INT__ones_like, UINT__ones_like, LONG__ones_like, ULONG__ones_like, LONGLONG__ones_like, ULONGLONG__ones_like, HALF__ones_like, FLOAT__ones_like, DOUBLE__ones_like, LONGDOUBLE__ones_like, CFLOAT__ones_like, CDOUBLE__ones_like, CLONGDOUBLE__ones_like, DATETIME__ones_like, TIMEDELTA__ones_like, NULL };
-static void * _ones_like_data[] = { (void *)NULL, (void *)NULL, (void *)NULL, (void *)NULL, (void *)NULL, (void *)NULL, (void *)NULL, (void *)NULL, (void *)NULL, (void *)NULL, (void *)NULL, (void *)NULL, (void *)NULL, (void *)NULL, (void *)NULL, (void *)NULL, (void *)NULL, (void *)NULL, (void *)NULL, (void *)NULL, (void *)NULL };
[…]

### 2014-02-24 src/build/numpy/build/src.linux-x86_64-2.7/numpy/core/include/numpy/config.h

@@ -1,151 +1,0 @@
-#define HAVE_ENDIAN_H 1
-#define SIZEOF_PY_INTPTR_T 8
-#define SIZEOF_PY_LONG_LONG 8
-#define MATHLIB m
-#define HAVE_SIN 1
-#define HAVE_COS 1
-#define HAVE_TAN 1
-#define HAVE_SINH 1
-#define HAVE_COSH 1
-#define HAVE_TANH 1
-#define HAVE_FABS 1
-#define HAVE_FLOOR 1
-#define HAVE_CEIL 1
-#define HAVE_SQRT 1
-#define HAVE_LOG10 1
-#define HAVE_LOG 1
-#define HAVE_EXP 1
-#define HAVE_ASIN 1
-#define HAVE_ACOS 1
-#define HAVE_ATAN 1
-#define HAVE_FMOD 1
-#define HAVE_MODF 1
-#define HAVE_FREXP 1
-#define HAVE_LDEXP 1
-#define HAVE_RINT 1
-#define HAVE_TRUNC 1
-#define HAVE_EXP2 1
-#define HAVE_LOG2 1
-#define HAVE_ATAN2 1
-#define HAVE_POW 1
-#define HAVE_NEXTAFTER 1
-#define HAVE_XMMINTRIN_H 1
-#define HAVE_EMMINTRIN_H 1
-#define HAVE___BUILTIN_ISNAN 1
-#define HAVE___BUILTIN_ISINF 1
-#define HAVE___BUILTIN_ISFINITE 1
-#define HAVE___BUILTIN_BSWAP32 1
-#define HAVE___BUILTIN_BSWAP64 1
-#define HAVE___BUILTIN_EXPECT 1
-#define HAVE__MM_LOAD_PS 1
-#define HAVE__MM_LOAD_PD 1
-#define HAVE_ATTRIBUTE_OPTIMIZE_UNROLL_LOOPS 1
-#define HAVE_SINF 1
-#define HAVE_COSF 1
-#define HAVE_TANF 1
-#define HAVE_SINHF 1
-#define HAVE_COSHF 1
-#define HAVE_TANHF 1
-#define HAVE_FABSF 1
-#define HAVE_FLOORF 1
-#define HAVE_CEILF 1
-#define HAVE_RINTF 1
-#define HAVE_TRUNCF 1
-#define HAVE_SQRTF 1
-#define HAVE_LOG10F 1
-#define HAVE_LOGF 1
-#define HAVE_LOG1PF 1
[…]

### 2014-02-24 src/build/numpy/build/src.linux-x86_64-2.7/numpy/core/src/multiarray/multiarray_tests.c

@@ -1,808 +1,0 @@
-#line 1 "numpy/core/src/multiarray/multiarray_tests.c.src"
-
-/*
- *****************************************************************************
- **       This file was autogenerated from a template  DO NOT EDIT!!!!      **
- **       Changes should be made to the original source (.src) file         **
- *****************************************************************************
- */
-
-#line 1
-#define NPY_NO_DEPRECATED_API NPY_API_VERSION
-#include <Python.h>
-#include "numpy/arrayobject.h"
-
-#include "npy_pycompat.h"
-
-/*
- * TODO:
- *  - Handle mode
- */
-
-#line 17
-static int copy_double(PyArrayIterObject *itx, PyArrayNeighborhoodIterObject *niterx,
-        npy_intp *bounds,
-        PyObject **out)
-{
-    npy_intp i, j;
-    npy_double *ptr;
-    npy_intp odims[NPY_MAXDIMS];
-    PyArrayObject *aout;
-
-    /*
-     * For each point in itx, copy the current neighborhood into an array which
-     * is appended at the output list
-     */
-    for (i = 0; i < itx->size; ++i) {
-        PyArrayNeighborhoodIter_Reset(niterx);
-
-        for (j = 0; j < PyArray_NDIM(itx->ao); ++j) {
-            odims[j] = bounds[2 * j + 1] - bounds[2 * j] + 1;
-        }
-        aout = (PyArrayObject*)PyArray_SimpleNew(
-                                PyArray_NDIM(itx->ao), odims, NPY_DOUBLE);
-        if (aout == NULL) {
-            return -1;
-        }
-
-        ptr = (npy_double*)PyArray_DATA(aout);
-
-        for (j = 0; j < niterx->size; ++j) {
-            *ptr = *((npy_double*)niterx->dataptr);
-            PyArrayNeighborhoodIter_Next(niterx);
-            ptr += 1;
-        }
-
-        PyList_Append(*out, (PyObject*)aout);
-        Py_DECREF(aout);
-        PyArray_ITER_NEXT(itx);
-    }
-
-    return 0;
-}
-
-#line 17
-static int copy_int(PyArrayIterObject *itx, PyArrayNeighborhoodIterObject *niterx,
-        npy_intp *bounds,
-        PyObject **out)
-{
-    npy_intp i, j;
[…]

### 2014-02-24 src/build/numpy/build/src.linux-x86_64-2.7/numpy/core/src/umath/operand_flag_tests.c

@@ -1,116 +1,0 @@
-#line 1 "numpy/core/src/umath/operand_flag_tests.c.src"
-
-/*
- *****************************************************************************
- **       This file was autogenerated from a template  DO NOT EDIT!!!!      **
- **       Changes should be made to the original source (.src) file         **
- *****************************************************************************
- */
-
-#line 1
-#define NPY_NO_DEPRECATED_API NPY_API_VERSION
-
-#include <math.h>
-#include <Python.h>
-#include <structmember.h>
-#include <numpy/arrayobject.h>
-#include <numpy/ufuncobject.h>
-#include "numpy/npy_3kcompat.h"
-
-
-static PyMethodDef TestMethods[] = {
-        {NULL, NULL, 0, NULL}
-};
-
-
-static void
-inplace_add(char **args, npy_intp *dimensions, npy_intp *steps, void *data)
-{
-    npy_intp i;
-    npy_intp n = dimensions[0];
-    char *in1 = args[0];
-    char *in2 = args[1];
-    npy_intp in1_step = steps[0];
-    npy_intp in2_step = steps[1];
-
-    for (i = 0; i < n; i++) {
-        (*(long *)in1) = *(long*)in1 + *(long*)in2;
-        in1 += in1_step;
-        in2 += in2_step;
-    }
-}
-
-
-/*This a pointer to the above function*/
-PyUFuncGenericFunction funcs[1] = {&inplace_add};
-
-/* These are the input and return dtypes of logit.*/
-static char types[2] = {NPY_LONG, NPY_LONG};
-
-static void *data[1] = {NULL};
-
-#if defined(NPY_PY3K)
-static struct PyModuleDef moduledef = {
-    PyModuleDef_HEAD_INIT,
-    "operand_flag_tests",
-    NULL,
-    -1,
-    TestMethods,
-    NULL,
-    NULL,
-    NULL,
-    NULL
-};
-
-#define RETVAL m
-PyMODINIT_FUNC PyInit_operand_flag_tests(void)
-{
-#else
-#define RETVAL
-PyMODINIT_FUNC initoperand_flag_tests(void)
-{
-#endif
-    PyObject *m = NULL;
-    PyObject *ufunc;
-
-#if defined(NPY_PY3K)
-    m = PyModule_Create(&moduledef);
-#else
-    m = Py_InitModule("operand_flag_tests", TestMethods);
-#endif
-    if (m == NULL) {
-        goto fail;
-    }
-
[…]

### 2014-02-24 src/build/numpy/build/src.linux-x86_64-2.7/numpy/core/src/umath/test_rational.c

@@ -1,1412 +1,0 @@
-#line 1 "numpy/core/src/umath/test_rational.c.src"
-
-/*
- *****************************************************************************
- **       This file was autogenerated from a template  DO NOT EDIT!!!!      **
- **       Changes should be made to the original source (.src) file         **
- *****************************************************************************
- */
-
-#line 1
-/* Fixed size rational numbers exposed to Python */
-
-#define NPY_NO_DEPRECATED_API NPY_API_VERSION
-
-#include <math.h>
-#include <Python.h>
-#include <structmember.h>
-#include <numpy/arrayobject.h>
-#include <numpy/ufuncobject.h>
-#include <numpy/npy_3kcompat.h>
-
-/* Relevant arithmetic exceptions */
-
-/* Uncomment the following line to work around a bug in numpy */
-/* #define ACQUIRE_GIL */
-
-static void
-set_overflow(void) {
-#ifdef ACQUIRE_GIL
-    /* Need to grab the GIL to dodge a bug in numpy */
-    PyGILState_STATE state = PyGILState_Ensure();
-#endif
-    if (!PyErr_Occurred()) {
-        PyErr_SetString(PyExc_OverflowError,
-                "overflow in rational arithmetic");
-    }
-#ifdef ACQUIRE_GIL
-    PyGILState_Release(state);
-#endif
-}
-
-static void
-set_zero_divide(void) {
-#ifdef ACQUIRE_GIL
-    /* Need to grab the GIL to dodge a bug in numpy */
-    PyGILState_STATE state = PyGILState_Ensure();
-#endif
-    if (!PyErr_Occurred()) {
-        PyErr_SetString(PyExc_ZeroDivisionError,
-                "zero divide in rational arithmetic");
-    }
-#ifdef ACQUIRE_GIL
-    PyGILState_Release(state);
-#endif
-}
-
-/* Integer arithmetic utilities */
-
-static NPY_INLINE npy_int32
-safe_neg(npy_int32 x) {
-    if (x==(npy_int32)1<<31) {
-        set_overflow();
-    }
-    return -x;
-}
-
-static NPY_INLINE npy_int32
-safe_abs32(npy_int32 x) {
-    npy_int32 nx;
-    if (x>=0) {
-        return x;
-    }
-    nx = -x;
-    if (nx<0) {
-        set_overflow();
-    }
-    return nx;
-}
-
-static NPY_INLINE npy_int64
-safe_abs64(npy_int64 x) {
[…]

### 2014-02-24 src/build/numpy/build/src.linux-x86_64-2.7/numpy/core/src/umath/umath_tests.c

@@ -1,491 +1,0 @@
-#line 1 "numpy/core/src/umath/umath_tests.c.src"
-
-/*
- *****************************************************************************
- **       This file was autogenerated from a template  DO NOT EDIT!!!!      **
- **       Changes should be made to the original source (.src) file         **
- *****************************************************************************
- */
-
-#line 1
-/* -*- c -*- */
-
-/*
- *****************************************************************************
- **                            INCLUDES                                     **
- *****************************************************************************
- */
-#define NPY_NO_DEPRECATED_API NPY_API_VERSION
-
-#include "Python.h"
-#include "numpy/arrayobject.h"
-#include "numpy/ufuncobject.h"
-
-#include "npy_pycompat.h"
-
-#include "npy_config.h"
-
-/*
- *****************************************************************************
- **                            BASICS                                       **
- *****************************************************************************
- */
-
-#define INIT_OUTER_LOOP_1       \
-    npy_intp dN = *dimensions++;\
-    npy_intp N_;                \
-    npy_intp s0 = *steps++;
-
-#define INIT_OUTER_LOOP_2       \
-    INIT_OUTER_LOOP_1           \
-    npy_intp s1 = *steps++;
-
-#define INIT_OUTER_LOOP_3       \
-    INIT_OUTER_LOOP_2           \
-    npy_intp s2 = *steps++;
-
-#define INIT_OUTER_LOOP_4       \
-    INIT_OUTER_LOOP_3           \
-    npy_intp s3 = *steps++;
-
-#define BEGIN_OUTER_LOOP_3      \
-    for (N_ = 0; N_ < dN; N_++, args[0] += s0, args[1] += s1, args[2] += s2) {
-
-#define BEGIN_OUTER_LOOP_4      \
-    for (N_ = 0; N_ < dN; N_++, args[0] += s0, args[1] += s1, args[2] += s2, args[3] += s3) {
-
-#define END_OUTER_LOOP  }
-
-
-/*
- *****************************************************************************
[…]

### 2014-02-24 src/build/numpy/build/src.linux-x86_64-2.7/numpy/distutils/__config__.py

@@ -1,33 +1,0 @@
-# This file is generated by /home/dhaivat/dev/panthera/src/build/numpy/-c
-# It contains system_info results at the time of building this package.
-__all__ = ["get_info","show"]
-
-blas_info={'libraries': ['blas'], 'library_dirs': ['/usr/lib'], 'language': 'f77'}
-lapack_info={'libraries': ['lapack'], 'library_dirs': ['/usr/lib'], 'language': 'f77'}
-atlas_threads_info={}
-blas_opt_info={'libraries': ['blas'], 'library_dirs': ['/usr/lib'], 'define_macros': [('NO_ATLAS_INFO', 1)], 'language': 'f77'}
-openblas_info={}
-atlas_blas_threads_info={}
-lapack_opt_info={'libraries': ['lapack', 'blas'], 'library_dirs': ['/usr/lib'], 'define_macros': [('NO_ATLAS_INFO', 1)], 'language': 'f77'}
-atlas_info={}
-lapack_mkl_info={}
-blas_mkl_info={}
-atlas_blas_info={}
-mkl_info={}
-
-def get_info(name):
-    g = globals()
-    return g.get(name, g.get(name + "_info", {}))
-
-def show():
-    for name,info_dict in globals().items():
-        if name[0] == "_" or type(info_dict) is not type({}): continue
-        print(name + ":")
-        if not info_dict:
-            print("  NOT AVAILABLE")
-        for k,v in info_dict.items():
-            v = str(v)
-            if k == "sources" and len(v) > 200:
-                v = v[:60] + " ...\n... " + v[-60:]
-            print("    %s = %s" % (k,v))
-    

### 2014-02-24 src/build/numpy/build/src.linux-x86_64-2.7/numpy/linalg/umath_linalg.c

@@ -1,6648 +1,0 @@
-#line 1 "numpy/linalg/umath_linalg.c.src"
-
-/*
- *****************************************************************************
- **       This file was autogenerated from a template  DO NOT EDIT!!!!      **
- **       Changes should be made to the original source (.src) file         **
- *****************************************************************************
- */
-
-#line 1
-/* -*- c -*- */
-
-/*
- *****************************************************************************
- **                            INCLUDES                                     **
- *****************************************************************************
- */
-#define NPY_NO_DEPRECATED_API NPY_API_VERSION
-
-#include "Python.h"
-#include "numpy/arrayobject.h"
-#include "numpy/ufuncobject.h"
-
-#include "npy_pycompat.h"
-
-#include "npy_config.h"
-
-#include <stddef.h>
-#include <stdio.h>
-#include <assert.h>
-#include <math.h>
-
-
-static const char* umath_linalg_version_string = "0.1.4";
-
-/*
- ****************************************************************************
- *                        Debugging support                                 *
- ****************************************************************************
- */
-#define TRACE_TXT(...) do { fprintf (stderr, __VA_ARGS__); } while (0)
-#define STACK_TRACE do {} while (0)
-#define TRACE\
-    do {                                        \
-        fprintf (stderr,                        \
-                 "%s:%d:%s\n",                  \
-                 __FILE__,                      \
-                 __LINE__,                      \
-                 __FUNCTION__);                 \
-        STACK_TRACE;                            \
-    } while (0)
-
-#if 0
-#include <execinfo.h>
-void
-dbg_stack_trace()
-{
-    void *trace[32];
-    size_t size;
-
-    size = backtrace(trace, sizeof(trace)/sizeof(trace[0]));
-    backtrace_symbols_fd(trace, size, 1);
-}
-
-#undef STACK_TRACE
-#define STACK_TRACE do { dbg_stack_trace(); } while (0)
-#endif
-
-/*
- *****************************************************************************
- *                    BLAS/LAPACK calling macros                             *
[…]

### 2014-02-24 poster/paper.tex
first version of poster

@@
 In particular, Hadoop, an open source project that allows developers to
-write distributed applications, has found use in academia and industry.
-But, Hadoop does not effectively utilize RAM memory and local storage to
-better performance. \textbf{This project develops and tests caching mechanisms
-to reduce waiting time in Hadoop.}
+write distributed applications, has found extensive use in academia and industry.
+For example, Yahoo, Facebook, Stanford Computer Science, etc.
+are all heavily involved with Hadoop development.
+But, Hadoop does not effectively utilize RAM memory and local storage to better performance.
+\textbf{This project develops and tests caching mechanisms, called \textbf{Panthera}, to reduce waiting time or latency in Hadoop, } thereby greatly increasing its efficiency and applicability in fields ranging from medical diagnosis to artificial intelligence.
       \end{block}
@@
 	\begin{itemize}
-	 \item Develop a caching layer for Hadoop that can cache both data and metadata.
+	 \item Develop a caching layer for the Hadoop File System that can cache both data and metadata.
 	 \item The solution must not depend on a modified Hadoop version; it should be a "drop-in" addition.
@@
 
-	  \begin{block}{Latency Types}
-		%discuss the types of latency and make the comparison
-		%between hard disk latency and regular lateny
-		In the Hadoop File System Architecture, there are several types of latency involved:
-		\begin{itemize}
-		\item \textbf{Network latency: } Information shared between client and server (e.g.
-between a DataNode and a client) has an attached waiting time.
-With \textbf{Panthera}, this latency is eliminated since the data is cached on the client.
-		\item \textbf{Hard drive latency: } Reading from the hard drive also has significant latency. Hadoop generally reads from the hard drive when accessing files.
-		\item \textbf{Memory latency: } Reading from RAM memory has an extremely small
-		waiting time associated with it.
-\textbf{Memory waiting time is less than 1/1000th hard drive waiting time.
-Panthera converts hard drive latency to memory latency, greatly reducing file access time.}
-		\end{itemize}
-	  \end{block}
-
 		\begin{block}{Data vs Metadata}
@@
 		\end{block}
+
+		\begin{block}{Current Hadoop File System Architecture}
+			\includegraphics[scale=1]{assets/hadoop_architecture.pdf}
+		\end{block}
+
+	  \begin{block}{Panthera Architecture}
+		\includegraphics[scale=1]{assets/panthera_architecture_complete.pdf}	  
+	  \end{block}
+
 	  \begin{block}{Panthera Metadata Architecture}
-	  \includegraphics[scale=1.2]{assets/panthera_metadata.pdf}
+	  \includegraphics[scale=1]{assets/panthera_metadata.pdf}
       \end{block}
@@
       \begin{block}{Panthera Data Architecture}
-      \includegraphics[scale=1.2]{assets/panthera_architecture.pdf}
+      \includegraphics[scale=1]{assets/panthera_architecture.pdf}
       \end{block}
@@
 		\item \textbf{Panthera is implemented on Google's Go programming language}. The language has excellent concurrency primitives and memory management, which are perfect for
-		\textbf{Panthera} which must run reliably with hundreds of clients.
+		for software ,such as \textbf{Panthera}, that must run reliably with hundreds of clients.
 		\item \textbf{Panthera} implements a large portion of the Hadoop File System protocol, enabling it to "speak" with the DataNode and NameNode.
@@
 	\begin{block}{Testing Methodology - Single Node vs Multinode}
-	All testing procedures were conducted on both \textbf{single node} and \textbf{multi-node} setups.\textbf{Single node} testing shows Panthera's performance without regard to network latency, whereas \textbf{multi node} testing incorporates network factors.
+	Testing procedures were conducted on both \textbf{single node} and \textbf{multi-node} setups.\textbf{Single node} testing shows Panthera's performance without regard to network latency, whereas \textbf{multinode} testing incorporates network factors.
 	\end{block}
@@
       \begin{column}{0.32\textwidth}
+
+	  \begin{block}{Latency Types}
+		%discuss the types of latency and make the comparison
+		%between hard disk latency and regular lateny
+		In the Hadoop File System Architecture, there are several types of latency involved:
+		\begin{itemize}
+		\item \textbf{Network latency: } Information shared between client and server (e.g.
+between a DataNode and a client) has an attached waiting time.
+With \textbf{Panthera}, this latency is eliminated since the data is cached on the client.
+		\item \textbf{Hard drive latency: } Reading from the hard drive has very significant latency. Hadoop generally reads from the hard drive when accessing files.
+		\item \textbf{Memory latency: } Reading from RAM memory has an extremely small
+		waiting time associated with it.
+\textbf{Memory waiting time is less than 1/1000th hard drive waiting time.
+Panthera converts hard drive latency to memory latency, thereby greatly reducing file access time.}
+		\end{itemize}
+	  \end{block}
+
     \begin{block}{Statistics}

### 2014-02-28 poster/paper.tex
poster fixes

@@
   \boldmath
-  \usepackage[orientation=portrait,size=a0,scale=1.2, debug]{beamerposter}                       % e.g. for DIN-A0 poster
+  \usepackage[orientation=portrait,size=a0,scale=1.15, debug]{beamerposter}                       % e.g. for DIN-A0 poster
   %\usepackage[orientation=portrait,size=a1,scale=1.4,grid,debug]{beamerposter}                  % e.g. for DIN-A1 poster, with optional grid and debug output
@@
 		\begin{block}{Current Hadoop File System Architecture}
-			\includegraphics[scale=1]{assets/hadoop_architecture.pdf}
+			\centerline{\includegraphics[scale=1.5]{assets/vanilla_hadoop.pdf}}
 		\end{block}
@@
 	  \begin{block}{Panthera Architecture}
-		\includegraphics[scale=1]{assets/panthera_architecture_complete.pdf}	  
+		\centerline{\includegraphics[scale=1.5]{assets/panthera_architecture_complete_larger.pdf}}	  
 	  \end{block}
@@
 		\begin{block}{Panthera Metadata Architecture}
-	  \includegraphics[scale=1]{assets/panthera_metadata.pdf}
-      \end{block}
-
-      \begin{block}{Panthera Data Architecture}
-      \includegraphics[scale=1]{assets/panthera_architecture.pdf}
+	 		\centerline{\includegraphics[scale=1.5]{assets/panthera_metadata_large.pdf}}
       	\end{block}
@@
     \begin{column}{0.32\textwidth}
+
+      \begin{block}{Panthera Data Architecture}
+      \includegraphics[scale=1.2]{assets/panthera_data_large.pdf}
+     \end{block}
 
@@
 	\begin{block}{Results}
-		\includegraphics[scale=1.2]{assets/box-plot-data.pdf}
+		\includegraphics[scale=1.10]{assets/box-plot-data-larger.pdf}
 		\vspace{1em}
-		\includegraphics[scale=1.2]{assets/box-plot-metadata.pdf}
+		\includegraphics[scale=1.1]{assets/box-plot-metadata-larger.pdf}
 		\vspace{1em}
-		\includegraphics[scale=1.2]{assets/box-plot-meta-multinode.pdf}
+		\includegraphics[scale=1.1]{assets/box-plot-meta-multinode-larger.pdf}
 	\end{block}

### 2014-02-28 analysis/box-plot.r

@@ -5,9 +5,13 @@ cached_data = unlist(na.omit(dat[1]))
 noncached_data = unlist(na.omit(dat[2]))
 
 df <- data.frame(metacached = cached_data, non_metacached = noncached_data)
-pdf("boxplot.pdf")
-boxplot(df, ylim=c(0, 4000000), col=c('cornflowerblue', 'chocolate1'), xaxt="n")
-title("Metacached (Panthera) vs Non-metacached")
-axis(1, at=1:2, labels=c("metacached (panthera)", "non-metacached"))
+pdf("boxplot.pdf", width=9.5, height=7)
+par(mar=c(5, 4.5, 4, 2))
+boxplot(df, ylim=c(0, 4000000), col=c('chocolate1', 'cornflowerblue'), xaxt="n", cex.main=1.5, whisklwd=4, staplelwd=4, outlwd=2, outcex=1.2)
+par(cex.main=1.9)
+title("Metacached (Panthera) vs Non-metacached (Status Quo)\nMultinode")
+par(cex.axis=1.8)
+axis(1, at=1:2, labels=c("metacached (Panthera)", "non-metacached"))
+par(cex.lab=1.8)
 title(ylab = "Access Time (nanoseconds, lower is better)")
 dev.off()

### 2014-02-28 analysis/histogram.r
added distribution for cached latency

@@ -4,6 +4,6 @@ cached_latency = unlist(na.omit(dat[1]))
 no_cached_latency = unlist(na.omit(dat[2]))
 
 pdf("histogram.pdf")
-hist(no_cached_latency, breaks=1000, col='chocolate1', xlim=c(0, 16000000), ylim=c(0, 250))
-hist(cached_latency, breaks=100, col='cornflowerblue', add=T)
+hist(no_cached_latency, breaks=1000, col='chocolate1', xlim=c(0, 1600000), ylim=c(0, 25))
+hist(cached_latency, breaks=10, col='cornflowerblue', add=T)
 dev.off()

### 2014-03-01 paper/paper.tex
paper added

@@
 \title{Panthera: A Study of Caching in Distributed Computing}
-\author{Dhaivat Pandya}
+\author{Dhaivat Pandya \\ Appleton North High School}
 \date{}
-
 %
@@
 \abstract{
-Describe your paper in 100-200 words, give or take.
-The command-line \texttt{wc} utility is really useful here!
-This particular sample paper is meant to demonstrate a variety of \LaTeX\ directives for producing a well-structured, consistently-formatted scholarly document.
-The actual content and outline may vary according to the needs of your specific research topic.
+The adoption of distributed computing systems has grown massively in the past few years.
+In particular, Apache Hadoop, which allows developers to create applications that run on a cluster of computers, is currently used throughout academia and industry.
+However, the Hadoop File System fails to effectively utilize memory and local storage in order to reduce waiting time (i.e.
+latency).
+In this project, a system, named Panthera, was developed, which creates a caching system for Hadoop metadata and data.
+Panthera caches both information about files and the file contents, thereby reducing waiting time in downloading and accessing the files and related information.
+It runs with an unmodified version of Hadoop, meaning that it can integrate easily into existing architecture.
+Panthera was tested using both single node and multi-
+node setups.
+The results showed an 90.8\% decrease in waiting time with the Panthera metadata cache, and a 95.7\% decrease with the Panthera data cache.
+Such drastic decreases in latency can greatly increase the efficiency of existing Hadoop applications and also allow the creation of algorithms that were previously not feasible.
+Panthera has numerous applications in fields ranging from market research to bioinformatics.
+All code associated with the project will be open sourced to further development in the area.
 }
@@
 \section{Introduction}
-The Hadoop distributed system \cite{hadoop} is an open source version of the revolutionary MapReduce system developed at Google.
+Distributed computing systems are currently employed across academia and industry.
+Applications vary from market research for department stores to bioinformatics.
+In particular, the Hadoop distributed system \cite{hadoop} has seen tremendous growth.
+Hadoop is an open source version of the revolutionary MapReduce system developed at Google.
 With it, developers can take easily take advantage of large, multi-node clusters to solve computational problems.
+
+Current Hadoop deployments exist in the thousands of nodes at companies such as Cloudera, Facebook, Yahoo, etc.
+However, especially in larger networks, file access latency can be a significant.
+The Hadoop File System currently does not use local (i.e.
+on the client) random access memory (RAM) to improve latency.
 
@@
 As the RAMCloud project has outlined \cite{ramcloud}, low latency can greatly extend the applications of distributed computing systems.
+For example, tree traversal algorithms are largely impractical in Hadoop, however, with lower latency, such algorithms may become useful and applicable.
 
 In this paper, we discuss \textit{Panthera}, a cache layer for Hadoop.
+In Section 2, we discuss the specific challenges associated with such a project.
+Section 3 details the architecture differences between a standard Hadoop installation and one with \textit{Panthera}.
 
@@
 
-There are several systems built on Hadoop that are in widespread use \cite{hbase, cloudbatch, pig}.
-Thus, for \textit{Panthera} to be practical, it must operate independently of the existing Hadoop codebase.
+There are several systems built on Hadoop that are in widespread use \cite{hbase, cloudbatch, pig}. The Hadoop project is also supported by large corporations with a 
+consistent release cycle.
+Thus, for \textit{Panthera} to be practical, it must operate independently of the existing Hadoop codebase, i.e.
+as a layer, rather than a modification.
+This implies that \textit{Panthera} is implemented completely independently of the Hadoop codebase.
+
 Additionally, for non-cache related requests, \textit{Panthera} must add insignificant latency.
-Finally, data and metadata request latency should be significantly with \textit{Panthera} in comparison to a vanilla Hadoop installation.
+Finally, data and metadata request latency should be significantly with \textit{Panthera} in comparison to a standard Hadoop installation.
 
 \section{Architecture}
-\begin{figure}
-\begin{center}
-	\begin{tikzpicture}
-		\begin{tikzpicture}[scale=1.5, transform shape]
-		\draw[style=solid] (0,0) circle (1) node {DataNode 1};
-		\draw[style=dashed] (0, 0) circle(1.3) node {DataNode 1};
-		\draw[style=solid] (3.0, 2.5) node {NameNode}  circle (1);
-		\draw[style=solid] (3.0, 0) node{DataNode 2} circle (1);
-		\draw[style=solid] (6.0, 0) node{DataNode 3} circle (1);
-		\draw[loosely dashed, line width=2pt]  (2.5,0.3)-- (2.5, 2.3);
-		\draw[loosely dashed, line width=2pt]  (0, 0.3) -- (2.5, 2.3);
-		\draw[loosely dashed, line width=2pt]  (5.0, 0.3) -- (2.5, 2.3);
-	\end{tikzpicture}
-	\end{tikzpicture}
-\end{center}
+\begin{figure}[!h]
+  \caption{Standard Hadoop Architecture (left), Panthera-Hadoop Architecture (right)}
+  \centering
+	\includegraphics[scale=0.4]{assets/hadoop_architecture.pdf}
+	\includegraphics[scale=0.4]{assets/panthera_architecture.pdf}
 \end{figure}
 
-\textit{Panthera} runs as a standalone program on every slave node (usually also a DataNode) of a given Hadoop Distributed File System cluster.
+Figure 1 outlines the architecture differences between a standard Hadoop File System architecture and an installation with Panthera.
+
+Within the standard conditions, the \textit{NameNode} serves as a metadata server, handling all requests from clients relating to metadata (e.g.
+directory listing, file sizes, block locations, etc.).
+\textit{DataNode(s)} hold data and serve it at request from the client.
+A more complete description of the Hadoop File System may be found in \cite{hadoop}.
+
+Panthera operates by intercepting requests and responses at the NameNode and the client.
+Both nodes have a running instance of \textit{Panthera}.
+Every request issued from the client is first reviewed by Panthera, and if possible, answered immediately from the cache.
+The \textit{Panthera} instance on the NameNode constantly monitors the filesystem for changes that need to be reflected in the cache (e.g.
+if a file is deleted, the cache on the client should have a copy of it).
+
+Panthera, since it caches both data and metadata, can be considered two separate subsystems, which are now discussed.
+
+\section{Metadata caching}
+
+\begin{figure}[!h]
+	\caption{Panthera Metadata System}
+	\centering
+		\includegraphics[scale=0.4]{assets/panthera_meta_architecture.pdf}
+\end{figure}
+
+The Panthera metadata system runs on the NameNode and communicates with the Panthera instance on the client.
+The challenge on the metadata portion consist of reducing computational latency to memory latency.
+For example, for some intensive methods (e.g.
+a recursive directory listing), it may be worthwhile to cache a response in order to prevent computing the same list again.
+
+However, the problem lies in identifying such methods.
+Obviously, there are some methods for which caching is not worthwhile since their execution time is lower or comparable to memory latency.
+Within Panthera, six different Hadoop query methods were identified as medium to high latency.
+Panthera identifies packets that contain these methods and subsequently caches them.
+
+Currently, a standard LRU cache is used for each method. Testing of different caching methods and predictive prefetching algorithms is under way.
+
+\section{Data caching}
+
+\begin{figure}[!h]
+	\caption{Panthera Data System}
+	\centering
+		\includegraphics[scale=0.4]{assets/panthera_data_architecture.pdf}
+\end{figure}
+
+Panthera's data caching system runs on the client.
+The architecture is similar to that of the metadata caching system.
+Every request that the client makes to the DataNodes is first processed by Panthera and responded to from the cache if possible.
+
+The challenges with data caching differ significantly from those in metadata caching.
+Here, the goal is to convert network and hard drive read latency into memory latency.
+As \cite{ramcloud, hadoop} note, memory latency can be extremely low in modern systems, especially in comparison to network and hard drive latencies.
+Thus, in some respects, the goal is "easier" to attain in the data cache.
+However, for smaller files, it may be difficult to obtain latency improvement.
+
+\section{Testing Methodology}
+
+In testing Panthera, very simple strategies were adopted.
+In order to test the metadata cache, a directory listing query was repeatedly run on a directory with 100 files in it, and the latency times for responding are measured, with Panthera and without.
+The reasoning for picking such a test lies in the fact that a directory listing in Hadoop involves all six of the cached methods as well as a few non-cached methods.
+Thus, the test provides a good estimate of the efficacy of the cache.
+
+Latency times for the data cache were obtained by repeatedly querying for a one megabyte large file. The test encapsulates the most common use case for Panthera and the Hadoop File System.
+
+Tests for the metadata cache were conducted on both single node and multinode setups, whereas the datanode cache testing was run on multinode setups.
+
+\section{Results: Data caching}
+\begin{figure}[!h]
+	\caption{Panthera data cache latencies}
+	\centering
+		\includegraphics[scale=0.4]{assets/box-plot-data.pdf}
+\end{figure}
+
+The Panthera data cache showed a 95.7\% decrease in the latency in comparison to a standard deviation non-datacached installation.
+In addition, the standard of deviation was lower by a factor of 27.
+For any deployed project, maximum levels of latency must be considered when designing a system.
+Thus, the \textit{status quo} of the non-cached version severely limits possibilities within Hadoop.
+
+\section{Results: Metadata caching}
+\begin{figure}[!h]
+	\caption{Panthera metadata cache latencies}
+	\centering
+		\includegraphics[scale=0.3]{assets/box-plot-metadata.pdf}
+		\includegraphics[scale=0.3]{assets/box-plot-metadata-multinode.pdf}
+\end{figure}
+
+Metadata caching has similarly incredible results.
+\textit{Panthera} showed a 9 fold or 90.8\% improvement in the total latency as measured on the client side in a multinode setup.
+Additionally, the standard deviation is 24 fold lower in \textit{Panthera}.
+
+\section{Conclusion}
+
+\textit{Panthera} fulfils the constraints mentioned at the outset.
+It is a layer on Hadoop rather than a main repository code modification.
+Considering the latency measurements, it is clear that it has also met the requirements in terms of benefit.
+\textit{Panthera} was able to reduce data access latency by 95.7\% and reduce basic metadata access latency by 90.8\%.
+Additionally, \textit{Panthera} greatly reduced the variation in the access latency, greatly aiding in the pre-development plans and extending the capabilities of distributed computing systems such as Hadoop.
+
+\section{Future work}
+\textit{Panthera} can be extremely useful in a wide variety of applications, specifically in those which require large datasets in separate files (e.g.
+unstructured data).
+In particular, I will be concentrating on bioinformatics and integrating Panthera's caching with algorithms used in the area.
+
+I am also working on a cache-based scheduler for Hadoop.
+Essentially, it will use information about what files are available in various clients and arrange Hadoop tasks such that the use of the cache is maximized.
+
+Finally, all code associated with this paper and \textit{Panthera} will be open sourced in order to further development of Hadoop and distributed computing.
+
+\nocite{*}
 
 \bibliographystyle{plain}
-
 \bibliography{references}

### 2014-03-01 poster/paper.tex
paper added

@@
       \textcolor{black}{
-      \textbf{\fontsize{110}{40}{\selectfont Panthera: A Study of Caching in 
-      Distributed Computing}}}
+      \textbf{\fontsize{110}{200}{\selectfont Panthera: A Study of Caching in }}
+      \vspace{0.5em}
+      \textbf{\fontsize{110}{200}{\selectfont Distributed Computing}}}
     \end{center}
@@
      \begin{column}{0.32\textwidth}
+      \begin{block}{Background}
+      \textbf{Distributed computing} has seen tremendous growth in the past few years. 
+      Essentially, \textbf{it allows computational algorithms to run across multiple
+      computers (i.e. nodes)}. Most distributed systems must make files accessible over 	networks of computers. \textbf{This 
+      project aims to reduce the waiting time (known as \textit{latency}) associated
+      with obtaining files within distributed computing systems}.
+      \end{block}
       \begin{block}{Introduction}
-      Distributed computing has seen tremendous growth in the past few years.
-In particular, Hadoop, an open source project that allows developers to
+      Hadoop, an open source project that allows developers to
 write distributed applications, has found extensive use in academia and industry.
@@
 are all heavily involved with Hadoop development.
-But, Hadoop does not effectively utilize RAM memory and local storage to better performance.
-\textbf{This project develops and tests caching mechanisms, called \textbf{Panthera}, to reduce waiting time or latency in Hadoop, } thereby greatly increasing its efficiency and applicability in fields ranging from medical diagnosis to artificial intelligence.
+But, Hadoop does not effectively utilize Random Access Memory (RAM) and local storage to better performance.
+\textbf{This project develops and tests caching mechanisms, called \textbf{Panthera}, to reduce waiting time (i.e.
+latency) in Hadoop, thereby greatly increasing its efficiency and applicability in fields ranging from medical diagnosis to artificial intelligence.}
       \end{block}
@@
 	\begin{itemize}
-	 \item Develop a caching layer for the Hadoop File System that can cache both data and metadata.
-	 \item The solution must not depend on a modified Hadoop version; it should be a "drop-in" addition.
-	 \item The layer should reduce file access latency/waiting time significantly in comparison to a vanilla Hadoop installation.
+	 \item \textbf{Develop a caching layer} for the Hadoop File System that can cache both data and metadata.
+	 \item The solution must not depend on a modified Hadoop version; it should be a \textbf{"drop-in" addition}. This will allow for easy adoption and integration into existing integration.
+	 \item The layer \textbf{should reduce file access latency}/waiting time significantly in comparison to \textbf{the control group: a standard Hadoop installation}.
 	\end{itemize}
@@
 		\begin{block}{Data vs Metadata}
-			Data within the Hadoop file system consists of the actual contents of a file.
-			Metadata is the information about files and directories, e.g. time a file was created, the list of files in a directory, etc. \textbf{Panthera performs both metadata
+			\textbf{Data} within the Hadoop file system \textbf{consists of the actual contents of a file.}
+			\textbf{Metadata is the information about files and directories}, e.g. time a file was created, the list of files in a directory, etc. \textbf{Panthera performs both metadata
 			and data caching; both have significantly different technical implications
@@
 
-		\begin{block}{Panthera Metadata Architecture}
-	 		\centerline{\includegraphics[scale=1.5]{assets/panthera_metadata_large.pdf}}
-      	\end{block}
-
      \end{column}
@@
 
-      \begin{block}{Panthera Data Architecture}
-      \includegraphics[scale=1.2]{assets/panthera_data_large.pdf}
+		\begin{block}{Panthera Metadata Architecture}
+	 		\centerline{\includegraphics[scale=1.2]{assets/panthera_metadata_large.pdf}}
       	\end{block}
 
-	\begin{block}{Panthera - Implementation}
-	\begin{itemize}
-		\item \textbf{Panthera is implemented on Google's Go programming language}. The language has excellent concurrency primitives and memory management, which are perfect for
-		for software ,such as \textbf{Panthera}, that must run reliably with hundreds of clients.
-		\item \textbf{Panthera} implements a large portion of the Hadoop File System protocol, enabling it to "speak" with the DataNode and NameNode.
-	\end{itemize}
+      \begin{block}{Panthera Data Architecture}
+      \centerline{\includegraphics[scale=1.2]{assets/panthera_data_architecture.pdf}}
      \end{block}
@@
 	\begin{block}{Testing Methodology - Single Node vs Multinode}
-	Testing procedures were conducted on both \textbf{single node} and \textbf{multi-node} setups.\textbf{Single node} testing shows Panthera's performance without regard to network latency, whereas \textbf{multinode} testing incorporates network factors.
+	Testing procedures were conducted on both \textbf{single node} and \textbf{multinode} setups.\textbf{ Single node} testing shows Panthera's performance without regard to network latency, whereas \textbf{multinode} testing incorporates network factors.
 	\end{block}
@@
 	\begin{block}{Results}
-		\includegraphics[scale=1.10]{assets/box-plot-data-larger.pdf}
+		\includegraphics[scale=1.08]{assets/box-plot-data-larger.pdf}
 		\vspace{1em}
-		\includegraphics[scale=1.1]{assets/box-plot-metadata-larger.pdf}
+		\includegraphics[scale=1.08]{assets/box-plot-metadata-larger.pdf}
 		\vspace{1em}
-		\includegraphics[scale=1.1]{assets/box-plot-meta-multinode-larger.pdf}
+		\includegraphics[scale=1.08]{assets/box-plot-meta-multinode-larger.pdf}
 	\end{block}
@@
 	  \end{block}
+	\begin{block}{Panthera - Implementation}
+	\begin{itemize}
+		\item \textbf{Panthera is implemented on Google's Go programming language}. The language has excellent concurrency primitives and memory management, which are perfect for
+		for software, such as \textbf{Panthera}, that must run reliably with hundreds of clients.
+		\item \textbf{Panthera} implements a large portion of the Hadoop File System protocol, enabling it to "speak" with the DataNode and NameNode.
+	\end{itemize}
+	\end{block}
 
@@
     		\item \textbf{Noncached/Panthera}: \textbf{95.9\% improvement in latency with Panthera}, waiting time of noncached is 24.5 times higher than that of Panthera.
+There is also a \textbf{16 fold decrease in standard deviation with Panthera}.
 
@@
 
-    		\item \textbf{Metadata - noncached/Panthera: } \textbf{90.3\% improvement with the Panthera metadata cache}. Panthera (developed during this project) decreases waiting time by a factor of 8.
+    		\item \textbf{Metadata - noncached/Panthera: } \textbf{90.3\% improvement with the Panthera metadata cache}.
+Panthera (developed during this project) decreases waiting time by a factor of 8.
+\textbf{Panthera reduces standard deviation by factor of 27 (!)}.
 

### 2014-03-01 analysis/data_latency_distr.r

@@ -1,0 +1,3 @@
+pdf("datacached_latency_distribution.pdf")
+ggplot(df, aes(x=cached)) + geom_histogram(colour="chocolate1", fill="chocolate1") + xlab("Panthera (datacached) latency") + ggtitle("Panthera (Datacached) Latency Distribution, nanoseconds")
+dev.off()

### 2014-03-01 analysis/point-plot.r

@@ -4,7 +4,7 @@ cached_latency = unlist(na.omit(dat[1]))
 no_cached_latency = unlist(na.omit(dat[2]))
 
 pdf("point-plot.pdf")
-plot(cached_latency, col="dodgerblue2", pch=1, cex = 1.5, lwd=2)
-title("Metacached (Panthera) vs Non-metacached")
-points(no_cached_latency, col="chocolate1", pch=1, cex = 1.5, lwd=2)
+plot(cached_latency, col="chocolate1", pch=1, cex = 1.5, lwd=2)
+points(no_cached_latency, col="cornflowerblue", pch=1, cex = 1.5, lwd=2)
+title("Metacached (Panthera) vs Non-metacached\nMultinode")
 dev.off()

### 2014-03-01 paper/paper.tex
paper font change

@@
-\documentclass{article}
+\documentclass[12pt]{article}
 
@@
 \section{Introduction}
-Distributed computing systems are currently employed across academia and industry.
-Applications vary from market research for department stores to bioinformatics.
+Distributed computing systems are currently employed across academia and industry, with applications varying from market research for department stores to bioinformatics.
 In particular, the Hadoop distributed system \cite{hadoop} has seen tremendous growth.
 Hadoop is an open source version of the revolutionary MapReduce system developed at Google.
-With it, developers can take easily take advantage of large, multi-node clusters to solve computational problems.
+With it, developers can easily take advantage of large, multi-node clusters to solve computational problems.
 
 Current Hadoop deployments exist in the thousands of nodes at companies such as Cloudera, Facebook, Yahoo, etc.
-However, especially in larger networks, file access latency can be a significant.
+However, especially in larger networks, file access latency can be a significant issue.
 The Hadoop File System currently does not use local (i.e.
@@
 Section 3 details the architecture differences between a standard Hadoop installation and one with \textit{Panthera}.
+Sections 4 and 5 detail the differences between metadata and data caching within \textit{Panthera}.
+Results and conclusions are covered in Sections 6, 7, 8, 9, concluding with a roadmap for future development in Section 10.
 
@@
 as a layer, rather than a modification.
-This implies that \textit{Panthera} is implemented completely independently of the Hadoop codebase.
+This implies that \textit{Panthera} must implemented without a single change in the existing Hadoop codebase, i.e.
+as a "drop-in" solution.
 

### 2014-03-01 paper/statistics.tex
proj

--- 2014-02-24 poster/paper.tex
@@
+\documentclass[12pt]{article}
+\usepackage[margin=1in]{geometry}
+\begin{document}	
+	\title{Statistics}
+	\date{}
+	\maketitle
+	\begin{itemize}
+    		\item \textbf{Data (single node, noncached)}
+    		\begin{itemize}
+    			\item Mean: 12548998 nanoseconds
+    			\item Standard deviation: 42408768 nanoseconds
+    			%\item Median: 1292182 nanoseconds
+    		\end{itemize}
+    		\item \textbf{Data (single node, cached/Panthera)}
+    		\begin{itemize}
+    			\item Mean: 511185.3 nanoseconds
+			\item Standard deviation: 2572151 nanoseconds
+    		\end{itemize}
+    		\item \textbf{Noncached/Panthera}: \textbf{95.9\% improvement in latency with Panthera}, waiting time of noncached is 24.5 times higher than that of Panthera.
+There is also a \textbf{16 fold decrease in standard deviation with Panthera}.
 
-  \documentclass[final]{beamer} % beamer 3.10: do NOT use option hyperref={pdfpagelabels=false} !
-
-  %\documentclass[final,hyperref={pdfpagelabels=false}]{beamer} % beamer 3.07: get rid of beamer warnings
-  \mode<presentation> {  %% check http://www-i6.informatik.rwth-aachen.de/~dreuw/latexbeamerposter.php for examples
-    \usetheme{Berlin}    %% you should define your own theme e.g. for big headlines using your own logos 
-  }
-\setbeamercolor{block body}{bg=white, fg=black}
-\setbeamerfont{block title}{size=\Large}
-  \usepackage[english]{babel}
-  \usepackage[latin1]{inputenc}
-  \usepackage{amsmath,amsthm, amssymb, latexsym}
-  \usepackage{float}
-\usepackage{color}
-\usepackage{pgfplots}
-\usepackage{anyfontsize}
-  %\usepackage{times}\usefonttheme{professionalfonts}  % times is obsolete
-  \usefonttheme[onlymath]{serif}
-  \boldmath
-  \usepackage[orientation=portrait,size=a0,scale=1.2, debug]{beamerposter}                       % e.g. for DIN-A0 poster
-  %\usepackage[orientation=portrait,size=a1,scale=1.4,grid,debug]{beamerposter}                  % e.g. for DIN-A1 poster, with optional grid and debug output
-  %\usepackage[size=custom,width=200,height=120,scale=2,debug]{beamerposter}                     % e.g. for custom size poster
-  %\usepackage[orientation=portrait,size=a0,scale=1.0,printer=rwth-glossy-uv.df]{beamerposter}   % e.g. for DIN-A0 poster with rwth-glossy-uv printer check
-  % ...
-\definecolor{OliveGreen}{HTML}{eeeeee}
-
-\begin{document}
-
-  \begin{frame}
-    \begin{center}
-
-      \textcolor{black}{
-      \textbf{\fontsize{110}{40}{\selectfont Panthera: A Study of Caching in 
-      Distributed Computing}}}
-    \end{center}
-
-    \begin{columns}[t]
-     \begin{column}{0.32\textwidth}
-      \begin{block}{Introduction}
-      Distributed computing has seen tremendous growth in the past few years.
-In particular, Hadoop, an open source project that allows developers to
-write distributed applications, has found extensive use in academia and industry.
-For example, Yahoo, Facebook, Stanford Computer Science, etc.
-are all heavily involved with Hadoop development.
-But, Hadoop does not effectively utilize RAM memory and local storage to better performance.
-\textbf{This project develops and tests caching mechanisms, called \textbf{Panthera}, to reduce waiting time or latency in Hadoop, } thereby greatly increasing its efficiency and applicability in fields ranging from medical diagnosis to artificial intelligence.
-      \end{block}
-
-    \begin{block}{Problem}
+    		\item \textbf{Metadata (single node, noncached)}
     		\begin{itemize}
-	 \item Develop a caching layer for the Hadoop File System that can cache both data and metadata.
-	 \item The solution must not depend on a modified Hadoop version; it should be a "drop-in" addition.
-	 \item The layer should reduce file access latency/waiting time significantly in comparison to a vanilla Hadoop installation.
+    			\item Mean: 8445265 nanoseconds
+    			\item Standard deviation: 14690532 nanoseconds
+    			%\item Median: 1563000 nanoseconds
     		\end{itemize}
 
-      \end{block}
+    		\item \textbf{Metadata (single node, cached)}
+    		\begin{itemize}
+    			\item Mean: 1055012 nanoseconds
+    			\item Standard deviation: 1737020 nanoseconds
+    			%\item Median: 179500 nanoseconds
+    		\end{itemize}
 
-      \begin{block}{Hypothesis}
-       The cache layer designed (called \textbf{Panthera}) will have lower file
-       access latency when compared to a standard Hadoop installation. Additionally, it
-       will be possible to design \textbf{Panthera} such that the Hadoop source code
-       remains unaffected.
-      \end{block}
+    		\item \textbf{Metadata (multinode, noncached)}
+		\begin{itemize}
+			\item Mean: 1638184 nanoseconds 
+			\item Standard deviation: 43708.26 nanoseconds   		
+    		\end{itemize}
 
-		\begin{block}{Data vs Metadata}
-			Data within the Hadoop file system consists of the actual contents of a file.
-			Metadata is the information about files and directories, e.g. time a file was created, the list of files in a directory, etc. \textbf{Panthera performs both metadata
-			and data caching; both have significantly different technical implications
-			and challenges.}
-		\end{block}
+    		\item \textbf{Metadata (multinode, Panthera cached)}
+    		\begin{itemize}
+			\item Mean: 176868.7 nanoseconds
+			\item Standard deviation: 1196944 nanoseconds 	 		
+    		\end{itemize}
 
-		\begin{block}{Current Hadoop File System Architecture}
-			\includegraphics[scale=1]{assets/hadoop_architecture.pdf}
-		\end{block}
-
-	  \begin{block}{Panthera Architecture}
-		\includegraphics[scale=1]{assets/panthera_architecture_complete.pdf}	  
-	  \end{block}
-
-	  \begin{block}{Panthera Metadata Architecture}
-	  \includegraphics[scale=1]{assets/panthera_metadata.pdf}
-      \end{block}
-
-      \begin{block}{Panthera Data Architecture}
-      \includegraphics[scale=1]{assets/panthera_architecture.pdf}
-      \end{block}
-
-     \end{column}
-     %##########################################################################################################
-
-    \begin{column}{0.32\textwidth}
-
-	\begin{block}{Panthera - Implementation}
-	\begin{itemize}
-		\item \textbf{Panthera is implemented on Google's Go programming language}. The language has excellent concurrency primitives and memory management, which are perfect for
-		for software ,such as \textbf{Panthera}, that must run reliably with hundreds of clients.
-		\item \textbf{Panthera} implements a large portion of the Hadoop File System protocol, enabling it to "speak" with the DataNode and NameNode.
+    		\item \textbf{Metadata - noncached/Panthera: } \textbf{90.3\% improvement with the Panthera metadata cache}.
+Panthera (developed during this project) decreases waiting time by a factor of 8.
+\textbf{Panthera reduces standard deviation by factor of 27 (!)}.
     \end{itemize}
-	\end{block}
-
-	\begin{block}{Testing Methodology - Single Node vs Multinode}
-	Testing procedures were conducted on both \textbf{single node} and \textbf{multi-node} setups.\textbf{Single node} testing shows Panthera's performance without regard to network latency, whereas \textbf{multinode} testing incorporates network factors.
-	\end{block}
-
-	\begin{block}{Testing Methodology - Procedure}
-	\begin{itemize}
-		\item \textbf{Metadata}: Repeatedly request directory listing for a directory with 100 files in it.
-		\item \textbf{Data}: Repeatedly request file contents for a 1MB large file.
-	\end{itemize}
-	\end{block}
-
-	\begin{block}{Results}
-		\includegraphics[scale=1.2]{assets/box-plot-data.pdf}
-		\vspace{1em}
-		\includegraphics[scale=1.2]{assets/box-plot-metadata.pdf}
-		\vspace{1em}
-		\includegraphics[scale=1.2]{assets/box-plot-meta-multinode.pdf}
-	\end{block}
-
-      \end{column}
-      %##########################################################
-
-      \begin{column}{0.32\textwidth}
-
-	  \begin{block}{Latency Types}
-		%discuss the types of latency and make the comparison
-		%between hard disk latency and regular lateny
-		In the Hadoop File System Architecture, there are several types of latency involved:
-		\begin{itemize}
-		\item \textbf{Network latency: } Information shared between client and server (e.g.
-between a DataNode and a client) has an attached waiting time.
-With \textbf{Panthera}, this latency is eliminated since the data is cached on the client.
-		\item \textbf{Hard drive latency: } Reading from the hard drive has very significant latency. Hadoop generally reads from the hard drive when accessing files.
-		\item \textbf{Memory latency: } Reading from RAM memory has an extremely small
-		waiting time associated with it.
-\textbf{Memory waiting time is less than 1/1000th hard drive waiting time.
-Panthera converts hard drive latency to memory latency, thereby greatly reducing file access time.}
-		\end{itemize}
-	  \end{block}
-
-    \begin{block}{Statistics}
-    \begin{itemize}
+\end{document}\begin{itemize}
     		\item \textbf{Data (single node, noncached)}
@@
 
-    		\item \textbf{Metadata - noncached/Panthera: } \textbf{90.3\% improvement with the Panthera metadata cache}. Panthera (developed during this project) decreases waiting time by a factor of 8.
-
-	\end{itemize}    	    
-    \end{block}
-
-    	\begin{block}{Conclusion}
-	The results validate the hypothesis presented.
-Panthera was successfully developed as a "drop-in" solution and it was able to reduce file access latency in comparison to a standard Hadoop installation.
-In particular, \textbf{the Panthera data cache led to a 95.9\% or 9 fold decrease in waiting time} and the \textbf{the Panthera metadata cache also gave a 90.3\% or 8 fold decrease in latency}.
-Additionally, the \textbf{the cache developed during this project had a far lower standard deviation for access times and fewer outliers}.
-For large scale systems such as those in use at major companies and universities, such a drastic decrease in latency can have incredible effects on the efficiency of distributed computing systems.
-
-	\end{block}
-
-	\begin{block}{Applications}
-	Hadoop has diverse applications and since Panthera was created as a "drop-in" solution, the \textbf{applications span an extremely wide range of disciplines}.
-	\begin{itemize}
-		\item Bioinformatics (e.g. CrossBow from Johns Hopkins University)
-		\item Numerical/scientific computing 
-		\item Image processing
-		\item Artificial Intelligence/Machine Learning (e.g. Apache Mahout)
-		\item Market Research (e.g. Target)
-	\end{itemize}
-	\end{block}
-
-	\begin{block}{Future Work}
-	\begin{itemize}
-		\item \textbf{All code related to this project will be open sourced} to further development in distributed computing.
-		\item Currently working on testing a \textbf{bioinformatics toolkit with Panthera}; initial results are very promising and show significant reduction in total running time.
-		\item Smart scheduling using Panthera; arranges tasks so that the cache is used most efficiently.
-
-	\end{itemize}
-	\end{block}
-      \end{column}
-    \end{columns}
-  \end{frame}
-
-  \end{document}
+    		\item \textbf{Metadata - noncached/Panthera: } \textbf{90.3\% improvement with the Panthera metadata cache}.
+Panthera (developed during this project) decreases waiting time by a factor of 8.
+\textbf{Panthera reduces standard deviation by factor of 27 (!)}.

### 2014-05-02 poster/paper.tex
poster fixes

@@
   \usepackage{float}
+  \usepackage{booktabs}
+\usepackage{mathptmx}
 \usepackage{color}
@@
   \boldmath
-  \usepackage[orientation=portrait,size=a0,scale=1.15, debug]{beamerposter}                       % e.g. for DIN-A0 poster
+  \usepackage[orientation=portrait,size=a0,scale=1.10, debug]{beamerposter}                       % e.g. for DIN-A0 poster
   %\usepackage[orientation=portrait,size=a1,scale=1.4,grid,debug]{beamerposter}                  % e.g. for DIN-A1 poster, with optional grid and debug output
@@
 But, Hadoop does not effectively utilize Random Access Memory (RAM) and local storage to better performance.
-\textbf{This project develops and tests caching mechanisms, called \textbf{Panthera}, to reduce waiting time (i.e.
-latency) in Hadoop, thereby greatly increasing its efficiency and applicability in fields ranging from medical diagnosis to artificial intelligence.}
+\textbf{This project develops and tests scheduling and caching mechanisms to reduce waiting time (i.e.
+latency) in Hadoop, thereby greatly increasing Hadoop's efficiency and applicability in fields ranging from medical diagnosis to artificial intelligence.}
+      \end{block}
+
+	\begin{block}{Caching}
+	There is significant waiting time/latency associated with retrieving files from distributed filesystems (Griffioen 1994). \textbf{Caches} are used to store
+	some files locally so that they can be accessed much more quickly. Panthera implements a cache and a cache-based scheduler for Hadoop.
 	\end{block}
@@
 	\begin{itemize}
-	 \item \textbf{Develop a caching layer} for the Hadoop File System that can cache both data and metadata.
-	 \item The solution must not depend on a modified Hadoop version; it should be a \textbf{"drop-in" addition}. This will allow for easy adoption and integration into existing integration.
+	 \item \textbf{Develop a caching layer} for the Hadoop Distributed File System that can cache \textbf{both data and metadata}.
+	 \item The solution should be a \textbf{"drop-in" addition}, allowing for easy adoption and integration into existing systems.
 	 \item The layer \textbf{should reduce file access latency}/waiting time significantly in comparison to \textbf{the control group: a standard Hadoop installation}.
+	 \item \textbf{Develop a scheduler} for the Hadoop which \textbf{schedules jobs based on what is available in the cache}.
+In addition, it should be be able to \textbf{prefetch files that could be needed in the future}.
 	\end{itemize}
@@
        will be possible to design \textbf{Panthera} such that the Hadoop source code
-       remains unaffected.
+       remains unaffected. It will also be feasible to create a cache-based 
+       scheduler for Hadoop.
       \end{block}
@@
 
-	  \begin{block}{Panthera Architecture}
-		\centerline{\includegraphics[scale=1.5]{assets/panthera_architecture_complete_larger.pdf}}	  
-	  \end{block}
-
      \end{column}
@@
     \begin{column}{0.32\textwidth}
+		\begin{block}{Panthera Architecture}
+		\vspace{1em}
+		\centerline{\includegraphics[scale=1.5]{assets/v2/panthera_hadoop_arch.pdf}}	  
+	  \end{block}   	  
+		%\begin{block}{Panthera Metadata Architecture}
+	 		%\centerline{\includegraphics[scale=1.2]{assets/panthera_metadata_large.pdf}}
+      	%\end{block}
 
-		\begin{block}{Panthera Metadata Architecture}
-	 		\centerline{\includegraphics[scale=1.2]{assets/panthera_metadata_large.pdf}}
+      %\begin{block}{Panthera Data Architecture}
+      %\centerline{\includegraphics[scale=1.2]{assets/panthera_data_architecture.pdf}}
+     %\end{block}
+
+	\begin{block}{Panthera - Implementation}
+	\begin{itemize}
+		\item \textbf{Panthera is implemented on Google's Go programming language} which has excellent concurrency primitives and memory management, making it perfect for
+		for software like \textbf{Panthera} that must run reliably with hundreds of clients.
+	\end{itemize}
 	\end{block}
 
-      \begin{block}{Panthera Data Architecture}
-      \centerline{\includegraphics[scale=1.2]{assets/panthera_data_architecture.pdf}}
-     \end{block}
-
-	\begin{block}{Testing Methodology - Single Node vs Multinode}
-	Testing procedures were conducted on both \textbf{single node} and \textbf{multinode} setups.\textbf{ Single node} testing shows Panthera's performance without regard to network latency, whereas \textbf{multinode} testing incorporates network factors.
-	\end{block}
-
-	\begin{block}{Testing Methodology - Procedure}
+	\begin{block}{Latency Testing Methodology}
+	Client and server (i.e. NameNode or DataNode) are on separate machines
+	for both metadata and data cache testing.
 	\begin{itemize}
@@
 	\end{itemize}
+
+	\end{block}
+		\begin{block}{Latency Results}
+		\vspace{1em}
+		\includegraphics[scale=1]{assets/v2/multinode_meta_box_plot.pdf}
+		\vspace{1.5em}
+		\includegraphics[scale=0.98]{assets/v2/getter_boxplot.pdf}
+		%\vspace{1em}
+		%\includegraphics[scale=0.98]{assets/v2}
 	\end{block}
 
-	\begin{block}{Results}
-		\includegraphics[scale=1.08]{assets/box-plot-data-larger.pdf}
-		\vspace{1em}
-		\includegraphics[scale=1.08]{assets/box-plot-metadata-larger.pdf}
-		\vspace{1em}
-		\includegraphics[scale=1.08]{assets/box-plot-meta-multinode-larger.pdf}
+	\begin{block}{Procedure Testing Methodology}
+	Panthera was also tested with running times of algorithms running on Hadoop (as opposed to simply downloading files). The input sample was a repeated copyof "The 
+	Adventures of Sherlock Holmes" (192 MB).
+	\begin{itemize}
+	\item \textbf{WordCount}: The classic distributed algorithm that counts occurrences
+	of all words in a given text.
+	\item \textbf{ReferenceCount}: Counts occurrences of words in a given text, but 
+	only considers the words that are in a given reference file (i.e. dictionary).
+	\end{itemize}
 	\end{block}
@@
       \begin{column}{0.32\textwidth}
+		\begin{block}{Procedure Results}
+		\vspace{1em}
+		\includegraphics[scale=1]{assets/v2/refgetter_boxplot.pdf}
+		\end{block}
 
@@
 		%between hard disk latency and regular lateny
-		In the Hadoop File System Architecture, there are several types of latency involved:
+		In the Hadoop File System Architecture, there are several types of latency involved (listed from highest to lowest latency):
 		\begin{itemize}
-		\item \textbf{Network latency: } Information shared between client and server (e.g.
-between a DataNode and a client) has an attached waiting time.
-With \textbf{Panthera}, this latency is eliminated since the data is cached on the client.
-		\item \textbf{Hard drive latency: } Reading from the hard drive has very significant latency. Hadoop generally reads from the hard drive when accessing files.
+		\item \textbf{Network latency}
+		\item \textbf{Hard drive latency}
 		\item \textbf{Memory latency: } Reading from RAM memory has an extremely small
-		waiting time associated with it.
-\textbf{Memory waiting time is less than 1/1000th hard drive waiting time.
-Panthera converts hard drive latency to memory latency, thereby greatly reducing file access time.}
-		\end{itemize}
-	  \end{block}
-	\begin{block}{Panthera - Implementation}
-	\begin{itemize}
-		\item \textbf{Panthera is implemented on Google's Go programming language}. The language has excellent concurrency primitives and memory management, which are perfect for
-		for software, such as \textbf{Panthera}, that must run reliably with hundreds of clients.
-		\item \textbf{Panthera} implements a large portion of the Hadoop File System protocol, enabling it to "speak" with the DataNode and NameNode.
+		waiting time associated with it. \textbf{Panthera converts hard drive latency to memory latency, thereby greatly reducing file access time.}
 		\end{itemize}
@@
     \begin{block}{Statistics}
-    \begin{itemize}
-    		\item \textbf{Data (single node, noncached)}
-    		\begin{itemize}
-    			\item Mean: 12548998 nanoseconds
-    			%\item Median: 1292182 nanoseconds
-    		\end{itemize}
-    		\item \textbf{Data (single node, cached/Panthera)}
-    		\begin{itemize}
-    			\item Mean: 511185.3 nanoseconds
-    			%\item Median: 149004 nanoseconds
-    		\end{itemize}
-    		\item \textbf{Noncached/Panthera}: \textbf{95.9\% improvement in latency with Panthera}, waiting time of noncached is 24.5 times higher than that of Panthera.
-There is also a \textbf{16 fold decrease in standard deviation with Panthera}.
+	\begin{table}
+	\setlength{\tabcolsep}{25pt}
+	\centering
+	\begin{tabular}{lc}
+		\toprule
+		\textbf{Type} & \textbf{Mean time improvement} \\
+		\midrule
+		Metadata   & 862.6\% \\
+		Data       & 863.4\% \\
+		Word count & \phantom025.1\% \\
+		Ref. count & 331.6\% \\
+	\bottomrule
+	\end{tabular}
+	\caption{Improvement with Panthera in the total amount of time required to execute algorithms/requirements (compared with status quo)}
+	\end{table}
 
-    		\item \textbf{Metadata (single node, noncached)}
-    		\begin{itemize}
-    			\item Mean: 8445265 nanoseconds
-    			%\item Median: 1563000 nanoseconds
-    		\end{itemize}
+ 	\vspace{1em}
+	\begin{table}
+	\setlength{\tabcolsep}{25pt}
+	\centering
+	\begin{tabular}{lc}
+		\toprule
+		\textbf{Type} & \textbf{Standard deviation contraction} \\
+		\midrule
+		Metadata   & 27.38x \\
+		Data       & 11.14x \\
+		Word count &  0.78x \\
+		Ref. count & 22.91x \\
+	\bottomrule
+	\end{tabular}
+	\caption{Reduction obtained with Panthera in the standard deviation of running times, i.e. a contraction of 9x implies that the standard deviation was reduced
+	by a factor of 9 with Panthera (compared to the status quo).}
+	\end{table}
 
-    		\item \textbf{Metadata (single node, cached)}
-    		\begin{itemize}
-    			\item Mean: 1055012 nanoseconds
-    			%\item Median: 179500 nanoseconds
-    		\end{itemize}
-
-    		\item \textbf{Metadata (multinode, noncached)}
-		\begin{itemize}
-			\item Mean: 1638184 nanoseconds    		
-    		\end{itemize}
-
-    		\item \textbf{Metadata (multinode, Panthera cached)}
-    		\begin{itemize}
-			\item Mean: 176868.7 nanoseconds    		
-    		\end{itemize}
-
-    		\item \textbf{Metadata - noncached/Panthera: } \textbf{90.3\% improvement with the Panthera metadata cache}.
-Panthera (developed during this project) decreases waiting time by a factor of 8.
-\textbf{Panthera reduces standard deviation by factor of 27 (!)}.
-
-	\end{itemize}    	    
     \end{block}
@@
     	The results validate the hypothesis presented.
-Panthera was successfully developed as a "drop-in" solution and it was able to reduce file access latency in comparison to a standard Hadoop installation.
-In particular, \textbf{the Panthera data cache led to a 95.9\% or 9 fold decrease in waiting time} and the \textbf{the Panthera metadata cache also gave a 90.3\% or 8 fold decrease in latency}.
-Additionally, the \textbf{the cache developed during this project had a far lower standard deviation for access times and fewer outliers}.
-For large scale systems such as those in use at major companies and universities, such a drastic decrease in latency can have incredible effects on the efficiency of distributed computing systems.
+    	\begin{itemize}
+    		\item Panthera was successfully developed as a "drop-in" solution
+    		\item It was able to reduce latency in comparison to a standard Hadoop installation. 
+    		\item A scheduler was also implemented which uses cache information to schedule
+    		Hadoop jobs.
+    		\item Metadata latency was \textbf{decreased by a factor of 7.36}
+    		\item Data latency was \textbf{decreased by a factor of 7.634}
+    		\item The running of time of WordCount (an algorithm with very few file accesses) was \textbf{improved by 25\%} with the code left unmodified.
+		\item The running time of RefCount was \textbf{improved by 331.6\%}.
+    	\end{itemize}
+
+    	As a "drop-in" solution, Panthera can be employed in a variety of 
+    	applications to improve performance without code modification. With the 
+    	demonstrated time and resource savings, Panthera can have incredible 
+    	impacts in distributed systems and in specific applications.
 

### 2014-05-02 src/main/main.go

@@ -4,7 +4,7 @@ import (
 	"net"
 	"hdfs_requests"
 	"util"
-	"io/ioutil"
+	//"io/ioutil"
 	"log"
 	"caches"
 	"fmt"
@@ -183,8 +183,9 @@ func main() {
 	cacheSet.GetListingCache = caches.NewGetListingCache(getListingCacheSize)
 	
 	//disable the metadata cache for now
-	//cacheSet.Disable()
+	cacheSet.Disable()
 
+  /*
 	server, err := net.Listen("tcp", config.ServerHost + ":" + config.ServerPort)
 	log.SetOutput(ioutil.Discard)
 	
@@ -193,7 +194,7 @@ func main() {
 		fmt.Println("Error ocurred in initializing the utilities: ", err)
 		return
 	}
-	util.TempLogger.Println("init()ed temporary logging")
+	util.TempLogger.Println("init()ed temporary logging") */
 
 	/* setup the data cache */
 	dataCacheSize := 10
@@ -216,7 +217,9 @@ func main() {
 	runDataNodeMap(dataNodeMap, dataCache)
 
 	/* start the server */
-	loop(server, cacheSet, &dataNodeMap)
+	//loop(server, cacheSet, &dataNodeMap)
+
+  for {}
 }
 
 

### 2014-05-02 paper/paper.tex

@@
 %
-\title{Panthera: A Study of Caching in Distributed Computing}
+\title{Panthera: Caching in Distributed Computing Systems}
 \author{Dhaivat Pandya \\ Appleton North High School}
@@
 latency).
-In this project, a system, named Panthera, was developed, which creates a caching system for Hadoop metadata and data.
-Panthera caches both information about files and the file contents, thereby reducing waiting time in downloading and accessing the files and related information.
+Thus, there is a very significant wastage of available computational resources.
+In this project, a system, named Panthera, was developed which reduces waiting time and conserves computational resources by using \textit{caching}, i.e.
+retaining downloaded files in local memory rather than discarding them, thereby reducing access time.
+Panthera caches both information about files (i.e.
+metadata) and the file contents (i.e.
+data).
 It runs with an unmodified version of Hadoop, meaning that it can integrate easily into existing architecture.
-Panthera was tested using both single node and multi-
-node setups.
-The results showed an 90.8\% decrease in waiting time with the Panthera metadata cache, and a 95.7\% decrease with the Panthera data cache.
+Panthera was tested using both single node and multi-node setups.
+The results showed a 90.8\% decrease in waiting time with the Panthera metadata cache and a 95.7\% decrease with the Panthera data cache.
 Such drastic decreases in latency can greatly increase the efficiency of existing Hadoop applications and also allow the creation of algorithms that were previously not feasible.
-Panthera has numerous applications in fields ranging from market research to bioinformatics.
-All code associated with the project will be open sourced to further development in the area.
+Panthera has numerous applications in fields ranging from bioinformatics and medicine to market research.
+All code associated with this project will be open sourced to further development in the area of distributed computing.
 }
@@
 \section{Introduction}
-Distributed computing systems are currently employed across academia and industry, with applications varying from market research for department stores to bioinformatics.
+As computational problems and their associated datasets have grown in size, it has become necessary to take advantage of distributed computing systems to solve them.
+Such systems allow programs to scale from one computer (i.e.
+one \textit{node}) to thousands of computers with few or no changes in the codebase.
 In particular, the Hadoop distributed system \cite{hadoop} has seen tremendous growth.
-Hadoop is an open source version of the revolutionary MapReduce system developed at Google.
+Hadoop is an open source version of the revolutionary MapReduce system developed at Google \cite{mapreduce}.
 With it, developers can easily take advantage of large, multi-node clusters to solve computational problems.
 
+In order to run algorithms across a network of computers, access to the dataset must be available to all nodes within a network.
+To facilitate this, Hadoop uses the Hadoop Distributed File System (HDFS) \cite{hdfs}, which allows one node to download files from other nodes.
+
 Current Hadoop deployments exist in the thousands of nodes at companies such as Cloudera, Facebook, Yahoo, etc.
-However, especially in larger networks, file access latency can be a significant issue.
+Especially in such large networks, \textit{access latency}, or the waiting time associated with accessing files, is a significant concern.
+Latency in distributed systems can have significant effects, and a reduction of the same comes with tremendous benefits.
+As the RAMCloud project has outlined \cite{ramcloud}, low latency can greatly extend the applications of distributed computing systems.
+For example, tree traversal algorithms are largely impractical in Hadoop \cite{ramcloud}, however, with lower latency, such algorithms may become useful and applicable.
 The Hadoop File System currently does not use local (i.e.
@@
 
-Latency in distributed systems can have significant effects, and a reduction of the same comes with tremendous benefits.
-As the RAMCloud project has outlined \cite{ramcloud}, low latency can greatly extend the applications of distributed computing systems.
-For example, tree traversal algorithms are largely impractical in Hadoop, however, with lower latency, such algorithms may become useful and applicable.
+In this paper, I present \textit{Panthera}, a \textit{cache layer} for Hadoop.
+Panthera functions by retaining recently accessed files in RAM, thus, when the files are needed once again, they do not have to be downloaded or read from the hard drive, thereby greatly reducing file access latency.
+Thus, using \textit{Panthera} can save on computational resources and reduce total execution time for a large variety of applications.
 
-In this paper, we discuss \textit{Panthera}, a cache layer for Hadoop.
 In Section 2, we discuss the specific challenges associated with such a project.
@@
 
-\section{Constraints}
+\section{Design Constraints}
 
@@
 as a "drop-in" solution.
+In order for \textit{Panthera} to remain independent of the Hadoop codebase, I had to reimplement a significant portion of the Hadoop File System network protocol, which in and of itself was an engineering challenge.
 
@@
 Finally, data and metadata request latency should be significantly with \textit{Panthera} in comparison to a standard Hadoop installation.
+
+\section{Metadata and Data}
+
+Within the Hadoop File System (HDFS), there is a clear distinction between \textit{metadata} and \textit{data} \cite{hdfs}.
+
+Metadata refers to information about a file or a directory, e.g.
+file size, file names, on which DataNode the contents of a file are located, etc.
+For HDFS, the NameNode handles all metadata.
+For every request, the NameNode recomputes metadata.
+\textit{Panthera} can reduce this computation by caching the results of a previous request, thereby conserving resources.
+
+Data refers to the actual contents of a file.
+Within HDFS, each file is split into 64 MB large \textit{blocks}, which are stored on DataNodes.
+Here, \textit{Panthera} serves to reduce access time by holding recently used blocks in RAM locally.
+
+Metadata and data caching provide significantly different technical challenges as the former involves conversion of computational latency to memory latency, whereas the latter involves a conversion of network and hard drive latency to memory latency.
+Note: memory latency is by far the lowest of all three in consideration.
+In many cases \cite{ramcloud} it is 1/1000th that of hard drive latency, which is the closest competitor.
 
@@
 
-Figure 1 outlines the architecture differences between a standard Hadoop File System architecture and an installation with Panthera.
+Figure 1 (see next page) outlines the architecture differences between a standard Hadoop File System architecture and an installation with Panthera.
 
@@
 
-Currently, a standard LRU cache is used for each method. Testing of different caching methods and predictive prefetching algorithms is under way.
+Currently, a standard LRU cache \cite{cache} is used for each method. Testing of different caching methods and predictive prefetching algorithms is under way.
 
@@
 
-Tests for the metadata cache were conducted on both single node and multinode setups, whereas the datanode cache testing was run on multinode setups.
+Tests for the metadata cache were conducted on both single node and multinode setups, whereas the data cache testing was run on multinode setups.
 
@@
 
-The Panthera data cache showed a 95.7\% decrease in the latency in comparison to a standard deviation non-datacached installation.
-In addition, the standard of deviation was lower by a factor of 27.
+\textbf{The Panthera data cache showed a 95.7\% decrease in the latency in comparison to a standard deviation non-datacached installation.
+In addition, the standard of deviation was lower by a factor of 27}.
 For any deployed project, maximum levels of latency must be considered when designing a system.
 Thus, the \textit{status quo} of the non-cached version severely limits possibilities within Hadoop.
+With \textit{Panthera}, not only is latency radically reduced, but also the variation in the latency is minimized.
 
@@
 Metadata caching has similarly incredible results.
-\textit{Panthera} showed a 9 fold or 90.8\% improvement in the total latency as measured on the client side in a multinode setup.
-Additionally, the standard deviation is 24 fold lower in \textit{Panthera}.
+\textbf{\textit{Panthera} showed a 9 fold or 90.8\% improvement in the total latency as measured on the client side in a multinode setup}.
+Additionally, the standard deviation is 24 fold lower with\textit{Panthera}.
+Thus, this project has been able to minimize the latency averages as well as spreads to an extremely significant degree.
 
@@
 
-\textit{Panthera} fulfils the constraints mentioned at the outset.
+\textit{Panthera} fulfills the constraints mentioned at the outset.
 It is a layer on Hadoop rather than a main repository code modification.
@@
 \textit{Panthera} was able to reduce data access latency by 95.7\% and reduce basic metadata access latency by 90.8\%.
-Additionally, \textit{Panthera} greatly reduced the variation in the access latency, greatly aiding in the pre-development plans and extending the capabilities of distributed computing systems such as Hadoop.
+Also, \textit{Panthera} was able to minimize the spread of latency times in both data and metadata requests.
+Thus, \textit{Panthera} opens up great possibilities within Hadoop due to the fact that algorithms that were previously impractical (e.g.
+distributed tree traversal algorithms) may now be implemented.
 
@@
 I am also working on a cache-based scheduler for Hadoop.
-Essentially, it will use information about what files are available in various clients and arrange Hadoop tasks such that the use of the cache is maximized.
+Essentially, it will intelligently use information about what files are available in various clients and arrange Hadoop tasks such that the use of the cache is maximized.
 

### 2014-05-04 poster/paper.tex
poster fixes

@@
 \setbeamerfont{block title}{size=\Large}
+\usepackage{lmodern}
+
   \usepackage[english]{babel}
@@
       Essentially, \textbf{it allows computational algorithms to run across multiple
-      computers (i.e. nodes)}. Most distributed systems must make files accessible over 	networks of computers. \textbf{This 
+      computers (i.e. \textit{nodes})}. Most distributed systems must make files accessible over 	networks of computers. \textbf{This 
       project aims to reduce the waiting time (known as \textit{latency}) associated
@@
       \begin{block}{Introduction}
-      Hadoop, an open source project that allows developers to
-write distributed applications, has found extensive use in academia and industry.
-For example, Yahoo, Facebook, Stanford Computer Science, etc.
+      \textbf{\textit{Hadoop}} is an open source and heavily used distributed computing sytem.
+It has found extensive applications in academia and industry.
+For example, Yahoo, Facebook, Microsoft, etc.
 are all heavily involved with Hadoop development.
@@
 
-	\begin{block}{Caching}
-	There is significant waiting time/latency associated with retrieving files from distributed filesystems (Griffioen 1994). \textbf{Caches} are used to store
-	some files locally so that they can be accessed much more quickly. Panthera implements a cache and a cache-based scheduler for Hadoop.
+	\begin{block}{Caching Theory}
+	There is significant waiting time/latency associated with retrieving files from distributed filesystems. \textbf{Caches} are used to store
+	some files in local RAM so that they can be accessed much more quickly. Panthera implements a cache and a cache-based scheduler for Hadoop.
 	\end{block}
@@
 	 \item \textbf{Develop a caching layer} for the Hadoop Distributed File System that can cache \textbf{both data and metadata}.
-	 \item The solution should be a \textbf{"drop-in" addition}, allowing for easy adoption and integration into existing systems.
+	 \item The solution should be a \textbf{``drop-in'' addition}, allowing for easy adoption and integration into existing systems.
 	 \item The layer \textbf{should reduce file access latency}/waiting time significantly in comparison to \textbf{the control group: a standard Hadoop installation}.
@@
 		\begin{block}{Data vs Metadata}
-			\textbf{Data} within the Hadoop file system \textbf{consists of the actual contents of a file.}
-			\textbf{Metadata is the information about files and directories}, e.g. time a file was created, the list of files in a directory, etc. \textbf{Panthera performs both metadata
+			\begin{itemize}
+				\item \textbf{Data} within the Hadoop file system \textbf{consists of the actual contents of a file.}
+				\item \textbf{Metadata is information \textit{about} files and directories}, e.g. time a file was created, the list of files in a directory, etc.
+				\item \textbf{Panthera performs both metadata
 			and data caching; both have significantly different technical implications
 			and challenges.}
+			\end{itemize}
 		\end{block}
@@
 		\begin{block}{Current Hadoop File System Architecture}
-			\centerline{\includegraphics[scale=1.5]{assets/vanilla_hadoop.pdf}}
+			\centerline{\includegraphics[scale=1.4]{assets/v2/vanilla_hadoop_arch.pdf}}
+		\end{block}
+
+		\begin{block}{Latency Types - Theory}
+		%discuss the types of latency and make the comparison
+		%between hard disk latency and regular lateny
+		In the Hadoop File System Architecture, there are several types of latencies involved (listed from highest to lowest magnitude):
+		\begin{itemize}
+		\item \textbf{Network latency}
+		\item \textbf{Hard drive latency}
+		\item \textbf{Memory latency: } Reading from RAM memory has an extremely small
+		waiting time associated with it ($<$ 1/100th that of the hard drive). \textbf{Panthera converts hard drive and network latency to memory latency, thereby greatly reducing file access time.}
+		\end{itemize}
 	  	\end{block}
@@
 		\begin{block}{Panthera Architecture}
-		\vspace{1em}
-		\centerline{\includegraphics[scale=1.5]{assets/v2/panthera_hadoop_arch.pdf}}	  
+		\vspace{0.5em}
+		\centerline{\includegraphics[scale=1.4]{assets/v2/panthera_hadoop_arch.pdf}}	  
 	  \end{block}   	  
@@
       	%\end{block}
+
+		\begin{block}{Panthera Scheduling Algorithm Theory}
+		\begin{itemize}
+			\item \textbf{Problem}: Given $\mathbf{J} = J_1, J_2, ... ,J_k$ jobs to be run
+			on a Hadoop cluster with caches $\mathbf{C} = C_1, C_2, ..., C_n$, what ordering will best take advantage of the caches?
+			\item \textbf{Panthera job score algorithm}: Each $J_i$ accesses a set of blocks $F_i$
+			and each $C_i$ holds a set of blocks $B_i$.
+			Every $J_i$ is assigned a score $x_i$:
+			\begin{equation}
+				x_i = \displaystyle{\frac{\left\vert{F_i \cap B_t}\right\vert}{ \left\vert{B_t}\right\vert}} \; \textrm{where} \;
+				B_t = \bigcup_{j = 1}^{n} B_j
+			\end{equation}
+			\item \textbf{Panthera scheduling algorithm}:
+			Sort $\mathbf{J}$ in descending order with respect to the $x_i$ associated
+			with each $J_i$. While $J_i$ is running, Panthera data for $J_{i+1}$ to further reduce latency.
+		\end{itemize}
+		\end{block}
 
@@
 	\begin{itemize}
-		\item \textbf{Panthera is implemented on Google's Go programming language} which has excellent concurrency primitives and memory management, making it perfect for
+		\item \textbf{Panthera is implemented in Google's Go programming language} which has excellent concurrency primitives and memory management, making it perfect for
 		for software like \textbf{Panthera} that must run reliably with hundreds of clients.
@@
 	\begin{block}{Latency Testing Methodology}
-	Client and server (i.e. NameNode or DataNode) are on separate machines
-	for both metadata and data cache testing.
+	Client and server are on separate machiner for latency testing.
 	\begin{itemize}
 		\item \textbf{Metadata}: Repeatedly request directory listing for a directory with 100 files in it.
-		\item \textbf{Data}: Repeatedly request file contents for a 1MB large file.
+		\item \textbf{Data}: Repeatedly request file contents for a 64MB large file (64MB = 1 block in Hadoop).
 	\end{itemize}
@@
 	\begin{block}{Procedure Testing Methodology}
-	Panthera was also tested with running times of algorithms running on Hadoop (as opposed to simply downloading files). The input sample was a repeated copyof "The 
-	Adventures of Sherlock Holmes" (192 MB).
+	Panthera was also tested with running times of algorithms running on Hadoop (as opposed to only downloading data or metadata). The input sample to both algorithms tested was a repeated copy of "The 
+	Adventures of Sherlock Holmes" (192 MB). Algorithms tested: 
 	\begin{itemize}
@@
 		\includegraphics[scale=1]{assets/v2/refgetter_boxplot.pdf}
-		\end{block}
-
-	  \begin{block}{Latency Types}
-		%discuss the types of latency and make the comparison
-		%between hard disk latency and regular lateny
-		In the Hadoop File System Architecture, there are several types of latency involved (listed from highest to lowest latency):
-		\begin{itemize}
-		\item \textbf{Network latency}
-		\item \textbf{Hard drive latency}
-		\item \textbf{Memory latency: } Reading from RAM memory has an extremely small
-		waiting time associated with it. \textbf{Panthera converts hard drive latency to memory latency, thereby greatly reducing file access time.}
-		\end{itemize}
 		\end{block}
@@
 		Data       & 11.14x \\
-		Word count &  0.78x \\
+		Word count &  \phantom01.02x \\
 		Ref. count & 22.91x \\
@@
 	\end{tabular}
-	\caption{Reduction obtained with Panthera in the standard deviation of running times, i.e. a contraction of 9x implies that the standard deviation was reduced
-	by a factor of 9 with Panthera (compared to the status quo).}
+	\caption{Reduction obtained with Panthera in the standard deviation of running times, i.e. a \textbf{contraction of 9x implies that the standard deviation was reduced
+	by a factor of 9 with Panthera (compared to the status quo).}}
 	\end{table}
@@
     		Hadoop jobs.
-    		\item Metadata latency was \textbf{decreased by a factor of 7.36}
-    		\item Data latency was \textbf{decreased by a factor of 7.634}
-    		\item The running of time of WordCount (an algorithm with very few file accesses) was \textbf{improved by 25\%} with the code left unmodified.
-		\item The running time of RefCount was \textbf{improved by 331.6\%}.
+    		\item Metadata latency was \textbf{decreased by a factor of 8.626}
+    		\item Data latency was \textbf{decreased by a factor of 8.634}
+    		\item The running of time of WordCount (an algorithm with very few file accesses) was \textbf{improved by 25.1\%} with the code left unmodified.
+		\item The running time of RefCount was \textbf{improved by 331.6\%} (equivalently, decreased by a factor of 3.316).
     	\end{itemize}
 
-    	As a "drop-in" solution, Panthera can be employed in a variety of 
-    	applications to improve performance without code modification. With the 
-    	demonstrated time and resource savings, Panthera can have incredible 
-    	impacts in distributed systems and in specific applications.
+    Thus, Panthera can significantly improve existing distributed systems
+    without any code modifications and also enable developers to 
+    implement algorithms that were previously impractical.
 
@@
 	\begin{block}{Applications}
-	Hadoop has diverse applications and since Panthera was created as a "drop-in" solution, the \textbf{applications span an extremely wide range of disciplines}.
+	Panthera can be applied in a vast range of disciplines since the Hadoop
+	distributed system is in widespread use. A few examples follow: 
 	\begin{itemize}
-		\item Bioinformatics (e.g. CrossBow from Johns Hopkins University)
-		\item Numerical/scientific computing 
-		\item Image processing
-		\item Artificial Intelligence/Machine Learning (e.g. Apache Mahout)
-		\item Market Research (e.g. Target)
+		\item Machine Learning
+			\begin{itemize}
+			\item Panthera was tested with Apache Mahout (a library of machine learning algorithms for Hadoop) and overall there was a 15\% reduction in total running time of
+			the k-means algorithm with two clusters.
+			\end{itemize}
+
+		\item Bioinformatics
+			\begin{itemize}
+				\item BioDoop: a bioinformatics framework for Hadoop.
+				\item Distributed sequence alignment
+			\end{itemize}
+
+		\item Predictive Analytics	
+			\begin{itemize}
+				\item Hadoop is heavily used for predictive analytics. Panthera can
+				be used to reduce the running time of queries and schedule queries
+				for maximal efficiency, reducing total turnaround time.
+			\end{itemize}
+		\item Image Processing
+		\item Numerical/Scientific Computing
+		\begin{itemize}
+			\item Apache Hama is a matrix computational package for Hadoop. Applications using it can be made more efficient without any modifications to their codebases due to Panthera's ``drop-in'' nature,.
+			\end{itemize}
 	\end{itemize}

### 2014-05-05 poster/paper.tex
poster changes

@@
      \begin{column}{0.32\textwidth}
-      \begin{block}{Background}
+      \begin{block}{Introduction}
       \textbf{Distributed computing} has seen tremendous growth in the past few years.
-      Essentially, \textbf{it allows computational algorithms to run across multiple
-      computers (i.e. \textit{nodes})}. Most distributed systems must make files accessible over 	networks of computers. \textbf{This 
-      project aims to reduce the waiting time (known as \textit{latency}) associated
-      with obtaining files within distributed computing systems}.
-      \end{block}
-      \begin{block}{Introduction}
+Essentially, it allows computational algorithms to execute over a network (i.e.
+\textit{\textbf{cluster}}) of computers. This project is focused on the \textbf{\textit{Hadoop}} distributed
+computing system, which has found extensive use throughout academia and
+industry.
+For example, major software companies such as Yahoo, Microsoft, Facebook, Cloudera, etc.
+all utilize Hadoop heavily.
+Hadoop provides a mechanism (called a \textit{\textbf{distributed filesystem}}) that
+allows each computer (i.e. \textit{\textbf{node}}) in a cluster to access the same set of files.
+However, it fails to utilize local Random Access Memory (RAM) resources to reduce the waiting
+time (i.e.
+\textbf{\textit{latency}}) associated with obtaining these files.
+\textbf{This project develops and tests scheduling and caching mechanisms to reduce latency in Hadoop, thereby greatly increasing Hadoop's efficiency and applicability in fields ranging from bioinformatics to artificial intelligence.}
+
+	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
+      %\textbf{Distributed computing} has seen tremendous growth in the past few years. Essentially, it allows 
+      %In particular, Hadoop is an open source distributed computing system
+      %used throughout academia and industry. For example, Yahoo, Facebook, Microsoft, etc.
+      %contribute to its development. Hadoop Essentially, it allows computational algorithms to execute across a \textit{cluster}
+      %of computers. In order to facilitate this, distributed computational systems must make files accessible all all the \textit{nodes} of a computer cluster. 
+      %\textbf{Distributed computing} has seen tremendous growth in the past few years. 
+      %Essentially, \textbf{it allows computational algorithms to execute across multiple
+      %computers (i.e.
+\textit{nodes})}.
+Most distributed systems must make files accessible over 	networks of computers.
 \textbf{\textit{Hadoop}} is an open source and heavily used distributed computing sytem.
@@
 \textbf{This project develops and tests scheduling and caching mechanisms to reduce waiting time (i.e.
-latency) in Hadoop, thereby greatly increasing Hadoop's efficiency and applicability in fields ranging from medical diagnosis to artificial intelligence.}
+latency) in Hadoop, thereby greatly increasing Hadoop's efficiency and applicability in fields ranging from bioinformatics to artificial intelligence.}
+      %%%%%%%%%%%%%%%%%%%%%%%%%%%55
+
+      %\textbf{Distributed computing} has seen tremendous growth in the past few years.
+Essentially, \textbf{it allows computational algorithms to execute across multiple computers (i.e.
+\textit{nodes})}.
+Most distributed computing systems must make files accessible over networks of computers so that a given algorithm being executed has access to its input data.
+However, obtaining these files has an associated waiting time, known as $\textit{\textbf{latency}}$.
+\textbf{This project aims to reduce latency within
+      %distributed computing systems to reduce the time algorithms take to run}.
+
+      %\textbf{This 
+      %project aims to reduce the waiting time (known as \textit{latency}) associated
+      %with obtaining files within distributed computing systems}.
+      %\end{block}
+      %\begin{block}{Introduction}
+      %\textbf{\textit{Hadoop}} is an open source and heavily used distributed computing sytem.
+It has found extensive applications in academia and industry.
+For example, Yahoo, Facebook, Microsoft, etc.
+are all heavily involved with Hadoop development.
+But, Hadoop does not effectively utilize Random Access Memory (RAM) and local storage to better performance.
+\textbf{This project develops and tests scheduling and caching mechanisms to reduce latency in Hadoop, thereby greatly increasing Hadoop's efficiency and applicability in fields ranging from bioinformatics to artificial intelligence.}
       \end{block}
 
-	\begin{block}{Caching Theory}
+	\begin{block}{Past Research}
+	\begin{itemize}
+		\item Several sources (e.g. T. White) have identified latency as a core problem in Hadoop's architecture. However, cache-based latency reduction has been unexplored.
+		\item Schedulers have been developed for Hadoop (by M. Zaharia, Y. He, etc.), though none have utilized file storage information.
+		\item A wide range of algorithms for operating system-level scheduling exist, but Hadoop's architecture requires a fairly ad-hoc approach.
+	\end{itemize}
+	\end{block}
+
+	\begin{block}{Caching}
 	There is significant waiting time/latency associated with retrieving files from distributed filesystems. \textbf{Caches} are used to store
@@
 	 \item The solution should be a \textbf{``drop-in'' addition}, allowing for easy adoption and integration into existing systems.
-	 \item The layer \textbf{should reduce file access latency}/waiting time significantly in comparison to \textbf{the control group: a standard Hadoop installation}.
+	 \item The layer \textbf{should reduce file access latency} significantly in comparison to \textbf{the control group: a standard Hadoop installation}.
 	 \item \textbf{Develop a scheduler} for the Hadoop which \textbf{schedules jobs based on what is available in the cache}.
@@
 		\item \textbf{Memory latency: } Reading from RAM memory has an extremely small
-		waiting time associated with it ($<$ 1/100th that of the hard drive). \textbf{Panthera converts hard drive and network latency to memory latency, thereby greatly reducing file access time.}
+		waiting time associated with it ($<$ 1/100th that of the hard drive). \textbf{Panthera converts hard drive and network latency to memory latency, and thus greatly reduces file access time.}
 		\end{itemize}
@@
 
-		\begin{block}{Panthera Scheduling Algorithm Theory}
-		\begin{itemize}
-			\item \textbf{Problem}: Given $\mathbf{J} = J_1, J_2, ... ,J_k$ jobs to be run
-			on a Hadoop cluster with caches $\mathbf{C} = C_1, C_2, ..., C_n$, what ordering will best take advantage of the caches?
-			\item \textbf{Panthera job score algorithm}: Each $J_i$ accesses a set of blocks $F_i$
-			and each $C_i$ holds a set of blocks $B_i$.
-			Every $J_i$ is assigned a score $x_i$:
-			\begin{equation}
-				x_i = \displaystyle{\frac{\left\vert{F_i \cap B_t}\right\vert}{ \left\vert{B_t}\right\vert}} \; \textrm{where} \;
-				B_t = \bigcup_{j = 1}^{n} B_j
-			\end{equation}
-			\item \textbf{Panthera scheduling algorithm}:
-			Sort $\mathbf{J}$ in descending order with respect to the $x_i$ associated
-			with each $J_i$. While $J_i$ is running, Panthera data for $J_{i+1}$ to further reduce latency.
-		\end{itemize}
-		\end{block}
-
       %\begin{block}{Panthera Data Architecture}
@@
 	\begin{itemize}
-		\item \textbf{Panthera is implemented in Google's Go programming language} which has excellent concurrency primitives and memory management, making it perfect for
-		for software like \textbf{Panthera} that must run reliably with hundreds of clients.
+		\item \textbf{Panthera is implemented in Google's Go programming language} which has excellent concurrency primitives and memory management, making it perfect for software that must run reliably with hundreds of clients.
 	\end{itemize}
@@
 	\begin{block}{Latency Testing Methodology}
-	Client and server are on separate machiner for latency testing.
+	Client and server are on separate machines for latency testing.
 	\begin{itemize}
 		\item \textbf{Metadata}: Repeatedly request directory listing for a directory with 100 files in it.
-		\item \textbf{Data}: Repeatedly request file contents for a 64MB large file (64MB = 1 block in Hadoop).
+		\item \textbf{Data}: Repeatedly request file contents for a 64MB file (64MB = 1 block in Hadoop).
 	\end{itemize}
@@
 	\begin{block}{Procedure Testing Methodology}
-	Panthera was also tested with running times of algorithms running on Hadoop (as opposed to only downloading data or metadata). The input sample to both algorithms tested was a repeated copy of "The 
+	Panthera was also tested with running times of algorithms executing on Hadoop (as opposed to only downloading data or metadata).
+The input sample to both algorithms tested was a repeated copy of "The
 	Adventures of Sherlock Holmes" (192 MB). Algorithms tested: 
 	\begin{itemize}
-	\item \textbf{WordCount}: The classic distributed algorithm that counts occurrences
-	of all words in a given text.
+	\item \textbf{WordCount}: A classic distributed algorithm that counts occurrences of all words in a given text.
 	\item \textbf{ReferenceCount}: Counts occurrences of words in a given text, but 
-	only considers the words that are in a given reference file (i.e. dictionary).
+	only considers the words that are in a given reference file.
 	\end{itemize}
+		\vspace{1.5em}
+		\centerline{\includegraphics[scale=1]{assets/v2/refgetter_boxplot.pdf}}
 		\end{block}
@@
       %##########################################################
-
             \begin{column}{0.32\textwidth}
-		\begin{block}{Procedure Results}
-		\vspace{1em}
-		\includegraphics[scale=1]{assets/v2/refgetter_boxplot.pdf}
+      		\begin{block}{Panthera Scheduling Algorithm Theory}
+		\begin{itemize}
+			\item \textbf{Problem}: Given $\mathbf{J} = J_1, J_2, ... ,J_k$ jobs to be run
+			on a Hadoop cluster with caches $\mathbf{C} = C_1, C_2, ..., C_n$, what ordering will best take advantage of the caches?
+			\item \textbf{Panthera job score}: Each $J_i$ accesses a set of blocks $F_i$
+			and each $C_i$ holds a set of blocks $B_i$.
+			Every $J_i$ is assigned a job score $x_i$:
+			\begin{equation}
+				x_i = \displaystyle{\frac{\left\vert{F_i \cap B_t}\right\vert}{ \left\vert{B_t}\right\vert}} \; \textrm{where} \;
+				B_t = \bigcup_{j = 1}^{n} B_j
+			\end{equation}
+			\item \textbf{Panthera scheduling algorithm}:
+			Sort $\mathbf{J}$ in descending order with respect to the $x_i$ associated
+			with each $J_i$. While $J_i$ is running, Panthera may prefetch data for $J_{i+1}$ to further reduce latency.
+		\end{itemize}
 		\end{block}
@@
 		\midrule
-		Metadata   & 862.6\% \\
-		Data       & 863.4\% \\
-		Word count & \phantom025.1\% \\
-		Ref. count & 331.6\% \\
+		Metadata   & 8.13x \\
+		Data       & 8.63x \\
+		Word count & 1.25x \\
+		Ref. count & 3.31x \\
 	\bottomrule
 	\end{tabular}
-	\caption{Improvement with Panthera in the total amount of time required to execute algorithms/requirements (compared with status quo)}
+	\caption{Improvement with Panthera in the total amount of time required to execute algorithms/requirements (compared with status quo).
+\textbf{An improvement of "3x" implies that the mean time was 3 times lower with Panthera compared to the status quo.}}
 	\end{table}
@@
 	\caption{Reduction obtained with Panthera in the standard deviation of running times, i.e. a \textbf{contraction of 9x implies that the standard deviation was reduced
-	by a factor of 9 with Panthera (compared to the status quo).}}
+	by a factor of 9 with Panthera compared to the status quo.}}
 	\end{table}
@@
     		Hadoop jobs.
-    		\item Metadata latency was \textbf{decreased by a factor of 8.626}
-    		\item Data latency was \textbf{decreased by a factor of 8.634}
+    		\item Metadata latency was \textbf{decreased by a factor of 8.13}
+    		\item Data latency was \textbf{decreased by a factor of 8.63}
     		\item The running of time of WordCount (an algorithm with very few file accesses) was \textbf{improved by 25.1\%} with the code left unmodified.
-		\item The running time of RefCount was \textbf{improved by 331.6\%} (equivalently, decreased by a factor of 3.316).
+		\item The running time of RefCount was \textbf{improved by a factor of 3.31} with its code also unmodified.
     	\end{itemize}
@@
     without any code modifications and also enable developers to 
-    implement algorithms that were previously impractical.
+    implement algorithms that were previously impractical (such as RefCount).
 
@@
 	\begin{itemize}
-		\item Machine Learning
+		\item \textbf{Machine Learning}
 			\begin{itemize}
-			\item Panthera was tested with Apache Mahout (a library of machine learning algorithms for Hadoop) and overall there was a 15\% reduction in total running time of
-			the k-means algorithm with two clusters.
+			\item Panthera was tested with Apache Mahout, which is a library of machine learning algorithms for Hadoop
 			\end{itemize}
 
-		\item Bioinformatics
+		\item \textbf{Bioinformatics}
 			\begin{itemize}
@@
 
-		\item Predictive Analytics	
+		\item \textbf{Predictive Analytics}	
 			\begin{itemize}
@@
 			\end{itemize}
-		\item Image Processing
-		\item Numerical/Scientific Computing
+		\item \textbf{Image Processing}
+		\item \textbf{Numerical/Scientific Computing}
 		\begin{itemize}
@@
 			\end{itemize}
+
 	\end{itemize}
+	Due to the drastic decreases in latency and running times that Panthera provides
+	(in addition to allowing the implementation of previously impractical algorithms), \textbf{Panthera can truly transform computational work across many
+	fields of study.}
 	\end{block}
@@
 	\begin{itemize}
-		\item \textbf{All code related to this project will be open sourced} to further development in distributed computing.
+		\item \textbf{All code developed during this project will be open sourced to further development in distributed computing.}
+		\item Expose an API to allow external caching requests (almost complete)
 		\item Currently working on testing a \textbf{bioinformatics toolkit with Panthera}; initial results are very promising and show significant reduction in total running time.
-		\item Smart scheduling using Panthera; arranges tasks so that the cache is used most efficiently.
+		\item Implement \textbf{predictive prefetching} algorithms on Panthera
+		\item Consider the effects of different cache replacement algorithms
 

### 2014-05-10 paper/paper.tex
poster lay

--- 2014-03-01 paper/paper.tex
@@
 
-% The geometry package allows for easy page formatting.
-\usepackage{geometry}
-\geometry{letterpaper}
+% The geometry package al	lows for easy page formatting.
+%\usepackage{geometry}
+%\geometry{letterpaper}
 
@@
 \usepackage{doc}
+\usepackage{fullpage}
+\usepackage{booktabs}
+\usepackage{subcaption}
+\usepackage{lmodern}
 
@@
 %
-\title{Panthera: A Study of Caching in Distributed Computing}
+\title{Panthera: Caching and Cache-based
+Scheduling in Distributed Computing Systems}
 \author{Dhaivat Pandya \\ Appleton North High School}
@@
 	The adoption of distributed computing systems has grown massively in the past few years.
-In particular, Apache Hadoop, which allows developers to create applications that run on a cluster of computers, is currently used throughout academia and industry.
-However, the Hadoop File System fails to effectively utilize memory and local storage in order to reduce waiting time (i.e.
+In particular, Apache Hadoop, which allows developers to create applications that run on a cluster of computers, is currently used throughout academia and industry in areas such as machine learning, medical diagnosis, natural language processing, etc.
+However, the Hadoop File System fails to effectively utilize random access memory (RAM) and local storage in order to reduce waiting time (i.e.
 latency).
-In this project, a system, named Panthera, was developed, which creates a caching system for Hadoop metadata and data.
-Panthera caches both information about files and the file contents, thereby reducing waiting time in downloading and accessing the files and related information.
-It runs with an unmodified version of Hadoop, meaning that it can integrate easily into existing architecture.
-Panthera was tested using both single node and multi-
-node setups.
-The results showed an 90.8\% decrease in waiting time with the Panthera metadata cache, and a 95.7\% decrease with the Panthera data cache.
-Such drastic decreases in latency can greatly increase the efficiency of existing Hadoop applications and also allow the creation of algorithms that were previously not feasible.
-Panthera has numerous applications in fields ranging from market research to bioinformatics.
-All code associated with the project will be open sourced to further development in the area.
+In this project, named \textit{Panthera}, caching and scheduling systems for Hadoop were developed.
+\textit{Panthera} caches both information about files and the file contents, thereby reducing waiting time in downloading and accessing the files and related information.
+It runs with an unmodified version of Hadoop, meaning it can integrate easily into existing architecture.
+The results showed that data access latency was 8.63 times lower with \textit{Panthera} and metadata access latency was 8.13 times lower.
+Such drastic decreases in latency can greatly increase the efficiency of existing Hadoop applications and also allow the creation of algorithms that were previously not feasible on Hadoop.
+\textit{Panthera} was tested with existing Hadoop algorithms and running times were up to 3.31 times lower than the control group of a standard Hadoop installation.
+In addition to the caching system, a scheduler and scheduling algorithm for Hadoop were developed.
+They use information available about the caches to decide the order of execution for computational jobs, thereby further reducing running time.
+Finally, \textit{Panthera} is widely applicable and can significantly speed up research in a very large range of fields ranging from bioinformatics to artificial intelligence.
+All code developed during this project will be open sourced to further development in the area.
 }
@@
 \section{Introduction}
-Distributed computing systems are currently employed across academia and industry, with applications varying from market research for department stores to bioinformatics.
+As computational problems and their associated datasets have grown in size, it has become necessary to take advantage of distributed computing systems to solve them \cite{taylor2010overview}.
+Such systems allow programs to scale from one computer (i.e.
+one \textit{node}) to thousands of computers with few or no changes in the codebase.
 In particular, the Hadoop distributed system \cite{hadoop} has seen tremendous growth.
-Hadoop is an open source version of the revolutionary MapReduce system developed at Google.
+Hadoop is an open source version of the revolutionary MapReduce system developed at Google \cite{mapreduce}.
 With it, developers can easily take advantage of large, multi-node clusters to solve computational problems.
@@
 Current Hadoop deployments exist in the thousands of nodes at companies such as Cloudera, Facebook, Yahoo, etc.
-However, especially in larger networks, file access latency can be a significant issue.
-The Hadoop File System currently does not use local (i.e.
+Especially in such large networks, \textit{access latency}, or the waiting time associated with accessing files through the Hadoop Distributed File System (\textit{HDFS}), is a significant concern.
+Latency in distributed systems can have significant effects and a reduction of the same comes with tremendous benefits.
+As the RAMCloud project has outlined \cite{ramcloud}, low latency can greatly extend the applications of distributed computing systems.
+For example, tree traversal algorithms are largely impractical in high latency systems such as Hadoop \cite{ramcloud}, however, with lowered latency, such algorithms may become useful and applicable.
+HDFS currently does not use local (i.e.
 on the client) random access memory (RAM) to improve latency.
 
-Latency in distributed systems can have significant effects, and a reduction of the same comes with tremendous benefits.
-As the RAMCloud project has outlined \cite{ramcloud}, low latency can greatly extend the applications of distributed computing systems.
-For example, tree traversal algorithms are largely impractical in Hadoop, however, with lower latency, such algorithms may become useful and applicable.
+In this paper, I present \textit{Panthera}, a \textit{cache layer} and \textit{cache-based scheduling system} for Hadoop.
+The \textit{Panthera} cache layer functions by retaining recently accessed files in random access memory, thus, when the files are needed once again, they do not have to be downloaded or read from the hard drive, thereby greatly reducing file access latency.
+Another portion of the project, the \textit{Panthera} scheduling system utilizes information about the data blocks cached in a cluster at a given time to schedule jobs.
+By doing so, it aims to maximize the benefit of caching to a series of jobs.
 
-In this paper, we discuss \textit{Panthera}, a cache layer for Hadoop.
+In this paper, we discuss both the \textit{Panthera} caching layer and scheduling system.
 In Section 2, we discuss the specific challenges associated with such a project.
@@
 Thus, for \textit{Panthera} to be practical, it must operate independently of the existing Hadoop codebase, i.e.
-as a layer, rather than a modification.
+as a layer, rather than modification.
 This implies that \textit{Panthera} must implemented without a single change in the existing Hadoop codebase, i.e.
 as a "drop-in" solution.
+Additionally, for non-cache related requests, \textit{Panthera} must add insignificant latency.
+Finally, data and metadata request latency should be significantly lowered with \textit{Panthera} in comparison to a standard Hadoop installation.
 
-Additionally, for non-cache related requests, \textit{Panthera} must add insignificant latency.
-Finally, data and metadata request latency should be significantly with \textit{Panthera} in comparison to a standard Hadoop installation.
+In order to develop a cache-based scheduling system, the \textit{Panthera} caching layer must be able to report information (e.g.
+blocks currently cached) over the network.
+The cache-based scheduler then uses information about blocks cached and the files each job requires to run in order to optimize the running order of a set of jobs (see Section \ref{scheduler-section} for a complete description of the \textit{Panthera} scheduling algorithm).
+
+\section{Metadata and Data}
+
+Within the Hadoop File System (HDFS), there is a clear distinction between \textit{metadata} and \textit{data} \cite{hdfs}.
+Metadata refers to information about a file or a directory, e.g.
+file size, file names, on which server the contents of a file are located, etc.
+For HDFS, the NameNode server handles all metadata.
+For every request, the NameNode recomputes metadata.
+\textit{Panthera} can reduce this computation by caching the results of a previous request, thereby conserving resources.
+
+Data refers to the actual content of a file.
+Within HDFS, each file is split into 64 MB large \textit{blocks}, which are stored on DataNodes.
+Here, \textit{Panthera} serves to reduce access time by holding recently used blocks in RAM locally thereby eliminating the need to download files from a remote server if they are present in the cache.
+
+Metadata and data caching provide significantly different technical challenges as the former involves conversion of computational latency to memory latency, whereas the latter involves a conversion of network and hard drive latency to memory latency.
+Note: memory latency is of the smallest magnitude of all three in consideration.
+In many cases \cite{ramcloud} it is 1/1000th hard drive latency.
 
@@
 \begin{figure}[!h]
-  \caption{Standard Hadoop Architecture (left), Panthera-Hadoop Architecture (right)}
 	\centering
@@
 	\includegraphics[scale=0.4]{assets/panthera_architecture.pdf}
+  \caption{Standard Hadoop Architecture (left), \textit{Panthera}-Hadoop Architecture (right)}
+  \label{arch}
+
 \end{figure}
 
-Figure 1 outlines the architecture differences between a standard Hadoop File System architecture and an installation with Panthera.
-
-Within the standard conditions, the \textit{NameNode} serves as a metadata server, handling all requests from clients relating to metadata (e.g.
+Figure~\ref{arch} outlines the architecture differences between a standard Hadoop File System architecture and an installation with \textit{Panthera}.
+In HDFS, the \textit{NameNode} serves as a metadata server, handling all requests from clients relating to metadata (e.g.
 directory listing, file sizes, block locations, etc.).
@@
 
-Panthera operates by intercepting requests and responses at the NameNode and the client.
+\textit{Panthera} operates by intercepting requests and responses at the NameNode and the client.
 Both nodes have a running instance of \textit{Panthera}.
-Every request issued from the client is first reviewed by Panthera, and if possible, answered immediately from the cache.
+Every request issued from the client is first reviewed by \textit{Panthera}, and if possible, answered immediately from the cache.
 The \textit{Panthera} instance on the NameNode constantly monitors the filesystem for changes that need to be reflected in the cache (e.g.
-if a file is deleted, the cache on the client should have a copy of it).
+if a file is deleted, the cache on the client should not have a copy of it).
 
-Panthera, since it caches both data and metadata, can be considered two separate subsystems, which are now discussed.
+\textit{Panthera}, since it caches both data and metadata, can be considered two separate subsystems, which are now discussed.
 
@@
 \begin{figure}[!h]
-	\caption{Panthera Metadata System}
+	\caption{\textit{Panthera} Metadata System}
 	\centering
@@
 
-The Panthera metadata system runs on the NameNode and communicates with the Panthera instance on the client.
+The \textit{Panthera} metadata system runs on the NameNode and communicates with the \textit{Panthera} instance on the client.
 The challenge on the metadata portion consist of reducing computational latency to memory latency.
@@
 Obviously, there are some methods for which caching is not worthwhile since their execution time is lower or comparable to memory latency.
-Within Panthera, six different Hadoop query methods were identified as medium to high latency.
-Panthera identifies packets that contain these methods and subsequently caches them.
+Within \textit{Panthera}, six different Hadoop query methods were identified as medium to high latency.
+\textit{Panthera} identifies packets that contain these methods and subsequently caches them.
 
-Currently, a standard LRU cache is used for each method. Testing of different caching methods and predictive prefetching algorithms is under way.
+Currently, a standard LRU cache \cite{cache} is used for each method. Testing of different caching methods and predictive prefetching algorithms is under way.
 
@@
 \begin{figure}[!h]
-	\caption{Panthera Data System}
+	\caption{\textit{Panthera} Data System}
 	\centering
@@
 
-Panthera's data caching system runs on the client.
+\textit{Panthera}'s data caching system runs on the client.
 The architecture is similar to that of the metadata caching system.
-Every request that the client makes to the DataNodes is first processed by Panthera and responded to from the cache if possible.
+Every request that the client makes to the DataNodes is first processed by \textit{Panthera} and responded to from the cache if possible.
 
@@
 As \cite{ramcloud, hadoop} note, memory latency can be extremely low in modern systems, especially in comparison to network and hard drive latencies.
-Thus, in some respects, the goal is "easier" to attain in the data cache.
+Thus, in some respects, the goal of lowered latency is "easier" to attain in the data cache.
 However, for smaller files, it may be difficult to obtain latency improvement.
+
+Hadoop, in professional and academic environments, is often used with series rather than individual jobs.
+Deciding how to order these jobs depending on cached file availability is a problem of great interest; \textit{Panthera} presents the solution of a cache-based scheduler for Hadoop.
+
+\section{Scheduling Algorithm} \label{scheduler-section}
+
+The scheduling problem faced may be stated as:
+
+\begin{quote}
+Given $\mathbf{J} = J_1, J_2, ... ,J_k$ jobs to be run on a Hadoop cluster with caches $\mathbf{C} = C_1, C_2, ..., C_n$, what ordering will best take advantage of the caches?
+\end{quote}
+
+Each $J_i$ accesses a set of blocks $F_i$ and each $C_i$ holds a set of blocks $B_i$. \textit{Panthera} assigns every $J_i$ a \textit{job score}, denoted $x_i$:
+			\begin{equation}
+				x_i = \displaystyle{\frac{\left\vert{F_i \cap B_t}\right\vert}{ \left\vert{B_t}\right\vert}} \; \textrm{where} \;
+				B_t = \bigcup_{j = 1}^{n} B_j
+			\end{equation}
+
+			The \textit{Panthera scheduling algorithm} consists of sorting $\mathbf{J}$ in descending order with respect to the $x_i$ associated
+			with each $J_i$.
+
+		While $J_i$ is running, \textit{Panthera} may prefetch data for $J_{i+1}$ to further reduce latency.
+An implementation of predictive prefetching algorithms as described in \cite{griffioen1994reducing} is possible in \textit{Panthera} and may serve to further reduce latency in specific applications.
 
@@
 
-In testing Panthera, very simple strategies were adopted.
-In order to test the metadata cache, a directory listing query was repeatedly run on a directory with 100 files in it, and the latency times for responding are measured, with Panthera and without.
+In order to test the metadata cache, a directory listing query was repeatedly run on a directory with 100 files in it, and the latency times for responding are measured, with \textit{Panthera} and without.
 The reasoning for picking such a test lies in the fact that a directory listing in Hadoop involves all six of the cached methods as well as a few non-cached methods.
 Thus, the test provides a good estimate of the efficacy of the cache.
+Latency times for the data cache were obtained by repeatedly querying for a 64 megabyte large file.
+The file size was chosen since it is the default block size for a Hadoop cluster and the \textit{Panthera} cache stores files as blocks.
+Both metadata and data cache testing was conducted on virtualized server instances running on hardware with solid state hard drives.
 
-Latency times for the data cache were obtained by repeatedly querying for a one megabyte large file. The test encapsulates the most common use case for Panthera and the Hadoop File System.
+In order to test the effects of \textit{Panthera} on the Hadoop computing system (rather than focusing on only HDFS access time), running times of two existing algorithms were recorded.
+The first is \textit{WordCount}, the classic distributed algorithm that counts the number of occurrences of each word in a text file.
+The second has been, termed \textit{RefCount}, has thus far been impractical on Hadoop due to excessive running times.
+It is similar to WordCount in that it counts occurrences of words, but it consults a \textit{reference file} (e.g.
+a limited dictionary) to determine which words are to be counted.
 
-Tests for the metadata cache were conducted on both single node and multinode setups, whereas the datanode cache testing was run on multinode setups.
-
-\section{Results: Data caching}
-\begin{figure}[!h]
-	\caption{Panthera data cache latencies}
+\section{Results}
+\begin{figure}
 \centering
-		\includegraphics[scale=0.4]{assets/box-plot-data.pdf}
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[scale=0.3]{assets/v2/getter_boxplot.pdf}
+  \caption{Data access latency}
+  \label{data}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[scale=0.3]{assets/v2/multinode_meta_box_plot.pdf}
+  \caption{RefCount job performance}
+  \label{meta}
+\end{subfigure}
+\caption{Metadata access latency}
+\label{fig:test}
 \end{figure}
 
-The Panthera data cache showed a 95.7\% decrease in the latency in comparison to a standard deviation non-datacached installation.
-In addition, the standard of deviation was lower by a factor of 27.
-For any deployed project, maximum levels of latency must be considered when designing a system.
-Thus, the \textit{status quo} of the non-cached version severely limits possibilities within Hadoop.
+	\begin{table}
+	\centering
+			\begin{tabular}{lcc}
+			\toprule
+			\textbf{Type} & \textbf{Mean Time Improvement} & \textbf{Standard Dev. Contraction} \\
+			\midrule
+			Metadata   & 8.13x & 27.38x \\
+			Data       & 8.63x & 11.14x \\
+			WordCount & 1.25x & \phantom01.02x \\
+			Ref. count & 3.31x & 22.91x \\
+			\bottomrule
+			\end{tabular}
+		\caption{Latency improvement with \textit{Panthera}}
+		\label{results}
+	\end{table}
 
-\section{Results: Metadata caching}
-\begin{figure}[!h]
-	\caption{Panthera metadata cache latencies}
+The \textit{Panthera} data cache decreased file access latency by a factor of 8.63 and metadata access latency was lowered by a factor of 8.13, both representing drastic decreases in HDFS latency.
+In addition, the standard deviation was also significantly decreased with \textit{Panthera}: by a factor 27.38 for metadata and a factor of 11.14 for data.
+With a lower standard deviation, the predictability of performance for Hadoop applications increases.
+
+The tests with existing Hadoop algorithms show similarly positive results.
+\textit{Panthera} is able to reduce running time for the WordCount algorithm by 25\% and is able to do the same for the Ref.
+Count algorithm by 331\%.
+It is important to keep in mind that these performance boosts were attained with the original algorithms left entirely unmodified.
+The difference between the two lies in the fact that WordCount has little interaction with HDFS once the initial input files have been downloaded, whereas Ref.
+Count often obtains the reference file from HDFS.
+Notice that the testing was conducted with only one slave node.
+Adding slave nodes to a cluster would increase the number of file requests since each node requires a portion of the input file.
+
+\begin{figure}
 \centering
-		\includegraphics[scale=0.3]{assets/box-plot-metadata.pdf}
-		\includegraphics[scale=0.3]{assets/box-plot-metadata-multinode.pdf}
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[scale=0.3]{assets/v2/wordcount_boxplot.pdf}
+  \caption{WordCount job performance}
+  \label{fig:sub1}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[scale=0.3]{assets/v2/refgetter_boxplot.pdf}
+  \caption{RefCount job performance}
+  \label{refgetter}
+\end{subfigure}
+\caption{Algorithm running time comparison}
+\label{fig:test}
 \end{figure}
-
-Metadata caching has similarly incredible results.
-\textit{Panthera} showed a 9 fold or 90.8\% improvement in the total latency as measured on the client side in a multinode setup.
-Additionally, the standard deviation is 24 fold lower in \textit{Panthera}.
 
@@
 
-\textit{Panthera} fulfils the constraints mentioned at the outset.
-It is a layer on Hadoop rather than a main repository code modification.
+\textit{Panthera} fulfils all of the constraints mentioned at the outset.
+It is a layer on Hadoop rather than a main repository code modification, allowing existing Hadoop applications to remain entirely unmodified.
 Considering the latency measurements, it is clear that it has also met the requirements in terms of benefit.
-\textit{Panthera} was able to reduce data access latency by 95.7\% and reduce basic metadata access latency by 90.8\%.
-Additionally, \textit{Panthera} greatly reduced the variation in the access latency, greatly aiding in the pre-development plans and extending the capabilities of distributed computing systems such as Hadoop.
+\textit{Panthera} was able to reduce data access latency by a factor of 8.63 and reduce basic metadata access latency by 8.13.
+Also, \textit{Panthera} was able to minimize the spread of latency times in both data and metadata requests.
+Thus, \textit{Panthera} opens up great possibilities within Hadoop due to the fact that algorithms that were previously impractical (e.g.
+distributed tree traversal algorithms) may now be implemented.
+Additionally, a scheduler algorithm was devised and a cache-based scheduler was implemented for Hadoop which enables further reduction of latency via greater utilization of cached resources.
+
+Previous work \cite{hdfs, borthakur2011apache} has lamented the latency issues that have plagued HDFS and Hadoop as barriers to adoption.
+\textit{Panthera} significantly reduces the scope of these issues and allows the application of Hadoop in areas where its reach has been limited.
+
+\section{Applications}
+One of the major benefits of the systems developed is that they are independent of the existing Hadoop codebase and they do not require modifications to existing Hadoop applications.
+Thus, \textit{Panthera} may be deployed in a wide range of scenarios, such as the following:
+
+\begin{itemize}
+	\item Bioinformatics: Hadoop is currently in use in bioinformatics research \cite{taylor2010overview}.
+Panthera can be employed to reduce access latencies of various datasets, e.g.
+protein data banks, etc.
+Additionally, domain-specific prefetching can be performed to further improve performance.
+
+	\item Machine learning: Apache Mahout is a machine learning library for Hadoop.
+Panthera has been tested with the K-Means implementation it provides and there is evidence of significant performance improvement.
+
+	\item Predictive analytics: A quickly growing field that has thoroughly
+	embraced Hadoop \cite{chen2012business}.
+Metadata caching can be used to decrease human operator waiting time.
+Often, queries consist of several queued jobs, meaning that the \textit{Panthera} scheduler can be used to optimize
+	job ordering.
+\end{itemize}
 
 \section{Future work}
-\textit{Panthera} can be extremely useful in a wide variety of applications, specifically in those which require large datasets in separate files (e.g.
-unstructured data).
-In particular, I will be concentrating on bioinformatics and integrating Panthera's caching with algorithms used in the area.
+One of the most immediate avenues for further research is in \textit{domain-specific prefetching}.
+Different problem domains (e.g.
+natural language processing, predictive analytics, etc.) have different dataset requirements.
+It may be possible to predict what files a given job will need depending
+on what problem domain the job is in.
+A generalized version of the same problem would be \textit{predictive prefetching} \cite{griffioen1994reducing} which involves the prediction and prefetching of files that will be required in the future.
+In both these problems, the Hadoop environment provides unique
+challenges.
 
-I am also working on a cache-based scheduler for Hadoop.
-Essentially, it will use information about what files are available in various clients and arrange Hadoop tasks such that the use of the cache is maximized.
-
-Finally, all code associated with this paper and \textit{Panthera} will be open sourced in order to further development of Hadoop and distributed computing.
+Another area worthy of further exploration is the effect of different cache replacement algorithms in Hadoop. Though various cache replacement algorithms
+have been explored thoroughly at the local filesystem level, their study 
+in conjunction with existing Hadoop algorithms may provide novel results, i.e.
+the access patterns of certain procedures may work best with a certain
+cache replacement scheme \cite{puzak1985analysis}.
 

### 2014-05-13 src/caches/get_file_info_cache.go
open sourcing code\!

@@ -35,16 +35,21 @@ func (gfi_cache *GetFileInfoCache) Disable() {
 	gfi_cache.Cache.Enabled = false
 }
 
-//Query the cache. Returns nil if req is not found in the cache or the Enabled is set to false.
-//gfi_cache.Cache.Query should NOT be called since it suses the wrong kind of equality comparator
-func (gfi_cache *GetFileInfoCache) Query(req namenode_rpc.ReqPacket) namenode_rpc.ResponsePacket {
+//Query the cache. Returns nil if req is not found in the cache or the Enabled is set to 
+//false.
+//gfi_cache.Cache.Query should NOT be called since it suses the wrong kind of equality 
+//comparator
+func (gfi_cache *GetFileInfoCache) Query(
+	req namenode_rpc.ReqPacket) namenode_rpc.ResponsePacket {
+	
 	util.DebugLogger.Println("in GetFileInfoCache.Query()")
 	//if the cache is not enabled, we keep returning nil
 	if !gfi_cache.enabled {
 		return nil
 	}
 
-	equals := EqualityFunc(func(req1 namenode_rpc.ReqPacket, req2 namenode_rpc.ReqPacket) bool {
+	equals := EqualityFunc(func(req1 namenode_rpc.ReqPacket, req2 namenode_rpc.ReqPacket) 
+	bool {
 		//the conditions mean the directory being queried and the method called
 		//are the same
 		methodName1 := string(req1.GetMethodName())

### 2014-05-13 src/caches/get_listing_cache.go

@@ -33,7 +33,6 @@ func (glc *GetListingCache) Disable() {
 }
 
 
-//TODO POTENTIAL BUG this method has not been unit tested
 func (glc *GetListingCache) Query(req namenode_rpc.ReqPacket) namenode_rpc.ResponsePacket {
 	//if the cache is not enabled, we keep returning nil
 	if !glc.IsEnabled() {

### 2014-05-13 src/caches/request_cache.go

@@ -35,7 +35,8 @@ type RequestCache struct {
 	//NameNode
 	RequestResponse map[PacketNumber](namenode_rpc.PacketPair)
 
-	//records the order in which packet numbers were placed in RequestResponse
+	//records the order in which packet numbers were placed in 
+	//RequestResponse
 	PacketNumbers []PacketNumber
 
 	Hits int
@@ -61,14 +62,17 @@ func (rc *RequestCache) Disable() {
 	rc.Enabled = false
 }
 
-//this a private method because it assumes that the mutex has already been locked
-func (rc *RequestCache) add(rp namenode_rpc.ReqPacket, resp namenode_rpc.ResponsePacket) error {
+//this a private method because it assumes that the mutex has already 
+//been locked
+func (rc *RequestCache) add(rp namenode_rpc.ReqPacket, 
+resp namenode_rpc.ResponsePacket) error {
 
 	//the central assumption does not hold if 
 	//we do not have equal packet numbers
 	if rp != nil && resp != nil {
 		if rp.GetPacketNumber() != resp.GetPacketNumber() {
-			return errors.New("Packet numbers are not equal for request and response")
+			return errors.New("Packet numbers are not equal for request 
+			and response")
 		}
 	}
 
@@ -101,7 +105,8 @@ func (rc *RequestCache) add(rp namenode_rpc.ReqPacket, resp namenode_rpc.Respons
 	return nil
 }
 
-func (rc *RequestCache) Add(rp namenode_rpc.ReqPacket, resp namenode_rpc.ResponsePacket) error {
+func (rc *RequestCache) Add(rp namenode_rpc.ReqPacket, 
+	resp namenode_rpc.ResponsePacket) error {
 	rc.Lock()
 	//unlock the mutex after done w/ processing this function
 	defer rc.Unlock()
@@ -109,7 +114,8 @@ func (rc *RequestCache) Add(rp namenode_rpc.ReqPacket, resp namenode_rpc.Respons
 	return rc.add(rp, resp)
 }
 
-//same as Add() but adds a PacketPair with a nil in place of the response packet
+//same as Add() but adds a PacketPair with a nil in place of the response 
+//packet
 func (rc *RequestCache) AddRequest(req namenode_rpc.ReqPacket) error {
[…]

### 2014-05-13 src/caches/request_cache_test.go

@@ -14,7 +14,6 @@ func TestRequestCacheConstructor(t *testing.T) {
 	}
 }
 
-//TODO revise structure of this unit test
 func TestRequestCacheAdd(t *testing.T) {
 	rs := NewRequestCache(2)
 
@@ -80,7 +79,6 @@ func TestRequestCacheQuery(t *testing.T) {
 		t.Fail()
 	}
 
-	//TODO this dereferencing going on here is too complicated
 	if !reflect.DeepEqual(resp, rc.Query(rp)) {
 		fmt.Println("Failed query result: ", rc.Query(rp))
 		fmt.Println("First response packet: ", rc.RequestResponse[PacketNumber(resp.GetPacketNumber())].Response)

### 2014-05-13 src/configuration/load_test.go

@@ -22,7 +22,8 @@ func TestLoadFile (t *testing.T) {
 
 	conf, err := LoadFile("example_conf.json")
 	if err != nil {
-		fmt.Println("Encountered error in loading configuration file: ", err)
+		fmt.Println("Encountered error in loading configuration file: ", 
+		err)
 		t.FailNow()
 	}
 	fmt.Println("Example conf: ", exampleConf)

### 2014-05-13 src/configuration/locations.go

@@ -45,7 +45,8 @@ type DataNodeMap map[Port]*DataNodeLocation
 func MakeDataNodeMap(dnls []*DataNodeLocation, portOffset int) DataNodeMap {
 	res := make(DataNodeMap)
 	for i := 0; i < len(dnls); i++ {
-		fmt.Println("For port: ", strconv.Itoa(portOffset+i), ", dnls[i]: ", dnls[i])
+		fmt.Println("For port: ", strconv.Itoa(portOffset+i), 
+		", dnls[i]: ", dnls[i])
 		res[Port(strconv.Itoa(portOffset+i))] = dnls[i]
 	}
 

### 2014-05-13 src/configuration/locations_test.go

@@ -44,12 +44,14 @@ func TestDataNodeMapConstructor (t *testing.T) {
 	dnLocationMap := MakeDataNodeMap(dnLocations, 2000)
 
 	if !reflect.DeepEqual(dnLocationMap[Port("2000")], dnLocations[0]) {
-		fmt.Println("Failed, not equal: ", dnLocationMap["2000"], dnLocations[0])
+		fmt.Println("Failed, not equal: ", dnLocationMap["2000"], 
+		dnLocations[0])
 		t.Fail()
 	}
 
 	if !reflect.DeepEqual(dnLocationMap[Port("2000")], dnl) {
-		fmt.Println("Failed second, not equal:", dnLocationMap["2000"], dnl)
+		fmt.Println("Failed second, not equal:", dnLocationMap["2000"], 
+		dnl)
 		t.Fail()
 	}
 

### 2014-05-13 src/data_requests/data_req_processor.go

@@ -88,7 +88,8 @@ type Processor struct {
 	startTimeCached time.Time
 }
 
-func NewProcessor(dataCache *caches.DataCache, nodeLocation *configuration.DataNodeLocation) *Processor {
+func NewProcessor(dataCache *caches.DataCache, 
+nodeLocation *configuration.DataNodeLocation) *Processor {
 	p := Processor{
 		dataCache: dataCache,
 		retryDataNode: true,
@@ -155,14 +156,17 @@ func (p *Processor) setCurrentRequest(dataRequest datanode_rpc.ReqPacket) {
 	util.DataReqLogger.Println(p.id, " Set current request.")
 }
 
-func (p *Processor) queryCache(dataRequest *datanode_rpc.DataRequest) *datanode_rpc.DataResponse {
+func (p *Processor) queryCache(
+	dataRequest *datanode_rpc.DataRequest) *datanode_rpc.DataResponse {
 	resp := p.dataCache.Query(*dataRequest)
-	util.DataReqLogger.Println(p.id, " Returned from cache request: ", resp)
+	util.DataReqLogger.Println(p.id, " Returned from cache request: ", 
+	resp)
 	return resp
 }
 
 //handle the request if it is an instance of datanode_rpc.DataRequest
-func (p *Processor) handleDataRequest(conn net.Conn, dataNode net.Conn, dataRequest *datanode_rpc.DataRequest) error {
+func (p *Processor) handleDataRequest(conn net.Conn, dataNode net.Conn, 
+	dataRequest *datanode_rpc.DataRequest) error {
 	if dataRequest == nil {
 		return errors.New("Data request value is nil")
 	}
@@ -179,9 +183,12 @@ func (p *Processor) handleDataRequest(conn net.Conn, dataNode net.Conn, dataRequ
 		p.RecordCachedLatency();
 		p.skipResponse = true
 	} else {
-		util.DataReqLogger.Println(p.id, " Cache miss. (cache size: ", p.dataCache.CurrSize(), ")")
-		util.DataReqLogger.Println(p.id, " Cache contents: ", p.dataCache.CachedRequests())
-		util.DataReqLogger.Println(p.id, " Cached responses: ", p.dataCache.CachedResponses())
+		util.DataReqLogger.Println(p.id, " Cache miss. 
+		(cache size: ", p.dataCache.CurrSize(), ")")
+		util.DataReqLogger.Println(p.id, " Cache contents: ", 
+		p.dataCache.CachedRequests())
[…]

### 2014-05-13 src/data_requests/put_req_processor.go

@@ -21,7 +21,8 @@ as a relay between Hadoop and the DataNode.
 //onto the DataNode. This method is needed because 
 //once the request is read, it still has to be forwarded
 //to the DataNode
-func ForwardPutDataRequest(request *datanode_rpc.PutDataRequest, conn net.Conn, dataNode net.Conn) {
+func ForwardPutDataRequest(request *datanode_rpc.PutDataRequest, 
+conn net.Conn, dataNode net.Conn) {
 	dataBytes, err := request.Bytes()
 	if err != nil {
 		util.DebugLogger.Println("Could not ForwardPutDataRequest(): ", err)
@@ -51,7 +52,8 @@ func (p *PutRequestProcessor) HandleConnection(conn net.Conn, dataNode net.Conn)
 		bytesRead, err := conn.Read(buf)
 		if err != nil {
 			//we can't really do anything if there's an error: just gotta keep going
-			util.DebugLogger.Println("Error occurred while reading in PutRequestProcessor.HandleConnection: ", err)
+			util.DebugLogger.Println("Error occurred while reading in 
+			PutRequestProcessor.HandleConnection: ", err)
 			continue
 		}
 
@@ -59,7 +61,8 @@ func (p *PutRequestProcessor) HandleConnection(conn net.Conn, dataNode net.Conn)
 		buf = buf[0:bytesRead]
 		_, err = dataNode.Write(buf)
 		if err != nil {
-			util.DebugLogger.Println("Errror occurred while writing in PutRequestProcessor.HandleConnection: ", err)
+			util.DebugLogger.Println("Errror occurred while writing in 
+			PutRequestProcessor.HandleConnection: ", err)
 		}
 	}
 }
@@ -71,14 +74,16 @@ func (p *PutRequestProcessor) HandleHDFS(conn net.Conn, dataNode net.Conn) {
 		bytesRead, err := dataNode.Read(buf)
 
 		if err != nil {
-			util.DebugLogger.Println("Error occurred while reading in PutRequestProcessor.HandleHDFS: ", err)
+			util.DebugLogger.Println("Error occurred while reading in
+			 PutRequestProcessor.HandleHDFS: ", err)
 			continue
 		}
 
 		buf = buf[0:bytesRead]
 		_, err = conn.Write(buf)
 		if err != nil {
-			util.DebugLogger.Println("Error occurred while writing in PutRequestProcessor.HandleHDFS", err)
[…]

### 2014-05-13 src/datanode_rpc/datanode_rpc.go

@@ -18,11 +18,7 @@ type ReqPacket interface {
 	Equals(ReqPacket) bool
 }
 
-//TODO POTENTIAL BUG:
-//this packet structure has been
-//obtained from the Wireshark
-//specification; have to double
-//check the validity of the structure
+
 type DataRequest struct {
 	ProtocolVersion uint16
 	Command uint8
@@ -228,7 +224,6 @@ func NewDataRequest() *DataRequest {
 	return &dr
 }
 
-//TODO this code is incredibly bashy
 //Compares two dataRequests. Note that the 
 //ClientId and ClientIdLength are not compared
 //since those are largely irrelevant.
@@ -295,8 +290,7 @@ func (dr *DataRequest) Equals(p ReqPacket) bool {
 	return false
 }
 
-//TODO error checking in this function is just nonsensical; find some
-//other way to do it.
+
 func (dr *DataRequest) Bytes() ([]byte, error) {
 	byte_buffer := new(bytes.Buffer)
 
@@ -349,7 +343,8 @@ func (dr *DataRequest) Bytes() ([]byte, error) {
 		return []byte{}, err
 	}
 
-	err = binary.Write(byte_buffer, binary.BigEndian, dr.AccessPasswordLength)
+	err = binary.Write(byte_buffer, binary.BigEndian, 
+	dr.AccessPasswordLength)
 	if err != nil {
 		return []byte{}, err
 	}
@@ -411,7 +406,8 @@ func LiveReadInitial(byte_buffer io.Reader) (*InitialRead, error) {
 //called to decide what kind of packet to load
 //Note: prevPDR is a boolean value which tells us whether or not the 
 //the previous request was a PutDataRequest.
-func LoadRequestPacket(byteBuffer io.Reader, prevPDR bool) (ReqPacket, error) {
+func LoadRequestPacket(byteBuffer io.Reader, prevPDR bool) (ReqPacket, 
+error) {
 	if prevPDR {
 		putFileDataRequest := NewPutFileDataRequest()
 		putFileDataRequest.LiveLoad(byteBuffer)
@@ -421,7 +417,8 @@ func LoadRequestPacket(byteBuffer io.Reader, prevPDR bool) (ReqPacket, error) {
 	util.DebugLogger.Println("Starting LiveReadInitial()...")
 	initialRead, err := LiveReadInitial(byteBuffer)
 	util.DebugLogger.Println("InitialRead structure: ", initialRead)
[…]

### 2014-05-13 src/datanode_rpc/datanode_rpc_test.go

@@ -123,8 +123,10 @@ func TestDataResponseLoad(t *testing.T) {
 	dResp.Load(DataResponseTestCase)
 
 	if !reflect.DeepEqual(*dResp, DataResponseTestExpected) {
-		fmt.Println("Not equal, expected: ", DataResponseTestExpected, " got: ", *dResp)
-		fmt.Println("Expected data: ", string(DataResponseTestExpected.Data), "got bytes: ", string(dResp.Data))
+		fmt.Println("Not equal, expected: ", DataResponseTestExpected, 
+		" got: ", *dResp)
+		fmt.Println("Expected data: ", string(DataResponseTestExpected.Data), 
+		"got bytes: ", string(dResp.Data))
 		t.Fail()
 	}
 }
@@ -135,8 +137,10 @@ func TestDataResponseLoadTwo(t *testing.T) {
 	dResp.Load(DataResponseTestCase2)
 
 	if !reflect.DeepEqual(*dResp, DataResponseTestExpected2) {
-		fmt.Println("Not equal, expected: ", DataResponseTestExpected2, " got: ", *dResp)
-		fmt.Println("Expected data: ", string(DataResponseTestExpected2.Data), "got bytes: ", string(dResp.Data))
+		fmt.Println("Not equal, expected: ", DataResponseTestExpected2, 
+		" got: ", *dResp)
+		fmt.Println("Expected data: ", string(DataResponseTestExpected2.Data), 
+		"got bytes: ", string(dResp.Data))
 		t.Fail()		
 	}
 }

### 2014-05-13 src/hdfs_requests/processor.go

@@ -126,7 +126,6 @@ func NewProcessor(event_chan chan ProcessorEvent, cacheSet *caches.CacheSet,
 //this gets called by both HandleConnection in order to put a 
 //Request into one of the caches. Once a corresponding call is made
 //to CacheResponse(), the cache spot is "filled"
-//TODO need to implement
 func (p *Processor) CacheRequest(req *namenode_rpc.RequestPacket) {
 	//fmt.Println("Method name from CacheRequest(): ", string(req.MethodName))
 
@@ -147,13 +146,10 @@ func (p *Processor) CacheRequest(req *namenode_rpc.RequestPacket) {
 //one of the caches. It looks up a request with the same packet number
 //and if such a request exists in the cache, it places itself into that
 //spot.
-//TODO need to implement
 func (p *Processor) CacheResponse(resp namenode_rpc.ResponsePacket) {
 	packetNum := caches.PacketNumber(resp.GetPacketNumber())
 
 	//we can hook up a response with a request
-	//TODO figure out a better way to make a generic
-	//caching mechanism
 	if p.cacheSet.GfiCache.Cache.HasPacketNumber(packetNum) {
 		p.cacheSet.GfiCache.Cache.AddResponse(resp)
 	}
@@ -185,10 +181,9 @@ func (p *Processor) readLength(conn net.Conn) (uint32, error) {
 }
 
 
-//takes a request object, and if it is DatanodeRegistration related, it modifies the storageID
+//takes a request object, and if it is DatanodeRegistration related, it 
+//modifies the storageID
 //and port number (all operations done in place on req)
-//TODO this needs a fix to consider the case where the new port string length != old port string
-//length
 func (p *Processor) ModifyBlockReport(req *namenode_rpc.RequestPacket) []byte {
 	//fmt.Println("Modifying block report...")
 	//we need to change the port numbers on the blockReport calls
@@ -208,7 +203,8 @@ func (p *Processor) ModifyBlockReport(req *namenode_rpc.RequestPacket) []byte {
 	
 	storageParameter := req.Parameters[1]
 	
-	location := configuration.NewDataNodeLocationAddr(string(storageParameter.Type))
[…]

### 2014-05-13 src/hdfs_requests/processor_test.go

@@ -41,7 +41,8 @@ func TestNewProcess(t *testing.T) {
 }
 
 //GFI request
-var RequestPacketTestCase []byte = []byte{0, 0, 0, 60, 0, 0, 0, 2, 0, 10, 103, 101, 116, 
+var RequestPacketTestCase []byte = []byte{0, 0, 0, 60, 0, 0, 0, 2, 0, 
+		10, 103, 101, 116, 
 		76, 105, 115, 116, 105, 110, 103, 0, 0, 0, 2, 0, 16, 106, 
 		97, 118, 97, 46, 108, 97, 110, 103, 46, 83, 116, 114, 105, 
 		110, 103, 0, 12, 47, 117, 115, 101, 114, 47, 104, 100, 117, 

### 2014-05-13 src/hdfs_requests/request_state.go

@@ -32,9 +32,11 @@ func NewRequestState() *RequestState {
 //this method is called a new request is received from 
 //the client. It resets the byteBuffer and readFirstPacket
 //fields and caches the current response.
-func (rs *RequestState) Empty(packetNumber uint32) *namenode_rpc.GenericResponsePacket {
+func (rs *RequestState) Empty(packetNumber uint32) 
+*namenode_rpc.GenericResponsePacket {
 	rs.ReadFirstPacket = false
-	genericResp := namenode_rpc.NewGenericResponsePacket(rs.ByteBuffer.Bytes(), rs.PacketNumber)
+	genericResp := namenode_rpc.NewGenericResponsePacket(
+	rs.ByteBuffer.Bytes(), rs.PacketNumber)
 	rs.ByteBuffer.Reset()
 	rs.PacketNumber = packetNumber
 

### 2014-05-13 src/main/main.go

@@ -24,7 +24,8 @@ var config *configuration.Configuration;
 var cpuprofile = flag.String("cpuprofile", "", "write cpu profile to file")
 
 //main reactor function called by main
-func loop(server net.Listener, caches *caches.CacheSet, dnMap *configuration.DataNodeMap) {
+func loop(server net.Listener, caches *caches.CacheSet, 
+dnMap *configuration.DataNodeMap) {
 	fmt.Println("looping...")
 	eventChannel := make(chan hdfs_requests.ProcessorEvent)
 	
@@ -40,8 +41,10 @@ func loop(server net.Listener, caches *caches.CacheSet, dnMap *configuration.Dat
 		util.DebugLog("Client Accepted; no errors received...");
 
 		//set up connection to HDFS
-		util.DebugLog("Connecting to HDFS host: " + string(config.HdfsHostname) + ":" + string(config.HdfsPort))
-		hdfs, hdfs_err := net.Dial("tcp", config.HdfsHostname + ":" + config.HdfsPort)
+		util.DebugLog("Connecting to HDFS host: " + 
+		string(config.HdfsHostname) + ":" + string(config.HdfsPort))
+		hdfs, hdfs_err := net.Dial("tcp", config.HdfsHostname + ":" 
+		+ config.HdfsPort)
 		if hdfs_err != nil {
 			util.LogError(hdfs_err.Error())
 			continue
@@ -74,7 +77,8 @@ func startCacheInfoServer(dataCache *caches.WritableDataCache) {
 
 //listen on a port connected to one of the datanodes
 //will be run as a goroutine
-func loopData(listener net.Listener, location *configuration.DataNodeLocation, cache *caches.DataCache) {
+func loopData(listener net.Listener, 
+	location *configuration.DataNodeLocation, cache *caches.DataCache) {
 	dataCacheSize := 15
 	dataCache := caches.NewWritableDataCache(dataCacheSize)
 
@@ -85,7 +89,8 @@ func loopData(listener net.Listener, location *configuration.DataNodeLocation, c
 		conn, err := listener.Accept()
 		util.DebugLogger.Println("Data client accepted.")
 		if err != nil {
-			util.LogError("Could not accept connection from the DataNode: " + err.Error())
+			util.LogError("Could not accept connection from the DataNode: "
+			 + err.Error())
 		}
 
 		//connection object to the datanode location
[…]

### 2014-05-13 src/namenode_rpc/namenode_rpc.go

@@ -76,7 +76,8 @@ type ResponseObject struct {
 }
 
 //this method figures out, given a request, what kind of new packet to build
-func BuildResponsePacket(buf []byte, packetNumber uint32, currRequest *RequestPacket) ResponsePacket {
+func BuildResponsePacket(buf []byte, packetNumber uint32, 
+	currRequest *RequestPacket) ResponsePacket {
 	if currRequest == nil {
 		return NewGenericResponsePacket(buf, packetNumber)
 	}
@@ -119,7 +120,8 @@ func (grp *GenericResponsePacket) GetBuf() []byte {
 	return grp.Buf
 }
 
-func NewGenericResponsePacket(buf []byte, packNum uint32) *GenericResponsePacket {
+func NewGenericResponsePacket(buf []byte, packNum uint32) 
+*GenericResponsePacket {
 	grp := GenericResponsePacket{Buf: buf, 
 		PacketNumber: packNum}
 	return &grp
@@ -267,14 +269,16 @@ func (rp *RequestPacket) Load(buf []byte) error {
 
 		rp.Parameters[i] = *(NewParameter())
 		//now read in all the fields one by one
-		binary.Read(byte_buffer, binary.BigEndian, &(rp.Parameters[i].TypeLength))
+		binary.Read(byte_buffer, binary.BigEndian, &(
+		rp.Parameters[i].TypeLength))
 
 		//create space for and read in the type of this parameter
 		rp.Parameters[i].Type = make([]byte, rp.Parameters[i].TypeLength)
 		byte_buffer.Read(rp.Parameters[i].Type)
 
 
-		binary.Read(byte_buffer, binary.BigEndian, &(rp.Parameters[i].ValueLength))
+		binary.Read(byte_buffer, binary.BigEndian, &(
+		rp.Parameters[i].ValueLength))
 
 		//create space for and read in the value of this parameter
 		rp.Parameters[i].Value = make([]byte, rp.Parameters[i].ValueLength)
@@ -402,7 +406,6 @@ type GetFileInfoResponse struct {
 	ObjectNameLength uint16
 	ObjectName []byte
 
-	//TODO not exactly sure why there are two of these...
 	ObjectNameLength2 uint16
 	ObjectName2 []byte
 
@@ -419,7 +422,6 @@ type GetFileInfoResponse struct {
 	ModifiedTime uint64
 	AccessTime uint64
 
-	//TODO not exactly sure what the two 
 	//file permission headers specify
 	FilePermission2 uint16
 
[…]

### 2014-05-13 src/namenode_rpc/namenode_rpc_test.go

@@ -143,15 +143,11 @@ func TestRequestPacketLoad(t *testing.T) {
 		t.Fail()
 	}
 
-	//TODO this part makes absolutely no sense, but 
-	//it is how Wireshark seems to be reporting it
-	//TODO does the [B have a special meaning attached to it?
 	//test second parameter for type and value
 	if p2.TypeLength != 2 {
 		t.Fail()
 	}
 
-	//TODO WTF does this mean
 	if string(p2.Type) !=  "[B" {
 		t.Fail()
 	}
@@ -169,13 +165,20 @@ func TestNewGetFileInfoResponse(t *testing.T) {
 }
 
 //this is the test case, but in hexademical because converting is hard
-var GetFileInfoResponseTestCase []byte = []byte{0,0,0,1,0,0,0,0,0,46,111,114,103,
-	46,97,112,97,99,104,101,46,104,97,100,111,111,112,46,104,100,102,115,46,112,114,
-	111,116,111,99,111,108,46,72,100,102,115,70,105,108,101,83,116,97,116,117,115,0,
-	46,111,114,103,46,97,112,97,99,104,101,46,104,97,100,111,111,112,46,104,100,102,
-	115,46,112,114,111,116,111,99,111,108,46,72,100,102,115,70,105,108,101,83,116,97,
-	116,117,115,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,1,65,230,55,105,95,
-	0,0,0,0,0,0,0,0,1,237,6,104,100,117,115,101,114,10,115,117,112,101,114,103,114,111,
+var GetFileInfoResponseTestCase []byte = []byte{0,0,0,1,0,0,0,0,0,46,111,
+114,103,
+	46,97,112,97,99,104,101,46,104,97,100,111,111,112,46,104,100,102,115,46,
+	112,114,
+	111,116,111,99,111,108,46,72,100,102,115,70,105,108,101,83,116,97,116,
+	117,115,0,
+	46,111,114,103,46,97,112,97,99,104,101,46,104,97,100,111,111,112,46,104,
+	100,102,
+	115,46,112,114,111,116,111,99,111,108,46,72,100,102,115,70,105,108,101,
+	83,116,97,
+	116,117,115,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,1,65,230,
+	55,105,95,
[…]

### 2014-05-13 src/network_tests/test.go

@@ -21,7 +21,8 @@ func main() {
 
 	bytesRead, err := conn.Read(byteBuffer)
 	if bytesRead < 0 || err != nil {
-		fmt.Println("Error occurred in reading from the server: ", err.Error())
+		fmt.Println("Error occurred in reading from the server: ", 
+		err.Error())
 		return
 	}
 

### 2014-05-13 src/proxy/proxy.go

@@ -11,7 +11,8 @@ import (
 )
  
 var localAddr *string = flag.String("l", "0.0.0.0:1035", "local address")
-var remoteAddr *string = flag.String("r", "188.226.198.184:1102", "remote address")
+var remoteAddr *string = flag.String("r", "188.226.198.184:1102", 
+"remote address")
  
 func proxyConn(conn *net.TCPConn) {
 	rAddr, err := net.ResolveTCPAddr("tcp", *remoteAddr)

### 2014-05-13 src/testing_client/testing_client.go

@@ -24,7 +24,15 @@ var GetListingRequest []byte = []byte{0,0,0,60,0,0,0,2,
 	105,110,103,0,12,47,117,115,101,114,47,104,100,117,
 	115,101,114,0,2,91,66,0,0,0,0,0}
 
-var ProtocolRequest []byte = []byte{0,0,0,57,46,111,114,103,46,97,112,97,99,104,101,46,104,97,100,111,111,112,46,104,100,102,115,46,112,114,111,116,111,99,111,108,46,67,108,105,101,110,116,80,114,111,116,111,99,111,108,1,0,6,104,100,117,115,101,114,0,0,0,0,108,0,0,0,0,0,18,103,101,116,80,114,111,116,111,99,111,108,86,101,114,115,105,111,110,0,0,0,2,0,16,106,97,118,97,46,108,97,110,103,46,83,116,114,105,110,103,0,46,111,114,103,46,97,112,97,99,104,101,46,104,97,100,111,111,112,46,104,100,102,115,46,112,114,111,116,111,99,111,108,46,67,108,105,101,110,116,80,114,111,116,111,99,111,108,0,4,108,111,110,103,0,0,0,0,0,0,0,61,0}
+var ProtocolRequest []byte = []byte{0,0,0,57,46,111,114,103,46,97,112,
+97,99,104,101,46,104,97,100,111,111,112,46,104,100,102,115,46,112,114,
+111,116,111,99,111,108,46,67,108,105,101,110,116,80,114,111,116,111,99,
+111,108,1,0,6,104,100,117,115,101,114,0,0,0,0,108,0,0,0,0,0,18,103,101,
+116,80,114,111,116,111,99,111,108,86,101,114,115,105,111,110,0,0,0,2,0,
+16,106,97,118,97,46,108,97,110,103,46,83,116,114,105,110,103,0,46,111,
+114,103,46,97,112,97,99,104,101,46,104,97,100,111,111,112,46,104,100,
+102,115,46,112,114,111,116,111,99,111,108,46,67,108,105,101,110,116,
+80,114,111,116,111,99,111,108,0,4,108,111,110,103,0,0,0,0,0,0,0,61,0}
 
 var InitiateRequest []byte = []byte{104,114,112,99,4,80,0}
 

### 2014-05-13 src/util/logging.go

@@ -39,8 +39,10 @@ var TempLogger *log.Logger
 //cleared at every run
 var TempLoggerLogFile = "../../logs/temp.log"
 
-//NOTE this is a relative path; in deployment, the executable needs to be in the same directory
-//this file otherwise there will be problems in loading the logging configuration (it will
+//NOTE this is a relative path; in deployment, the executable needs to
+// be in the same directory
+//this file otherwise there will be problems in loading the logging 
+//configuration (it will
 //just default to the development values).
 var LoggingConfFile = "logging.json"
 
@@ -55,7 +57,6 @@ func InitLoggingConfiguration() error {
 }
 
 func InitTempLogger() error {
-	//tempLogFile, err := os.OpenFile(TempLoggerLogFile, os.O_WRONLY | os.O_CREATE, 0666)
 	tempLogFile, err := os.Create(TempLoggerLogFile)
 	
 	if err != nil {
@@ -67,7 +68,8 @@ func InitTempLogger() error {
 }
 
 func InitDataReqLogger() error {
-	dataLogFile, err := os.OpenFile(DataReqLogFile, os.O_RDWR | os.O_CREATE | os.O_APPEND, 0666)
+	dataLogFile, err := os.OpenFile(DataReqLogFile, os.O_RDWR 
+	| os.O_CREATE | os.O_APPEND, 0666)
 	if err != nil {
 		return err
 	}
@@ -86,7 +88,8 @@ func InitDebugLogger() error {
 }
 
 func InitNoCacheLatencyLogger() error {
-	noCacheLogFile, err := os.OpenFile(NoCacheLatencyLogFile, os.O_RDWR | os.O_CREATE | os.O_APPEND, 0666)
+	noCacheLogFile, err := os.OpenFile(NoCacheLatencyLogFile, os.O_RDWR 
+	| os.O_CREATE | os.O_APPEND, 0666)
 	if err != nil {
 		return err
 	}
@@ -98,7 +101,8 @@ func InitNoCacheLatencyLogger() error {
 }
 
 func InitCachedLatencyLogger() error {
-	cachedLogFile, err := os.OpenFile(CachedLatencyLogFile, os.O_RDWR | os.O_CREATE | os.O_APPEND, 0666)
+	cachedLogFile, err := os.OpenFile(CachedLatencyLogFile, os.O_RDWR 
[…]

### 2014-05-13 src/writable_processor/writable_processor.go

@@ -181,7 +181,8 @@ func (w *WritableProcessor) handleReadBlockResponse(
 	}
 }
 
-func (w *WritableProcessor) processReadBlock(requestHeader *writables.DataRequestHeader, 
+func (w *WritableProcessor) processReadBlock(
+requestHeader *writables.DataRequestHeader, 
 	conn writables.ReaderWriter, dataNode writables.ReaderWriter) {
 
 	blockRequest, err := w.readReadBlockRequest(conn)
@@ -213,7 +214,8 @@ func (w *WritableProcessor) processReadBlock(requestHeader *writables.DataReques
 }
 
 //this method is called from generalProcessing()
-func (w *WritableProcessor) handleGeneralResponse(conn writables.ReaderWriter, dataNode writables.ReaderWriter) {
+func (w *WritableProcessor) handleGeneralResponse(
+conn writables.ReaderWriter, dataNode writables.ReaderWriter) {
 	for {
 		//check the channel to make sure that the socket isn't closed
 		msg := w.readComm()
@@ -242,7 +244,9 @@ func (w *WritableProcessor) handleGeneralResponse(conn writables.ReaderWriter, d
 	}
 }
 
-func (w *WritableProcessor) forwardRequestHeader(requestHeader *writables.DataRequestHeader, dataNode writables.ReaderWriter) error {
+func (w *WritableProcessor) forwardRequestHeader(
+requestHeader *writables.DataRequestHeader, dataNode writables.ReaderWriter) 
+error {
 	resBuf := new(bytes.Buffer)
 	err := requestHeader.Write(resBuf)
 	if err != nil {
@@ -266,7 +270,9 @@ func (w *WritableProcessor) GeneralProcessing(conn net.Conn,
 //a simple relay is set up so that data that comes in is sent directly
 //to the DataNode without modification. It also starts (as a goroutine)
 //handleGeneralResponse() that relays packets from the DataNode to the client
-func (w *WritableProcessor) generalProcessing(conn writables.ReaderWriter, dataNode writables.ReaderWriter, replaceResponse bool) {
+func (w *WritableProcessor) generalProcessing(
+conn writables.ReaderWriter, dataNode writables.ReaderWriter, 
+replaceResponse bool) {
 	util.TempLogger.Println(w.id, "General processing called.")
 
 	//start the response method
@@ -286,7 +292,8 @@ func (w *WritableProcessor) generalProcessing(conn writables.ReaderWriter, dataN
[…]

### 2014-05-13 src/writables/data_transfer_protocol.go

@@ -575,7 +575,8 @@ func (header *BlockResponseHeader) AdjustChecksumSize(dataLen int64) int64 {
 	bytesPerChecksum := int64(header.Checksum.BytesPerChecksum)
 	checksumSize := int64(header.Checksum.Size())
 
-	requiredSize := ((dataLen + bytesPerChecksum - 1)/bytesPerChecksum)*checksumSize
+	requiredSize := ((dataLen + bytesPerChecksum - 1)/bytesPerChecksum)*
+	checksumSize
 	return requiredSize
 }
 

### 2014-05-13 src/writables/data_transfer_protocol_test.go

@@ -90,7 +90,8 @@ func TestReadRequestHeaderEquals(t *testing.T) {
 
 	if !r1.Equals(r2) {
 		fmt.Println("R1 wasn't equal to R2 when it was supposed to.")
-		fmt.Println("Client names equal? : ", r1.ClientName.Equals(r2.ClientName))
+		fmt.Println("Client names equal? : ", r1.ClientName.Equals(
+		r2.ClientName))
 		t.Fail()
 	}
 

### 2014-05-13 src/writables/writables.go

@@ -242,7 +242,9 @@ func GenericRead(packet interface{}, reader Reader) error {
 		//Hadoop's protocol (if it isn't, don't use GenericRead)
 		case []byte:
 			if i <= 0 {
-				return errors.New("There is a []byte as the first element of the packet structure; don't know the length, so cannot proceed.")
+				return errors.New("There is a []byte as the first element 
+				of the packet structure; don't know the length, so cannot 
+				proceed.")
 			}
 
 			length := reflect.ValueOf(packet).Elem().Field(i-1).Interface()
@@ -543,7 +545,8 @@ func WriteVInt(val int64, writer Writer) {
   for idx := length; idx != 0; idx-- {
   	shiftbits := (idx - 1) * 8
   	mask := 0xFF << uint(shiftbits)
-  	binary.Write(writer, binary.BigEndian, int8((val & int64(mask)) >> uint(shiftbits)))
+  	binary.Write(writer, binary.BigEndian, int8((val & int64(mask)) >> 
+  	uint(shiftbits)))
   }
 }
 

### 2014-05-13 src/writables/writables_test.go

@@ -9,10 +9,14 @@ import (
 )
 
 //test case
-var DataNodeRegistrationTest []byte = []byte{0,24,100,104,97,105,118,97,116,45,71,65,45,56,55,
-	48,65,45,85,68,51,58,50,48,49,48,0,42,68,83,45,50,48,57,54,56,50,54,49,51,54,45,49,50,55,46,
-	48,46,49,46,49,45,50,48,49,48,45,49,51,57,53,50,48,53,55,51,57,56,51,56,195,155,195,100,255,
-	255,255,215,108,110,46,95,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,255,0,0,0,0}
+var DataNodeRegistrationTest []byte = []byte{0,24,100,104,97,105,118,
+	97,116,45,71,65,45,56,55,
+	48,65,45,85,68,51,58,50,48,49,48,0,42,68,83,45,50,48,57,54,56,50,
+	54,49,51,54,45,49,50,55,46,
+	48,46,49,46,49,45,50,48,49,48,45,49,51,57,53,50,48,53,55,51,57,56,
+	51,56,195,155,195,100,255,
+	255,255,215,108,110,46,95,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,
+	0,0,0,0,0,0,0,255,0,0,0,0}
 
 var dnrBuffer *bytes.Buffer
 
@@ -304,7 +308,8 @@ func TestDNRReadIsBlockTokenEnabled(t *testing.T) {
 	err := dnr.ReadIsBlockTokenEnabled(dnrBuffer)
 	if err != nil {
 		t.Fail()
-		fmt.Println("Error occurred in TestDNRReadIsBlockTokenEnabled: ", err)
+		fmt.Println("Error occurred in TestDNRReadIsBlockTokenEnabled: ", 
+		err)
 	}
 
 	//should be false

### 2014-05-13 paper/paper.tex

@@
 In particular, the Hadoop distributed system \cite{hadoop} has seen tremendous growth.
-Hadoop is an open source version of the revolutionary MapReduce system developed at Google \cite{mapreduce}.
+Hadoop is an open source implementation of the revolutionary MapReduce system developed at Google \cite{mapreduce}.
 With it, developers can easily take advantage of large, multi-node clusters to solve computational problems.
 
+Hadoop consists of a computational layer termed \textit{MapReduce} and a custom distributed filesystem named \textit{Hadoop Distributed File System (HDFS)}.
+HDFS is used to allow all nodes in a cluster access to the same dataset.
+For instance, one may consider running an algorithm that counts the frequency of each word in a large text file.
+If this algorithm is run on a Hadoop cluster, each node must have access to the file in question in order to determine word frequency.
+
 Current Hadoop deployments exist in the thousands of nodes at companies such as Cloudera, Facebook, Yahoo, etc.
-Especially in such large networks, \textit{access latency}, or the waiting time associated with accessing files through the Hadoop Distributed File System (\textit{HDFS}), is a significant concern.
+Especially in such large networks, \textit{access latency}, or the waiting time associated with accessing files and related information through the Hadoop Distributed File System (\textit{HDFS}), is a significant concern \cite{hdfs}.
 Latency in distributed systems can have significant effects and a reduction of the same comes with tremendous benefits.
 As the RAMCloud project has outlined \cite{ramcloud}, low latency can greatly extend the applications of distributed computing systems.
-For example, tree traversal algorithms are largely impractical in high latency systems such as Hadoop \cite{ramcloud}, however, with lowered latency, such algorithms may become useful and applicable.
+For example, algorithms that require files other than the input file are largely impractical in high latency systems such as Hadoop \cite{ramcloud, spark}.
+A similar case exists with jobs (or sets of jobs) that consist of multiple passes over the same dataset; many of these exist in the field of machine learning, analytics, etc.
+\cite{spark}.
+However, with lowered latency, such algorithms may become useful and applicable.
 HDFS currently does not use local (i.e.
@@
 
-In this paper, I present \textit{Panthera}, a \textit{cache layer} and \textit{cache-based scheduling system} for Hadoop.
-The \textit{Panthera} cache layer functions by retaining recently accessed files in random access memory, thus, when the files are needed once again, they do not have to be downloaded or read from the hard drive, thereby greatly reducing file access latency.
+I present \textit{Panthera}, a \textit{cache layer} and \textit{cache-based scheduling system} for Hadoop.
+The \textit{Panthera} cache layer functions by retaining recently accessed files and related information in random access memory.
+Thus, when the files are needed once again, they do not have to be downloaded or read from the hard drive, thereby greatly reducing file access latency.
 Another portion of the project, the \textit{Panthera} scheduling system utilizes information about the data blocks cached in a cluster at a given time to schedule jobs.
@@
 
-In this paper, we discuss both the \textit{Panthera} caching layer and scheduling system.
-In Section 2, we discuss the specific challenges associated with such a project.
-Section 3 details the architecture differences between a standard Hadoop installation and one with \textit{Panthera}.
-Sections 4 and 5 detail the differences between metadata and data caching within \textit{Panthera}.
-Results and conclusions are covered in Sections 6, 7, 8, 9, concluding with a roadmap for future development in Section 10.
+In this paper, the \textit{Panthera} caching layer and scheduling system are presented.
+In Section \ref{sect:architecture}, motivation and construction of the \textit{Panthera} architecture are explained.
+Section~\ref{sect:caching_layer} describes the caching layer portion of this project.
+Section~\ref{sect:scheduler} presents the scheduling algorithm and implementation details.
+The results and data are noted in Section~\ref{sect:results}.
+Finally, Sections \ref{sect:conclusions}, \ref{sect:applications}, \ref{sect:future_work} conclude with an interpretation of the results, discussion of potential applications and future work.
 
-\section{Design Constraints}
+\section{Architecture} \label{sect:architecture}
+\subsection{Design Constraints}  \label{design_constraints}
+The primary goal of the \textit{Panthera} project is to reduce access latency in HDFS.
+Therefore, data and metadata request latency should be significantly lowered with \textit{Panthera} in comparison to the control group of a standard Hadoop installation.
 
-There are several systems built on Hadoop that are in widespread use \cite{hbase, cloudbatch, pig}. The Hadoop project is also supported by large corporations with a 
+There are many systems built on Hadoop that are in widespread use \cite{hbase, cloudbatch, pig}. The Hadoop project is also supported by large corporations with a 
 consistent release cycle.
-Thus, for \textit{Panthera} to be practical, it must operate independently of the existing Hadoop codebase, i.e.
+Thus, for \textit{Panthera} to be practical, it must opedrate independently of the existing Hadoop codebase, i.e.
 as a layer, rather than modification.
 This implies that \textit{Panthera} must implemented without a single change in the existing Hadoop codebase, i.e.
-as a "drop-in" solution.
-Additionally, for non-cache related requests, \textit{Panthera} must add insignificant latency.
-Finally, data and metadata request latency should be significantly lowered with \textit{Panthera} in comparison to a standard Hadoop installation.
+as a ``drop-in'' solution.
+This is due to the fact that a deviation from the mainline Hadoop codebase would have to updated at the same pace as the vanilla Hadoop, which would be contrary the benefits of Hadoop's large and active development community.
+Also, since \textit{Panthera} does not require any code modifications to existing Hadoop applications, allowing it to be easily integrated into current software stacks.
 
@@
 blocks currently cached) over the network.
-The cache-based scheduler then uses information about blocks cached and the files each job requires to run in order to optimize the running order of a set of jobs (see Section \ref{scheduler-section} for a complete description of the \textit{Panthera} scheduling algorithm).
+The cache-based scheduler then uses information about blocks cached and the files each job requires to run in order to optimize the running order of a set of jobs (see Section \ref{sect:scheduling} for a complete description of the \textit{Panthera} scheduling algorithm).
 
-\section{Metadata and Data}
+\subsection{Metadata and Data} \label{metadata_and_data}
 
-Within the Hadoop File System (HDFS), there is a clear distinction between \textit{metadata} and \textit{data} \cite{hdfs}.
+Within the Hadoop File System (HDFS), there is a clear architectural distinction between \textit{metadata} and \textit{data} \cite{hdfs}.
 Metadata refers to information about a file or a directory, e.g.
-file size, file names, on which server the contents of a file are located, etc.
+file size, file names, the location of a file in the cluster, etc.
 For HDFS, the NameNode server handles all metadata.
@@
 
-\section{Architecture}
-\begin{figure}[!h]
+\subsection{Panthera Architecture} \label{architecture_sect}
+\begin{figure}
+\centering
+\begin{subfigure}{.5\textwidth}
   \centering
   \includegraphics[scale=0.4]{assets/hadoop_architecture.pdf}
+  \caption{Vanilla Hadoop architecture}
+  \label{fig:vanilla_arch}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
   \includegraphics[scale=0.4]{assets/panthera_architecture.pdf}
-  \caption{Standard Hadoop Architecture (left), \textit{Panthera}-Hadoop Architecture (right)}
-  \label{arch}
-
+  \caption{Panthera-Hadoop architecture}
+  \label{fig:panthera_arch}
+\end{subfigure}
+\label{fig:arch}
 \end{figure}
 
-Figure~\ref{arch} outlines the architecture differences between a standard Hadoop File System architecture and an installation with \textit{Panthera}.
+Figure~\ref{architecture_sect} outlines the architecture differences between a standard Hadoop File System architecture and an installation with \textit{Panthera}.
 In HDFS, the \textit{NameNode} serves as a metadata server, handling all requests from clients relating to metadata (e.g.
@@
 \textit{DataNode(s)} hold data and serve it at request from the client.
-A more complete description of the Hadoop File System may be found in \cite{hadoop}.
+In order to read or write a file, a client first obtains permission from the NameNode and subsequently communicates with a DataNode to complete the operation.
+Though the aforementioned background is sufficient for this paper, a more complete description of the Hadoop File System may be found in \cite{hadoop}.
 
@@
 Every request issued from the client is first reviewed by \textit{Panthera}, and if possible, answered immediately from the cache.
-The \textit{Panthera} instance on the NameNode constantly monitors the filesystem for changes that need to be reflected in the cache (e.g.
-if a file is deleted, the cache on the client should not have a copy of it).
+If the response is not found in the cache, certain requests are cached and subsequent responses are associated with the request in consideration.
 
-\textit{Panthera}, since it caches both data and metadata, can be considered two separate subsystems, which are now discussed.
+The \textit{Panthera} instance on the NameNode serves as a cache refresh mechanism.
+Though sparse in the Hadoop environment since input files are usually not written to in the MapReduce computational paradigm, files that are cached may become out-of-date, i.e.
+the cache becomes "stale".
+The \textit{Panthera} instance on the NameNode monitors the filesystem for changes and reports certain changes to Panthera caches, which then remove stale files from the cache.
+The refresh mechanism is particularly important for metadata caching where the cached information is far more likely to change during the execution of a job.
 
-\section{Metadata caching}
+The \textit{Panthera} cache layer, since it caches both data and metadata, can be considered two separate subsystems, which are now discussed.
+
+\section{Panthera Caching Layer} \label{sect:caching_layer}
+A significant portion of Panthera is a caching layer for Hadoop, the details of which are now covered. 
+\subsection{Metadata caching} \label{sect:metadata_caching}
 
@@
 The \textit{Panthera} metadata system runs on the NameNode and communicates with the \textit{Panthera} instance on the client.
-The challenge on the metadata portion consist of reducing computational latency to memory latency.
+The challenge on the metadata portion consists of converting computational latency to memory latency.
 For example, for some intensive methods (e.g.
@@
 Obviously, there are some methods for which caching is not worthwhile since their execution time is lower or comparable to memory latency.
-Within \textit{Panthera}, six different Hadoop query methods were identified as medium to high latency.
-\textit{Panthera} identifies packets that contain these methods and subsequently caches them.
+Within Hadoop, there are several methods that incur very significant computational latency, e.g.
+a recursive directory listing.
+Only these methods are cached within Panthera.
 
-Currently, a standard LRU cache \cite{cache} is used for each method. Testing of different caching methods and predictive prefetching algorithms is under way.
+Currently, a standard LRU cache \cite{cache} is used for each method. Testing of different caching methods and predictive prefetching algorithms \cite{griffioen} is under way.
 
-\section{Data caching}
+\subsection{Data caching} \label{sect:data_caching}
 
@@
 As \cite{ramcloud, hadoop} note, memory latency can be extremely low in modern systems, especially in comparison to network and hard drive latencies.
-Thus, in some respects, the goal of lowered latency is "easier" to attain in the data cache.
-However, for smaller files, it may be difficult to obtain latency improvement.
+Thus, in some respects, the goal of lowered latency is ``easier'' to attain in the data cache in comparison to the metadata cache.
 
+\section{Scheduler} \label{sect:scheduler}
+\subsection{Hadoop Job Structure}
 Hadoop, in professional and academic environments, is often used with series rather than individual jobs.
+For example, in machine learning, various algorithms (termed \textit{iterative machine learning algorithms}) consist of multiple Hadoop jobs rather than a single job \cite{spark}.
+In Hadoop-based querying systems \cite{hive, pig}, a set of queries may consist of jobs that are entirely independent of one-another.
 Deciding how to order these jobs depending on cached file availability is a problem of great interest; \textit{Panthera} presents the solution of a cache-based scheduler for Hadoop.
 
-\section{Scheduling Algorithm} \label{scheduler-section}
+\subsection{Cluster Saturation}
+It may be asked why it is not possible to run independent jobs in parallel on the Hadoop cluster.
+Each node in a given cluster has a maximum number of \textit{map} or \textit{reduce} calls (i.e.
+the basic computational units in Hadoop) it may process at a given time.
+If a job saturates nodes in a cluster, running another job concurrently would provide no running time benefit.
+Thus, a scheduler is useful since it may organize independent jobs in a beneficial ordering.
+In \textit{Panthera}, this ordering is based on the files or blocks present in the cluster's caches.
+
+\subsection{Scheduling Algorithm} \label{sect:scheduling}
 
@@
 
-\section{Testing Methodology}
+\section{Results} \label{sect:results}
 
+\subsection{Testing Methodology}
 In order to test the metadata cache, a directory listing query was repeatedly run on a directory with 100 files in it, and the latency times for responding are measured, with \textit{Panthera} and without.
@@
 
-\section{Results}
+\subsection{Results}
 \begin{figure}
@@
 
-\section{Conclusion}
+\section{Conclusions}\label{sect:conclusions}
 
@@
 
-\section{Applications}
+\section{Applications}\label{sect:applications}
 One of the major benefits of the systems developed is that they are independent of the existing Hadoop codebase and they do not require modifications to existing Hadoop applications.
@@
 
-\section{Future work}
+\section{Future work}\label{sect:future_work}
 One of the most immediate avenues for further research is in \textit{domain-specific prefetching}.
